---
title: "Honda et al. (2019): Updated Armitage et al. (2014) Model"
author: "Gregory S. Honda"
date: "July 13, 2018"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Honda et al. (2019): Updated Armitage et al. (2014) Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}---

```{r setup, include=FALSE, EVAL=F}
knitr::opts_chunk$set(echo = TRUE)
```

## Updated Armitage 2014 Top Panel Figure 3

This produces an updated version of the top panel of Figure 3 from the paper Armitage, J. M.; Wania, F.; Arnot, J. A. Environ. Sci. Technol. 2014, 48, 9770-9779.

```{r, eval=F}
library(data.table)
library(magrittr)
library(ggplot2)
library(httk)
armitage.dt <- copy(armitage_input)
armitage.dt[,well_number:=384] %>% 
  .[,option.bottom:=TRUE] %>% 
  .[,option.plastic:=TRUE] %>% 
  .[,Tsys:=37] %>% 
  .[,Tref:=298.15] %>% 
  .[,FBSf:=0.1] %>% 
  .[,nomconc:=50]

armitage.dt2 <- armitage_estimate_sarea(tcdata = armitage.dt)
armitage_output1 <- armitage_eval(tcdata = armitage.dt2)
armitage_output2 <- armitage_eval(tcdata = armitage.dt2[,ac50:=1])
armitage_output3 <- armitage_eval(tcdata = armitage.dt2[,ac50:=0.001])



```

## Plot results

```{r, echo=TRUE, eval=F}
armitage_output <- rbind(armitage_output1[,xfill:="50 \U00B5M"],
                         armitage_output2[,xfill:="1 \U00B5M"],
                         armitage_output3[,xfill:="1 nM"])
armitage_output[,xfill:=factor(xfill,levels=c("50 \U00B5M","1 \U00B5M","1 nM"))]

ggplot(armitage_output) +
  geom_point(aes(x=gkow,y=log10(ccells/MW*1e9),color=xfill),shape=1) +
  labs(color=expression("AC"["50"]),x="log P", y = "log(ng chemical/g cells)") +
  theme_bw()
```


