---
title: "Pearce et al. (2017) Figures in Partition Coefficient Evaluation"
author: "Robert Pearce"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pearce et al. (2017): Creating Partition Coefficient Evaluation Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette generates the data and figures from "Evaluation and calibration 
of high-throughput predictions of chemical distribution to tissues".

Robert G. Pearce, R. Woodrow Setzer, Jimena L. Davis & John F. Wambaugh

Journal of Pharmacokinetics and Pharmacodynamics volume 44, pages549-565 (2017)

https://dx.doi.org/10.1007%2Fs10928-017-9548-7

## Abstract
Toxicokinetics (TK) provides critical information for integrating chemical 
toxicity and exposure assessments in order to determine potential chemical 
risk (i.e., the margin between toxic doses and plausible exposures). For 
thousands of chemicals that are present in our environment, in vivo TK 
data are lacking. The publicly available R package "httk" (version 1.8, 
named for "high throughput TK") draws from a database of in vitro data and 
physico-chemical properties in order to run physiologically-based TK (PBTK) 
models for 553 compounds. The PBTK model parameters include tissue:plasma 
partition coefficients (Kp) which the httk software predicts using the 
model of Schmitt (Toxicol In Vitro 22 (2):457-467, 2008). In this paper we 
evaluated and modified httk predictions, and quantified confidence using 
in vivo literature data. We used 964 rat Kp measured by in vivo experiments 
for 143 compounds. Initially, predicted Kp were significantly larger than 
measured Kp for many lipophilic compounds (log10 octanol:water partition 
coefficient > 3). Hence the approach for predicting Kp was revised to 
account for possible deficiencies in the in vitro protein binding assay, 
and the method for predicting membrane affinity was revised. These changes 
yielded improvements ranging from a factor of 10 to nearly a factor of 
10,000 for 83 Kp across 23 compounds with only 3 Kp worsening by more than 
a factor of 10. The vast majority (92%) of Kp were predicted within a 
factor of 10 of the measured value (overall root mean squared error of 
0.59 on log10-transformed scale). After applying the adjustments, 
regressions were performed to calibrate and evaluate the predictions for 
12 tissues. Predictions for some tissues (e.g., spleen, bone, gut, lung) 
were observed to be better than predictions for other tissues (e.g., skin, 
brain, fat), indicating that confidence in the application of in silico 
tools to predict chemical partitioning varies depending upon the tissues 
involved. Our calibrated model was then evaluated using a second data set 
of human in vivo measurements of volume of distribution (Vss) for 498 
compounds reviewed by Obach et al. (Drug Metab Dispos 36(7):1385-1405, 
2008). We found that calibration of the model improved performance: a 
regression of the measured values as a function of the predictions has a 
slope of 1.03, intercept of ??? 0.04, and R2 of 0.43. Through careful 
evaluation of predictive methods for chemical partitioning into tissues, 
we have improved and calibrated these methods and quantified confidence 
for TK predictions in humans and rats.
```{r knitrPRrep, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>', fig.width=4, fig.height=3)
```
```{r InitialPrep, eval=TRUE}
library(httk)
library(gdata)
library(ggplot2)
library(viridis)
library(CensRegMod)
library(gmodels)
library(gplots)
library(scales)
library(colorspace)
# We love to give warning messages whenever assumptions are used by HTTK,
# but they will overwhelm the output of this vignette so we turn them
# off:
options(warn = -1)
```
We first filter the measured rat Kp data, pc.data. Then the old and new Kp predictions are made, along with error and improvement measures, 
and these are all consolidated into a table for analysis and plotting. Note that the final table contains log10-transformed values and error 
and improvements derived from subtracting these values. Only relevant rat values are used. Compounds with Funbound.plasma and partition 
coefficients of zero are removed as well as compounds with approximated Funbound.plasma values.

```{r FilterData}
pc.table <- NULL
pc.data <- subset(httk::pc.data,fu != 0 & Exp_PC != 0 & Tissue %in% c("Adipose","Bone","Brain","Gut",
    "Heart","Kidney","Liver","Lung","Muscle","Skin","Spleen","Blood Cells") & 
    tolower(Species) == 'rat' & !CAS %in% c('10457-90-6','5786-21-0','17617-23-1','69-23-8','2898-12-6',
    '57562-99-9','59-99-4','2955-38-6','155-97-5','41903-57-5','58-55-9','77-32-7','59-05-2','60-54-8'))
cas.list <- get_cheminfo(model='schmitt',species='rat')
cas.list <-  cas.list[cas.list %in% pc.data[,'CAS']]
ma.data.list <- subset(chem.physical_and_invitro.data,!is.na(logMA))[,'CAS']
for(this.cas in cas.list){
  parameters <- parameterize_schmitt(chem.cas=this.cas,species='rat')
  init.parameters <- parameters
  charge <- calc_ionization(chem.cas=this.cas,pH=7.4)$fraction_charged
  if(!this.cas %in% ma.data.list){
    init.parameters$MA <- 10^(0.999831 - 0.016578*38.7 + 0.881721 * log10(parameters$Pow))
  }
  pcs <- predict_partitioning_schmitt(parameters=parameters,species='rat',regression=F)
  init.pcs <- predict_partitioning_schmitt(parameters=init.parameters,species='rat',regression=F)
  for(this.tissue in subset(pc.data,CAS==this.cas)[,'Tissue']){
    if(this.tissue == 'Blood Cells') this.pc <- 'rbc'
    else this.pc <- this.tissue
    pc.table <- rbind(pc.table,cbind(as.data.frame(this.cas),as.data.frame(this.tissue),
    as.data.frame(log10(init.pcs[[which(substr(names(init.pcs),2,nchar(names(init.pcs))-3)
                                  == tolower(this.pc))]] * init.parameters$Funbound.plasma)),
    as.data.frame(log10(pcs[[which(substr(names(pcs),2,nchar(names(pcs))-3)
                             == tolower(this.pc))]] * parameters$unadjusted.Funbound.plasma)),
    as.data.frame(log10(init.pcs[[which(substr(names(init.pcs),2,nchar(names(init.pcs))-3)
                                  == tolower(this.pc))]] * init.parameters$unadjusted.Funbound.plasma)),
    as.data.frame(log10(pcs[[which(substr(names(pcs),2,nchar(names(pcs))-3)
                             == tolower(this.pc))]] * parameters$Funbound.plasma)),
    as.data.frame(log10(subset(pc.data,CAS==this.cas & Tissue==this.tissue)[,'Exp_PC'])),
    as.data.frame(subset(pc.data,CAS==this.cas & Tissue==this.tissue)[,'LogP']),as.data.frame(charge),
    as.data.frame(as.character(subset(pc.data,CAS == this.cas)[1,'A.B.N'])),
    as.data.frame(subset(pc.data,CAS == this.cas)[1,'fu'])))
  }
}
colnames(pc.table) <- c('CAS','Tissue','fup.adjustment','ma.adjustment','init.Predicted',
                        'Predicted','Experimental','logP','charge','type','fup')
init.error <- pc.table[,'Experimental'] - pc.table[,'init.Predicted']
fup.error <- pc.table[,'Experimental'] - pc.table[,'fup.adjustment']
ma.error <- pc.table[,'Experimental'] - pc.table[,'ma.adjustment']
final.error <- pc.table[,'Experimental'] - pc.table[,'Predicted']
fup.improvement <- abs(init.error) - abs(fup.error)
ma.improvement <- abs(init.error) - abs(ma.error)
final.improvement <- abs(init.error) - abs(final.error)
pc.table <- cbind(pc.table,fup.improvement,ma.improvement, final.improvement,
                  final.error,init.error,ma.error,fup.error)

scientific_10 <- function(x) {                                  
  out <- gsub("1e", "10^", scientific_format()(x))              
  out <- gsub("\\+","",out)                                     
  out <- gsub("10\\^01","10",out)                               
  out <- parse(text=gsub("10\\^00","1",out))                    
}  


init.plot <- ggplot() + geom_point(data=pc.table,aes(10^(init.Predicted),10^(Experimental))) +
    geom_abline() + labs(y=expression(paste("Measured ",K[p])), x=expression(paste("Predicted ",K[p]))) +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16),
    plot.title=element_text(size=18,hjust = 0.5)) + scale_x_log10(label=scientific_10,limits=c(0.01,10^4.5)) +
    scale_y_log10(label=scientific_10,limits=c(0.01,10^4.5)) + ggtitle('(A)')
print(init.plot)

final.plot <- ggplot() + geom_point(data=pc.table,aes(10^(Predicted),10^(Experimental))) +
    geom_abline() + labs(y=expression(paste("Measured ",K[p])),x=expression(paste("Predicted ",K[p]))) +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16),
    plot.title=element_text(size=18,hjust=0.5)) + scale_x_log10(label=scientific_10,limits=c(0.01,10^4.5)) +
    scale_y_log10(label=scientific_10,limits=c(0.01,10^4.5)) + ggtitle('(B)')
print(final.plot)

fup.change.plot <-  ggplot() +
    geom_point(data=pc.table[order(pc.table[,'fup.improvement'],decreasing=F),],
    aes(10^(fup.adjustment),10^(Experimental),color=fup.improvement)) + geom_abline() +
    labs(y=expression(paste("Measured ",K[p])),x=expression(paste("Predicted ",K[p])),color='Improvement') +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16)) +
    scale_x_log10(label=scientific_10,limits=c(0.01,10^4.5)) + 
    scale_y_log10(label=scientific_10,limits=c(0.01,10^4.5)) +
    scale_color_viridis(direction=-1,option='inferno')
print(fup.change.plot)

ma.subset <- subset(pc.table,!CAS %in% ma.data.list)
ma.change.plot <- ggplot() +
    geom_point(data=ma.subset[order(ma.subset[,'ma.improvement'],decreasing=F),]
    ,aes(10^(ma.adjustment),10^(Experimental),color=ma.improvement)) + geom_abline() +
    labs(y=expression(paste("Measured ",K[p])),x=expression(paste("Predicted ",K[p])),color='Improvement') +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16)) +
    scale_x_log10(label=scientific_10,limits=c(0.01,10^4.5)) + 
    scale_y_log10(label=scientific_10,limits=c(0.01,10^4.5)) +
    scale_color_viridis(direction=-1,option='inferno')
print(ma.change.plot)
```
Now we calculate and plot the regressions for all tissues, together with their 95% confidence intervals.

```{r Regressions}
regressions <- NULL

for(tissue in as.character(unique(pc.table[,'Tissue']))){
  fit <- lm(Experimental ~ Predicted ,data=subset(pc.table,Tissue==tissue))
  smry <- summary(fit)
  est <- estimable(fit, cm=diag(2), beta0=c(0,1), joint.test=TRUE)
  regressions <- rbind(regressions,cbind(tissue,as.data.frame(fit$coefficients[['(Intercept)']]),
      as.data.frame(fit$coefficients[['Predicted']]),
      as.data.frame(smry$coefficients[['Predicted','Pr(>|t|)']]),
      as.data.frame(smry$sigma),as.data.frame(smry$r.squared),
      as.data.frame(smry[[11]][1,1]),as.data.frame(smry[[11]][2,2]),
      as.data.frame(smry[[11]][1,2]),as.data.frame(smry$df[2]),as.data.frame(est[[3]])))
}
colnames(regressions) <- c('Tissue','Intercept','Slope','P-value','SE','R-squared',
                           'Int Var','Slp Var','Cov','df','estimable')
x.cf <- seq(-2,3.5,.01)
for(tissue in  as.character(unique(pc.table[,'Tissue']))){
  conf <- qt(0.975,df=subset(regressions,Tissue==tissue)[['df']]+1) *
      subset(regressions,Tissue==tissue)[['SE']] *
      sqrt(subset(regressions,Tissue==tissue)[['Int Var']] +
      x.cf^2 * subset(regressions,Tissue==tissue)[['Slp Var']] +
      2 * x.cf * subset(regressions,Tissue==tissue)[['Cov']] + 1)
  line <- subset(regressions,Tissue==tissue)[['Intercept']] +
      x.cf * subset(regressions,Tissue==tissue)[['Slope']]
  y.cf <- line + conf
  y.ncf <- line - conf

  cf <- cbind(as.data.frame(x.cf),as.data.frame(y.cf),as.data.frame(y.ncf))
  if(tissue == 'Blood Cells'){
    eval(parse(text= paste('Blood <- ggplot() + labs(y=expression(paste("Inferred ",K[p])),
                               x=expression(paste("Predicted ",K[p]))) + geom_abline(linetype = "dashed") +
                               geom_point(data=subset(pc.table,Tissue == \'',tissue,'\'),aes(10^(Predicted),
                               10^(Experimental)))  +  theme(axis.text=element_text(size=14),
                               axis.title=element_text(size=14),plot.title=element_text(size=14,hjust=0.5)) +
                               scale_x_log10(label=scientific_10,limits=c(0.01,1000)) + 
                               scale_y_log10(label=scientific_10,limits=c(0.01,1000)) +
                               geom_line(data=cf,aes(10^(x.cf),10^(y.cf))) +
                               geom_line(data=cf,aes(10^(x.cf),10^(y.ncf))) +
                               geom_abline(intercept=subset(regressions,Tissue==tissue)[[\'Intercept\']],
                               slope=subset(regressions,Tissue==tissue)[[\'Slope\']]) +
                               ggtitle(\'Red Blood Cells\')',sep='')))
  }else{
    eval(parse(text= paste(tissue,' <- ggplot() + labs(y=expression(paste("Measured ",K[p]))
                               ,x=expression(paste("Predicted ",K[p]))) + geom_abline(linetype = "dashed") +
                               geom_point(data=subset(pc.table,Tissue == \'',tissue,'\'),
                               aes(10^(Predicted),10^(Experimental))) + theme(axis.text=element_text(size=14),
                               axis.title=element_text(size=14),plot.title=element_text(size=14,hjust=0.5)) +
                               scale_x_log10(label=scientific_10,limits=c(0.01,1000)) + 
                               scale_y_log10(label=scientific_10,limits=c(0.01,1000)) +
                               geom_line(data=cf,aes(10^(x.cf),10^(y.cf))) +
                               geom_line(data=cf,aes(10^(x.cf),10^(y.ncf))) +
                               geom_abline(intercept=subset(regressions,Tissue==tissue)[[\'Intercept\']],
                               slope=subset(regressions,Tissue==tissue)[[\'Slope\']]) +
                               ggtitle(\'',tissue,'\')',sep='')))
  }
}
```
In vivo volume of distribution data are compared with the predictions, and errors are calculated.  Regressions and improvements are calculated and plotted.

```{r VolumeofDistributionEvaluation}
obach <- subset(Obach2008,CAS %in% get_cheminfo(model='schmitt'))
vd.table <- NULL

for(this.cas in obach[,'CAS']){
  parameters <- parameterize_schmitt(chem.cas=this.cas)
  init.parameters <- parameters
  if(!this.cas %in% ma.data.list){
    init.parameters$MA <- 10^(0.999831 - 0.016578*37 + 0.881721 * log10(parameters$Pow))
  }
  pcs <- predict_partitioning_schmitt(parameters=parameters,regression=F)
  init.pcs <- predict_partitioning_schmitt(parameters=init.parameters,regression=F)
  reg.pcs <- predict_partitioning_schmitt(parameters=parameters,regression=T)
  vdist <- calc_vdist(parameters=c(pcs,
    Funbound.plasma=parameters$Funbound.plasma),
    suppress.messages = TRUE)
  init.vdist <- calc_vdist(parameters=c(init.pcs,
    Funbound.plasma=parameters$unadjusted.Funbound.plasma),
    suppress.messages = TRUE)
  reg.vdist <- calc_vdist(parameters=c(reg.pcs,
    Funbound.plasma=parameters$Funbound.plasma),
    suppress.messages = TRUE)
  vd.table <- rbind(vd.table,cbind(as.data.frame(this.cas),as.data.frame(log10(init.vdist)),
                    as.data.frame(log10(vdist)),as.data.frame(log10(reg.vdist)),
                    as.data.frame(log10(subset(obach,CAS==this.cas)[['VDss (L/kg)']]))))
}
colnames(vd.table) <- c('CAS','init.vdist','adjusted.vdist','calibrated.vdist','Experimental')
init.error <- vd.table[,'Experimental'] - vd.table[,'init.vdist']
adjustment.error <- vd.table[,'Experimental'] - vd.table[,'adjusted.vdist']
calibration.error <- vd.table[,'Experimental'] - vd.table[,'calibrated.vdist']
adjustment.improvement <- abs(init.error) - abs(adjustment.error)
calibration.improvement <- abs(adjustment.error) - abs(calibration.error)
vd.table <- cbind(vd.table,adjustment.improvement,calibration.improvement,
                  init.error,adjustment.error,calibration.error)
fit <- lm(Experimental ~ calibrated.vdist,data=vd.table)
smry <- summary(fit)
calibrated.reg <- cbind(as.data.frame(fit$coefficients['(Intercept)']),
                        as.data.frame(fit$coefficients['calibrated.vdist']),
                        as.data.frame(smry$coefficients['calibrated.vdist','Pr(>|t|)']),
                        as.data.frame(smry$sigma),as.data.frame(smry$r.squared))
fit <- lm(Experimental ~ init.vdist,data=vd.table)
smry <- summary(fit)
init.reg <- cbind(as.data.frame(fit$coefficients['(Intercept)']),
                  as.data.frame(fit$coefficients['init.vdist']),
                  as.data.frame(smry$coefficients['init.vdist','Pr(>|t|)']),
                  as.data.frame(smry$sigma),as.data.frame(smry$r.squared))
fit <- lm(Experimental ~ adjusted.vdist,data=vd.table)
smry <- summary(fit)
adjustment.reg <- cbind(as.data.frame(fit$coefficients['(Intercept)']),
                       as.data.frame(fit$coefficients['adjusted.vdist']),
                       as.data.frame(smry$coefficients['adjusted.vdist','Pr(>|t|)']),
                       as.data.frame(smry$sigma),as.data.frame(smry$r.squared))
colnames(init.reg) <- colnames(adjustment.reg) <-
colnames(calibrated.reg) <- c('Intercept','Slope','P-value','Std Err','R-squared')

init.vd.plot <- ggplot(vd.table,aes(10^(init.vdist),10^(Experimental))) + geom_point() +
    geom_abline(intercept = init.reg[['Intercept']], slope = init.reg[["Slope"]]) +
    geom_abline(linetype = "dashed") + xlab("Predicted Volume of Distribution") +
    ylab("Measured Volume of Distribution") + theme(axis.text=element_text(size=16),
    axis.title=element_text(size=16),plot.title=element_text(size=18,hjust = 0.5)) +
    scale_x_log10(label=scientific_10,limits=c(10^(-1.5),10^(8.5))) + 
    scale_y_log10(label=scientific_10,limits=c(10^(-1.5),10^(8.5))) +
    ggtitle('(A)')
print(init.vd.plot)

adjustment.plot <- ggplot() +
    geom_point(data=vd.table[order(vd.table[,'adjustment.improvement'],decreasing=F),],
    aes(10^(adjusted.vdist),10^(Experimental),color=adjustment.improvement)) +
    geom_abline(intercept = adjustment.reg[['Intercept']],slope = adjustment.reg[["Slope"]]) +
    geom_abline(linetype = "dashed") + xlab("Predicted Volume of Distribution") +
    ylab("Measured Volume of Distribution") + theme(legend.position = c(.95, .95),
    legend.justification = c("right", "top"),legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),axis.text=element_text(size=16),
    axis.title=element_text(size=16),plot.title=element_text(size=18,hjust = 0.5)) +
    scale_x_log10(limits=c(10^(-1.5),10^(3))) + scale_y_log10(limits=c(10^(-1.5),10^(3))) +
    ggtitle('(B)') + scale_color_viridis(direction=-1,option='inferno')
print(adjustment.plot)

calibration.plot <- ggplot() +
    geom_point(data=vd.table[order(vd.table[,'calibration.improvement'],decreasing=F),],
    aes(10^(calibrated.vdist),10^(Experimental),color=calibration.improvement)) +
    geom_abline(intercept = calibrated.reg[['Intercept']],slope = calibrated.reg[["Slope"]]) +
    geom_abline(linetype = "dashed") + xlab("Predicted Volume of Distribution") +
    ylab("Measured Volume of Distribution") + theme(legend.position = c(.95, .95),
    legend.justification = c("right", "top"),legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),axis.text=element_text(size=16),
    axis.title=element_text(size=16),plot.title=element_text(size=18,hjust = 0.5)) +
    scale_x_log10(limits=c(10^(-1.5),10^(3))) + scale_y_log10(limits=c(10^(-1.5),10^(3))) +
    ggtitle('(C)') + scale_color_viridis(direction=-1,option='inferno')
print(calibration.plot)
```

Now we pull in the in vivo blood to plasma ratios, use these to calculate the inferred red blood cell to plasma ratios, and then make predictions for these values.
A censored regressions is performed, and predictions are plotted against errors.

```{r RBCtoplasma}
rb2p.data <- subset(chem.physical_and_invitro.data,!is.na(Human.Rblood2plasma))
measured.rb2p <- NULL
measured.krbc <- NULL
predicted.rb2p <- NULL
predicted.krbc <- NULL
cas <- NULL
charge <- NULL
fup <- NULL
logP <- NULL
pka_donor <- NULL
pka_accept <- NULL
for(this.cas in rb2p.data[rb2p.data[,'CAS'] %in% get_cheminfo(model='schmitt'),'CAS']){
  rb2p <- get_rblood2plasma(chem.cas=this.cas)
  krbc <- (rb2p + .44 - 1) / 0.44
  measured.rb2p <- c(measured.rb2p,rb2p)
  measured.krbc <- c(measured.krbc,krbc)
  parameters <- parameterize_schmitt(chem.cas=this.cas)
  pcs <- predict_partitioning_schmitt(parameters=parameters)
  predicted.krbc <- c(predicted.krbc,pcs[['Krbc2pu']] * parameters$Funbound.plasma)
  cas <- c(cas,this.cas)
  charge <- c(charge,calc_ionization(chem.cas=this.cas,pH=7.4)$fraction_charged)
  fup <- c(fup,parameters$unadjusted.Funbound.plasma)
  logP <-  c(logP,log10(parameters$Pow))
  pka_donor <- c(pka_donor,paste(parameters$pKa_Donor,collapse=','))
  pka_accept <- c(pka_accept,paste(parameters$pKa_Accept,collapse=','))
}
predicted.rb2p <-  1 - 0.44 + 0.44 * predicted.krbc
rb2p.table <- cbind(as.data.frame(cas),as.data.frame(predicted.rb2p),as.data.frame(measured.rb2p))
colnames(rb2p.table) <- c('cas','predicted.rb2p','measured.rb2p')
error <- log10(rb2p.table[,'measured.rb2p']) - log10(rb2p.table[,'predicted.rb2p'])
rb2p.table <- cbind(rb2p.table,error,charge,fup,logP)

error <- log10(measured.krbc) -  log10(predicted.krbc)
krbc.table <- cbind(as.data.frame(cas),as.data.frame(predicted.krbc),as.data.frame(measured.krbc),
                    as.data.frame(error),charge,fup,logP,pka_donor,pka_accept)

pdta <- data.frame(x = predicted.krbc,
                   y = measured.krbc)
pdta$y[pdta$y <= 0.1] <- 0.1
pdta$Censoring <- factor(c("Not Censored","Censored")[as.numeric(pdta$y <= 0.1) + 1])
y <- measured.krbc
x <- cbind(rep(1, length(y)),-1 * log10(predicted.krbc))
colnames(x) <- c("Intercept","Predicted")
cc <- as.numeric(y <= 0.1)
y[y < 0.1] <- 0.1
y <- -log10(y)

out <- em.cens(cc, x, y, dist="Normal")

censored.regression <- ggplot() +
    geom_point(data=pdta,aes(x=x,y=y, color=Censoring)) +
    scale_x_log10(limits=c(.0009,40)) + scale_y_log10(limits=c(.1,4),breaks=c(.1,.5,2.5)) +
    labs(y=expression(paste("Inferred ",K[p])),x=expression(paste("Predicted ",K[p]))) +
    geom_abline(intercept=0, slope=1, linetype='dashed') +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16),
    plot.title=element_text(size=18,hjust=0.5),legend.position = c(0.11, .8)) +
    geom_abline(slope=out$betas[2],intercept=-out$betas[1]) + ggtitle('(B)')

rb2p.plot <- ggplot(rb2p.table,aes(predicted.rb2p,measured.rb2p)) +
    geom_point()  + scale_x_log10(lim=c(.52,18)) + 
    scale_y_log10(lim=c(.52,2.5),breaks=c(0.5,1,2)) + geom_abline(linetype='dashed') +
    labs(y="Measured B:P Ratio ",
    x="Predicted B:P Ratio") +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16),
    plot.title=element_text(size=18,hjust=0.5)) + ggtitle('(A)')
print(rb2p.plot)
```

Lastly, we make the heatmap.
```{r MakeHeatmap}
heatmap.table <- NULL
for(this.cas in get_cheminfo(model='schmitt')){
    parms <- parameterize_schmitt(chem.cas=this.cas)
    pcs <- predict_partitioning_schmitt(parameters=parms)
    heatmap.table <- cbind(heatmap.table,log10(unlist(pcs)[1:11]*parms$Funbound.plasma))
}
rownames(heatmap.table) <-  c('Adipose','Bone','Brain','Gut','Heart',
                              'Kidney','Liver','Lung','Muscle','Skin','Spleen')
colnames(heatmap.table) <- rep("",dim(heatmap.table)[2])

pal <- function (n, h = c(260, -328), c = 80, l = c(30, 100), power = 1.5,
    fixup = TRUE, gamma = NULL, alpha = 1, ...)
{
    if (!is.null(gamma))
        warning("'gamma' is deprecated and has no effect")
    if (n < 1L)
        return(character(0L))
   h <- rep(h, length.out = 2L)
    c <- c[1L]
    l <- rep(l, length.out = 2L)
    power <- rep(power, length.out = 2L)
    rval <- seq(1, -1, length = n)
    rval <- hex(polarLUV(L = l[2L] - diff(l) * abs(rval)^power[2L],
        C = c * abs(rval)^power[1L], H = ifelse(rval > 0, h[1L],
            h[2L])), fixup = fixup, ...)
    if (!missing(alpha)) {
        alpha <- pmax(pmin(alpha, 1), 0)
        alpha <- format(as.hexmode(round(alpha * 255 + 1e-04)),
            width = 2L, upper.case = TRUE)
        rval <- paste(rval, alpha, sep = "")
    }
    return(rval)
}

hclust.ave <- function(x) hclust(x, method="ward.D2")
par(cex.main=1.5,cex.lab=2)
lhei <- c(1,4)
lwid <- c(1.25,2)
lmat <- rbind(c(2,3),c(4,1))
heatmap.2(heatmap.table,dendrogram='column',col=pal,trace="none",
          hclustfun=hclust.ave,key.xlab=expression(paste(log[10],K[p]," Value")),
          key.ylab=expression(paste("Number of ",K[p])),
          key.title="Partition Coefficients",xlab="Chemicals",cex.lab=2,
          lmat=lmat,margins=c(2,5),lwid=lwid,lhei=lhei)
```
