---
title: "Ring et al. (2017) Global sensitivity analysis"
author: "Caroline Ring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ring et al. (2017) Global sensitivity analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
```

This vignette includes the code to perform the global sensitivity analysis described in the HTTK-Pop paper.

```{r, eval = FALSE}
library(httk)
library(data.table)
library(parallel)
```

Basically, the idea of the Sobol method is to apportion variance: to estimate 
how much of the variance in a model's output is contributed by variance in each 
of the model parameters. The fraction of output variance attributable to each 
parameter is called the sensitivity index for that parameter.

The sensitivity indexes can be estimated using the following method. Generate 
two parameter matrices, independently of one another. Call them Matrix A and 
Matrix B. Evaluate the model using A and B, to get two vectors of output, fA 
and fB. Then, swap one column at a time between A and B. If you swap column i, 
the resulting matrices can be called ABi (for matrix A, with column i from 
matrix B) and BAi (for matrix B, with column i from matrix A). Evaluate the 
model for the column-swapped matrices to get another two vectors of output, 
fABi and fBAi. Then first-order and total sensitivity indices can be estimated 
from the correlation coefficients between output vectors (Glen & Isaacs 2012).

With HTTK-Pop, this means we need to generate two virtual populations with the 
same specifications. For each chemical, we then need to add `Funbound.plasma` 
and `Clint` columns to each population matrix, and convert it into a matrix of 
HTTK model parameters. That will give matrices A and B, for each chemical. Then, 
we need to swap columns, evaluate the model, and compute the sensitivity indices 
from the correlations.

Let's get started. First, let's generate the two virtual populations using 
HTTK-Pop. These are strictly physiological parameters -- no chemical dependence 
-- so we can just do this once and re-use the two virtual populations for each 
chemical.

We'll generate virtual populations for each of the ten ExpoCast subpopulations 
of interest, to see if and how global sensitivity changes by subpopulation. In 
fact, we've already generated one set of virtual populations in the "Generating 
subpopulations" vignette. Let's just re-use those as the first virtual 
population, and repeat the same procedure to generate a second virtual 
population, for each subpopulation.

We'll use the same seed as before (based on my name), but increment it by 1, so 
that the two virtual populations are not identical.
```{r, eval=FALSE}
seed.int <- TeachingDemos::char2seed('Caroline Ring', set=FALSE)+1
```
We set up the subpopulation specifications in the same way as before.
```{r, eval=FALSE}
nsamp<-25
ExpoCast.group<-list("Total",
                     "Age.6.11",
                     "Age.12.19",
                     "Age.20.65",
                     "Age.GT65",
                     "BMIgt30",
                     "BMIle30",
                     "Females",
                     "Males",
                     "ReproAgeFemale",
                     "Age.20.50.nonobese")

gendernum <- c(rep(list(NULL),7), 
               list(list(Male=0, Female=1000)), 
               list(list(Male=1000, Female=0)), 
               list(list(Male=0, Female=1000)), 
               list(NULL))

agelim<-c(list(c(0,79),
               c(6,11),
               c(12,19),
               c(20,65),
               c(66,79)),
          rep(list(c(0,79)),4),
          list(c(16,49)),
          list(c(20,50)))

bmi_category <- c(rep(list(c('Underweight', 
                             'Normal',
                             'Overweight',
                             'Obese')),
                      5),
                  list('Obese',
                       c('Underweight','Normal', 'Overweight')),
                  rep(list(c('Underweight', 
                             'Normal',
                             'Overweight',
                             'Obese')),
                      3),
                  list(c('Underweight', 'Normal', 'Overweight')))
```
Then we just evaluate another set of virtual populations, using the direct resampling method.
```{r, eval=FALSE}
tmpfun <- function(gendernum, agelim, bmi_category, ExpoCast_grp,
                   nsamp, method){
  result <- tryCatch({
    pops <- httk::httkpop_generate(
      method=method,
      nsamp=nsamp,
      gendernum = gendernum,
      agelim_years = agelim,
      weight_category = bmi_category)
    
    filepart <- switch(method,
                       'virtual individuals' = 'vi',
                       'direct resampling' = 'dr')
    saveRDS(object=pops,
            file=paste0('data/httkpop_',
                        filepart,
                        '_',
                        ExpoCast_grp,
                        '2.Rdata')) #Note we've added a 2 to the file name!
    return(0)
    }, error = function(err){
      print(paste('Error occurred:', err))
      return(1)
      })
  }


cluster <- parallel::makeCluster(2, #Change number of cores from 40 to 2
                                 outfile='subpopulations_parallel_out2.txt')

evalout <- parallel::clusterEvalQ(cl=cluster,
                                  {library(data.table)
                                   library(httk)})
parallel::clusterExport(cl = cluster,
                        varlist = 'tmpfun')
#Set seeds on all workers for reproducibility
parallel::clusterSetRNGStream(cluster, 
                              seed.int)
out_dr <- parallel::clusterMap(cl=cluster,
                               fun = tmpfun,
                               gendernum=gendernum,
                               agelim=agelim,
                               bmi_category=bmi_category,
                               ExpoCast_grp = ExpoCast.group,
                               MoreArgs = list(nsamp = nsamp,
                                               method = 'direct resampling')
                               )
parallel::stopCluster(cluster)
```
Great. Now, let's think of the procedure as looping first over the ten 
subpopulations of interest, and then over the chemicals in the HTTK data set. 
So first, let's think about what we need to do for each chemical, given a subpopulation.

We need to draw `Funbound.plasma` and `Clint` samples for the two virtual 
populations and convert into HTTK parameter matrices. That will give us matrix A 
and matrix B. Then, we need to evaluate yA and yB. Then, we need to loop over 
the HTTK model parameters -- the columns of A and B -- and swap one column at a 
time between A and B. For parameter i, we then need to evaluate yABi and yBAi. 
Then, we need to compute the correlations, and from them the sensitivity indices 
for parameter i. 

So we can write some pseudocode like this:

* for subpopulation subpop
* load previously-generated virtual population 1
* generate virtual population 2
* for chemical chem
* draw Funbound.plasma and Clint for virtual population 1
* draw Funbound.plasma and Clint for virtual population 2
* convert virtual population 1 into HTTK parameter matrix A
* convert virtual population 2 into HTTK parameter matrix B
* evaluate HTTK model for A and B, yielding yA and yB
* for HTTK parameter i{
* swap column i between A and B, forming ABi and BAi
* evaluate HTTK model for ABi and BAi, yielding yABi and yBAi
* compute correlations among yA, yB, yABi, yBAi
* compute sensitivity indexes for parameter i
* end loop over HTTK parameters
* end loop over chemicals
* end loop over subpopulations

# For each parameter
Let's start with the innermost loop: what we want to do for each parameter i.
```{r, eval=FALSE}
doforeachparam <- function(i,
                           indiv.A,
                           indiv.B,
                           this.chemcas,
                           model,
                           species='Human',
                           css.method,
                           f.A, #already evaluated model for A
                           f.B, #already evaluated model for B
                           g0, #normalized f.A
                           g0prime, #normalized f.B
                           p0){ #spurious correlation betweenf.A and f.B
  
  #make copies so as not to change these data.tables outside the function
  indiv.A <- copy(indiv.A)
  indiv.B <- copy(indiv.B)
  
  #First, let's swap column i.
  indiv.ABi <- copy(indiv.A)
  indiv.ABi[, (i):= indiv.B[, i, with=FALSE]]
  indiv.BAi <- copy(indiv.B)
  indiv.BAi[, (i):=indiv.A[, i, with=FALSE]]
  
  #Convert to HTTK parameters for matrix ABi and matrix BAi
  indiv.ABi.httk <- httk::convert_httk(indiv.model.bio = indiv.ABi,
                                          model = model,
                                          this.chem = this.chemcas)
  indiv.BAi.httk <- httk::convert_httk(indiv.model.bio = indiv.BAi,
                                          model = model,
                                          this.chem = this.chemcas)
  
  #If model is 3compartmentss, convert Funbound.plasma to Funbound.blood
  if (model=="3compartmentss"){
    #First, get the default parameters used for the Schmitt method of estimating
    #partition coefficients.
    pschmitt <- httk::parameterize_schmitt(chem.cas=this.chemcas,
                                           species='Human')
    convert_Fup <- function(DT.bio, DT.httk, pschmitt, this.chemcas){
      DT.bio <- copy(DT.bio)
      DT.httk <- copy(DT.httk)
    #next, replace the single default value for Funbound.plasma with the vector
    #of Funbound.plasma values from the virtual population data.table.
    pschmitt$Funbound.plasma<-DT.httk[, Funbound.plasma]
    #Now, predict the partitioning coefficients using Schmitt's method. The
    #result will be a list of numerical vectors, one vector for each
    #tissue-to-plasma partitioning coefficient, and one element of each vector
    #for each individual. The list element names specify which partition
    #coefficient it is, e.g. Kliver2plasma, Kgut2plasma, etc.
    PCs <- httk::predict_partitioning_schmitt(parameters=pschmitt,
                                              chem.cas=this.chemcas,
                                              species='Human')
    #Compute predicted Rblood2plasma (amount in blood/amount in plasma)
    Rb2p <- 1 - DT.bio$hematocrit + DT.bio$hematocrit * 
      PCs[["Krbc2pu"]] * 
      DT.httk$Funbound.plasma
    
  Funbound.blood <- DT.httk$Funbound.plasma/Rb2p
  return(list(Rb2p=Rb2p,
              Funbound.blood=Funbound.blood))
  }
  
  Fconvert.ABi <- convert_Fup(DT.bio=indiv.ABi,
                                                DT.httk=indiv.ABi.httk,
                                                pschmitt=pschmitt,
                                                this.chemcas=this.chemcas)
  
  Fconvert.BAi <- convert_Fup(DT.bio=indiv.BAi,
                                                DT.httk=indiv.BAi.httk,
                                                pschmitt=pschmitt,
                                                this.chemcas=this.chemcas)
  
  indiv.ABi.httk[, Funbound.plasma:=Fconvert.ABi$Funbound.blood]
  indiv.BAi.httk[, Funbound.plasma:=Fconvert.BAi$Funbound.blood]
  }
  #Next, let's evaluate the model.
  if (tolower(css.method)=='analytic') {
    f.ABi <- httk::calc_analytic_css(chem.cas=this.chemcas,
                                     parameters=indiv.ABi.httk,
                                     daily.dose=1, 
                                     output.units="uM",
                                     model=model,
                                     species=species,
                                     suppress.messages=TRUE,
                                     recalc.blood2plasma=TRUE)
    f.BAi <- httk::calc_analytic_css(chem.cas=this.chemcas,
                                     parameters=indiv.BAi.httk,
                                     daily.dose=1, 
                                     output.units="uM",
                                     model=model,
                                     species=species,
                                     suppress.messages=TRUE,
                                     recalc.blood2plasma=TRUE)
    
    if (model=="3compartmentss"){
      f.ABi <- f.ABi/Fconvert.ABi$Rb2p
      f.BAi <- f.BAi/Fconvert.BAi$Rb2p
    }
    }
  else if (tolower(css.method)=='full'){
    f.ABi <- apply(X=indiv.ABi.httk, 
                   MARGIN=1,
                   FUN=function(x) httk::calc_css(chem.cas=this.chemcas,
                                                  parameters=as.list(x),
                                                  daily.dose=1, 
                                                  output.units="uM",
                                                  model=model,
                                                  species=species,
                                                  suppress.messages=TRUE,
                                                  f.change=1e-5)[['avg']])
    f.BAi <- apply(X=indiv.BAi.httk, 
                   MARGIN=1,
                   FUN=function(x) httk::calc_css(chem.cas=this.chemcas,
                                                  parameters=as.list(x),
                                                  daily.dose=1, 
                                                  output.units="uM",
                                                  model=model,
                                                  species=species,
                                                  suppress.messages=TRUE,
                                                  f.change=1e-5)[['avg']])
    }
  
  #Finally, let's estimate the correlations and the sensitivity indexes. This code uses the same notation as Glen and Isaacs, so you can follow along from their paper.
  gj <- (f.ABi-mean(f.ABi))/sqrt(var(f.ABi) * (length(f.ABi)-1) / length(f.ABi))
  gjprime<-(f.BAi-mean(f.BAi))/sqrt(var(f.BAi) * (length(f.BAi)-1) / length(f.BAi))
  
  csj <- mean(g0prime*gj)
  csminusj <- mean(g0*gj)
  cdj <- sum(g0prime*gj + g0*gjprime)/(2*length(g0))
  cdminusj <- sum(g0*gj + g0prime*gjprime)/(2*length(g0))
  pj <- sum(g0*g0prime + gj*gjprime)/(2*length(g0))
  caj <- (cdj-pj*cdminusj)/(1-pj^2)
  caminusj <- (cdminusj - pj*cdj)/(1-pj^2)
  Si <- cdj - pj*(caminusj)/(1-caj*caminusj)
  Ti <- 1 - cdminusj + pj*caj/(1-caj*caminusj)
  return(list(Ti=Ti,
              Si=Si,
              p0=p0,
              pj=pj,
              csj=csj,
              csminusj=csminusj,
              cdj=cdj,
              cdminusj=cdminusj,
              caj=caj,
              caminusj=caminusj))
  }
```
 For each chemical
Great. Now take one step out, and define a function for what we need to do for each chemical.
```{r, eval=FALSE}
doforeachchem <- function(this.chemcas, #CAS for one chemical 
                          model, #HTTK model to use
                          species='Human', #Species for HTTK
                          nsamp, #Number of people in virtual population
                          css.method, #'analytic' or 'full'
                          sigma.factor=0.3, #coefficient of variation
                          indiv.model.bio1, #output of httkbio()
                          indiv.model.bio2, #output of httkbio()
                          ExpoCast.group, #subpopulation of interest
                          poormetab, #TRUE or FALSE
                          fup.censored.dist){ #TRUE or FALSE
                           
  #To avoid making changes outside the function
  indiv.model.bio1 <- copy(indiv.model.bio1)
  indiv.model.bio2 <- copy(indiv.model.bio2)
  
  #First draw Funbound.plasma and Clint
  indiv.A <- cbind(indiv.model.bio1,
                   httk::draw_fup_clint(this.chem=this.chemcas,
                                     nsamp=nrow(indiv.model.bio1),
                                     sigma.factor=sigma.factor,
                                     poormetab=poormetab,
                                     fup.censored.dist=fup.censored.dist))
  indiv.B <- cbind(indiv.model.bio2, 
                   httk::draw_fup_clint(this.chem=this.chemcas,
                                     nsamp=nrow(indiv.model.bio2),
                                     sigma.factor=sigma.factor,
                                     poormetab=poormetab,
                                     fup.censored.dist=fup.censored.dist))
  #Convert to HTTK parameters for matrix A and matrix B
  indiv.A.httk <- httk::convert_httk(indiv.model.bio = indiv.A,
                                        model = model,
                                        this.chem = this.chemcas)
  indiv.B.httk <- httk::convert_httk(indiv.model.bio = indiv.B,
                                        model = model,
                                        this.chem = this.chemcas)
  
  #If model is 3compartmentss, convert Funbound.plasma to Funbound.blood
  if (model=="3compartmentss"){
    #First, get the default parameters used for the Schmitt method of estimating
    #partition coefficients.
    pschmitt <- httk::parameterize_schmitt(chem.cas=this.chemcas,
                                           species='Human')
    convert_Fup <- function(DT.bio, DT.httk, pschmitt, this.chemcas){
      DT.bio <- copy(DT.bio)
      DT.httk <- copy(DT.httk)
    #next, replace the single default value for Funbound.plasma with the vector
    #of Funbound.plasma values from the virtual population data.table.
    pschmitt$Funbound.plasma<-DT.httk[, Funbound.plasma]
    #Now, predict the partitioning coefficients using Schmitt's method. The
    #result will be a list of numerical vectors, one vector for each
    #tissue-to-plasma partitioning coefficient, and one element of each vector
    #for each individual. The list element names specify which partition
    #coefficient it is, e.g. Kliver2plasma, Kgut2plasma, etc.
    PCs <- httk::predict_partitioning_schmitt(parameters=pschmitt,
                                              chem.cas=this.chemcas,
                                              species='Human')
    Rb2p <- 1 - DT.bio$hematocrit + DT.bio$hematocrit * 
      PCs[["Krbc2pu"]] * 
      DT.httk$Funbound.plasma
    
  Funbound.blood <- DT.httk$Funbound.plasma/Rb2p
  return(list(Rb2p=Rb2p,
              Funbound.blood=Funbound.blood))
  }
  
  Fconvert.A <- convert_Fup(DT.bio=indiv.A,
                                                DT.httk=indiv.A.httk,
                                                pschmitt=pschmitt,
                                                this.chemcas=this.chemcas)
  
  Fconvert.B <- convert_Fup(DT.bio=indiv.B,
                                                DT.httk=indiv.B.httk,
                                                pschmitt=pschmitt,
                                                this.chemcas=this.chemcas)
  
  indiv.A.httk[, Funbound.plasma:=Fconvert.A$Funbound.blood]
  indiv.B.httk[, Funbound.plasma:=Fconvert.B$Funbound.blood]
  }
  #Evaluate model
  if (tolower(css.method)=='analytic') {
    f.A <- httk::calc_analytic_css(chem.cas=this.chemcas,
                                   parameters=indiv.A.httk,
                                   daily.dose=1, 
                                   output.units="uM",
                                   model=model,
                                   species=species,
                                   suppress.messages=TRUE,
                                   recalc.blood2plasma=TRUE)
    f.B <- httk::calc_analytic_css(chem.cas=this.chemcas,
                                   parameters=indiv.B.httk,
                                   daily.dose=1, 
                                   output.units="uM",
                                   model=model,
                                   species=species,
                                   suppress.messages=TRUE,
                                   recalc.blood2plasma=TRUE)
    
    if (model=="3compartmentss"){
      f.A <- f.A/Fconvert.A$Rb2p
      f.B <- f.B/Fconvert.B$Rb2p
    }
    }
  else if (tolower(css.method)=='full'){
    f.A <- apply(X=indiv.A.httk, 
                 MARGIN=1,
                 FUN=function(x) httk::calc_css(chem.cas=this.chemcas,
                                                parameters=as.list(x),
                                                daily.dose=1, 
                                                output.units="uM",
                                                model=model,
                                                species=species,
                                                suppress.messages=TRUE,
                                                f.change=1e-5)[['avg']])
    f.B <- apply(X=indiv.B.httk, 
                 MARGIN=1,
                 FUN=function(x) httk::calc_css(chem.cas=this.chemcas,
                                                parameters=as.list(x),
                                                daily.dose=1, 
                                                output.units="uM",
                                                model=model,
                                                species=species,
                                                suppress.messages=TRUE,
                                                f.change=1e-5)[['avg']])
    }
  #Compute normalized data (number of standard deviations from the mean)
  g0<-(f.A-mean(f.A))/sqrt(var(f.A) * (length(f.A)-1) / length(f.A))
  g0prime <- (f.B-mean(f.B))/sqrt(var(f.B) * (length(f.B)-1) / length(f.B))
  
  #Compute spurious correlation term between f.A and f.B
  p0 <- mean(g0*g0prime)
  
  #Next, define the list of HTTK parameters that constitute the physiological parameters for this model.
  #Strictly speaking this should be all of the columns in indiv.model.bio1, except for liver.density, which is a fixed value.
  phys.par <- names(indiv.model.bio1)[names(indiv.model.bio1)!=
                                        'liver.density']
  #Add Funbound.plasma
  parlist <- list('Funbound.plasma',
                  'Clint',
                  phys.par)
  #Note: this is a list with three members! 
  #Sensitivity is calculated to the *group* of physiological parameters *as a whole*, not to each one individually.
  
  #If Clint is fixed at zero,
  #don't test sensitivity to it
  if (all(indiv.A[, Clint]==0)){
    parlist <- parlist[parlist!='Clint']
    }
  
  #Now loop over parameters (and groups thereof)
  eff.list <- lapply(parlist,
                     doforeachparam,
                     indiv.A=indiv.A,
                     indiv.B=indiv.B,
                     this.chemcas=this.chemcas,
                     model=model,
                     species=species,
                     css.method=css.method,
                     f.A=f.A,
                     f.B=f.B,
                     g0=g0,
                     g0prime=g0prime,
                     p0=p0)
  
  parnames <- parlist
  #Name the sensitivity indexes by their parameters
  #For the group of physiological parameters, just name it "phys.par"
  parnames[sapply(parnames, function(x) length(x)>1)] <- 'phys.par'
  names(eff.list) <- parnames
  
  #If Clint was zero, then sensitivity to it wasn't calculated.
  #Set all the sensitivity quantities to NA to indicate this.
  if (all(indiv.A[, Clint]==0)){
    eff.list <- c(eff.list,
                  list(Clint=list(Ti=NA,
                                  Si=NA,
                                  p0=NA,
                                  pj=NA,
                                  csj=NA,
                                  csminusj=NA,
                                  cdj=NA,
                                  cdminusj=NA,
                                  caj=NA,
                                  caminusj=NA)))
    }
  #Construct a data.table to return
  eff.dt <- as.data.table(do.call(rbind.data.frame,
                                  eff.list),
                          keep.rownames=TRUE)
  eff.dt[, chemcas:=this.chemcas]
  return(eff.dt)
  }
```
# For each subpopulation
Now take one more step out and define a function to do for each subpopulation.
```{r, eval=FALSE}
doforeachsubpop <- function(ecg, #Short for "ExpoCast group"
                            nsamp,
                            chemlist, #list of chemicals to loop over
                            model, #HTTK model to use
                            species = 'Human', #HTTK species to use 
                            sigma.factor = 0.3, #Coefficient of variation
                            css.method = 'analytic', #or 'full'
                            poormetab, #TRUE or FALSE
                            fup.censored.dist, #TRUE or FALSE
                            cluster){ #The parallel cluster to use (we'll parallelize the loop over chemicals)
  #First read in the virtual population data.
  indiv1.dt<-readRDS(paste0('data/httkpop_',
                            'dr',
                            '_',
                            ecg,
                            '.Rdata'))
  indiv2.dt<-readRDS(paste0('data/httkpop_',
                            'dr',
                            '_',
                            ecg,
                            '2.Rdata'))
  
  #Convert the biological parameters first
  indiv.model.bio1 <- httk::httkpop_bio(indiv_dt=indiv1.dt)
  indiv.model.bio2 <- httk::httkpop_bio(indiv_dt=indiv2.dt)
  
  #Now loop over the chemicals
  eff.allchems <- rbindlist(
    parLapplyLB(cl=cluster,
                chemlist, 
                doforeachchem,
                model=model, 
                species=species, 
                sigma.factor=sigma.factor,
                css.method=css.method,
                indiv.model.bio1=indiv.model.bio1,
                indiv.model.bio2=indiv.model.bio2,
                ExpoCast.group=ecg,
                poormetab=poormetab,
                fup.censored.dist=fup.censored.dist,
                nsamp=nsamp))
  #Add a column to the data.table denoting the subpopulation
  eff.allchems[,ExpoCast.group:=ecg]
  #And return the output
  return(eff.allchems)
  }
```
# Loop over subpopulations
Excellent! All that remains is to write the loop over the subpopulations that 
will call all of these functions in turn. We'll also add loops over the values 
of `poormetab` and `Funbound.plasma`, to see if and how global sensitivity 
changes under different conditions on the distributions of those variables.

**WARNING: this code may take a LONG time to run!** 4 model evaluations per 
individual times 1000 individuals times 3 parameters times about 500 chemicals 
times 10 subpopulations = 6e7 model evaluations, plus overhead. While the model 
evaluations are vectorized for the analytic steady-state models -- so 1000 model 
evaluations take very little time -- this still can add up. On my machine, it 
took about 7 minutes for each combination of CLint and Funbound conditions, and 
about half an hour total -- and that is parallelized with 10 cores.
```{r, eval=FALSE}
model <- "3compartmentss"
species <- "Human"
sigma.factor <- 0.3
css.method <- "analytic"
chemlist <- httk::get_cheminfo(info="CAS",
                               exclude.fup.zero=FALSE)

cluster <- parallel::makeCluster(2, # Changed number of corse from 40 to 2 
                                 outfile='globalsens_parallel_out.txt')

evalout <- parallel::clusterEvalQ(cl=cluster,
                                  {library(data.table)
                                   library(httk)})
parallel::clusterExport(cl = cluster,
                        varlist = c('doforeachparam',
                                    'doforeachchem',
                                    'doforeachsubpop'))
#Set seeds on all workers for reproducibility
parallel::clusterSetRNGStream(cluster, 
                              TeachingDemos::char2seed("Caroline Ring"))

system.time({for (poormetab in c(TRUE, FALSE)){
  for (fup.censored.dist in c(TRUE, FALSE)) {
    eff.all<-rbindlist(lapply(ExpoCast.group,
                              doforeachsubpop,
                              nsamp=nsamp,
                              chemlist=chemlist,
                              model=model, 
                              species=species,
                              sigma.factor=sigma.factor,
                              css.method=css.method,
                              poormetab=poormetab,
                              fup.censored.dist=fup.censored.dist,
                              cluster=cluster))
    saveRDS(eff.all, 
            paste0('data/',
                   'sens_glenisaacs_nhanes_', 
                   'allchems_',
                   'allgroups_',
                   model,
                   '_',
                   css.method,
                   '_fup_censor_',
                   fup.censored.dist,
                   '_poormetab_',
                   poormetab,
                   "_FuptoFub",
                   '.Rdata'))
    }
  }
  })
parallel::stopCluster(cluster)
```

Finally, we want to do sensitivity analysis for independent Monte Carlo, because 
that lets us estimate sensitivity to each physiological parameter independently.

The first thing we need to do is generate two separate populations using 
independent Monte Carlo. To save space, we'll set this up as a function that 
takes the seed as an argument.

```{r, eval=FALSE}
indep_gen <- function(nsamp=1000, sigma.factor=0.3){
  
  COmean <- physiology.data[physiology.data$Parameter=='Cardiac Output',
                               'Human']
  indep.bio <- data.table(Qcardiacc=truncnorm::rtruncnorm(n=nsamp,
                                                          mean=COmean,
                                                          sd=sigma.factor*COmean,
                                                          a=0)/1000*60)
  indep.bio[, BW:=truncnorm::rtruncnorm(n=nsamp,
                                        mean=physiology.data[physiology.data$Parameter=='Average BW',
                                                                'Human'],
                                        sd=sigma.factor*physiology.data[physiology.data$Parameter=='Average BW',
                                                                           'Human'],
                                        a=0)]
  indep.bio[, plasma.vol:=truncnorm::rtruncnorm(n=nsamp,
                                                mean=physiology.data[physiology.data$Parameter=='Plasma Volume',
                                                                        'Human'],
                                                sd=sigma.factor*physiology.data[physiology.data$Parameter=='Plasma Volume',
                                                                                   'Human'],
                                                a=0)/1000] #convert mL/kg to L/kg
  indep.bio[, hematocrit:=truncnorm::rtruncnorm(n=nsamp,
                                                mean=physiology.data[physiology.data$Parameter=='Hematocrit',
                                                                        'Human'],
                                                sd=sigma.factor*physiology.data[physiology.data$Parameter=='Hematocrit',
                                                                                   'Human'],
                                                a=0,
                                                b=1)]
  indep.bio[, million.cells.per.gliver:=truncnorm::rtruncnorm(n=nsamp,
                                                              mean=110,
                                                              sd=sigma.factor*110,
                                                              a=0)]
  
  all.tissues <- tissue.data$Tissue[tissue.data$Tissue!='red blood cells']
  for (tissue in all.tissues){
    vol.mean <- tissue.data[tissue.data$Tissue==tissue,
                            'Human Vol (L/kg)']
    flow.mean <- tissue.data[tissue.data$Tissue==tissue,
                             'Human Flow (mL/min/kg^(3/4))']/
      1000*60
    if (tissue=='liver'){ #subtract gut flow from portal vein flow
      #to get arterial flow
      flow.mean <- (tissue.data[tissue.data$Tissue=='liver',
                                'Human Flow (mL/min/kg^(3/4))'] -
                      tissue.data[tissue.data$Tissue=='gut',
                                  'Human Flow (mL/min/kg^(3/4))'])/
        1000*60
      }
    indep.bio[, paste0('V',
                       tissue,
                       'c'):=truncnorm::rtruncnorm(n=nsamp,
                                                   mean=vol.mean,
                                                   sd=sigma.factor*vol.mean,
                                                   a=0)]
    indep.bio[, paste0('Q',
                       tissue,
                       'f'):=truncnorm::rtruncnorm(n=nsamp,
                                                   mean=flow.mean,
                                                   sd=sigma.factor*flow.mean,
                                                   a=0)/Qcardiacc]
    }
  
  
  
  indep.bio[, Qtotal.liverc:=(Qliverf+Qgutf)*Qcardiacc]
  indep.bio[, liver.density:=1.05]
  gfr.mean<-physiology.data[physiology.data$Parameter=='GFR',
                               'Human']*60/1000 #convert from ml/min/kg^(3/4) to L/hr/kg(3/4)
  indep.bio[, Qgfrc:=truncnorm::rtruncnorm(n=nsamp,
                                           mean=gfr.mean,
                                           sd=sigma.factor*gfr.mean,
                                           a=0)]
  indep.bio[, 
            Vartc:= plasma.vol/
              (1-hematocrit)/2] #L/kgBW
  indep.bio[, 
            Vvenc:= plasma.vol/
              (1-hematocrit)/2] #L/kgBW
  return(indep.bio)
  }
```

And then we call it twice to get two different populations.

```{r, eval=FALSE}
TeachingDemos::char2seed("Caroline Ring")
indep.bio1 <- indep_gen()
indep.bio2 <- indep_gen()
```

The innermost function (over parameters) will be the same, but we need to 
modify the function of things to do for each chemical, because we need to do 
sensitivity analysis for the physiological parameters independently, not as a 
group.

```{r, eval=FALSE}
 Commented out because of problem with Pandoc for Windows:
doforeachchem_indep <- function(this.chemcas, #CAS for one chemical 
                                model, #HTTK model to use
                                species='Human', #Species for HTTK
                                nsamp, #Number of people in virtual population
                                css.method, #'analytic' or 'full'
                                sigma.factor=0.3, #coefficient of variation
                                indep.model.bio1, #output of indep_gen()
                                indep.model.bio2, #output of indep_gen()
                                poormetab, #TRUE or FALSE
                                fup.censored.dist){  #TRUE or FALSE
  #To avoid making changes outside the function
  indiv.model.bio1 <- copy(indep.model.bio1)
  indiv.model.bio2 <- copy(indep.model.bio2)
  
  #First draw Funbound.plasma and Clint
  indiv.A <- httk::draw_fup_clint(this.chem=this.chemcas,
                                     indiv.bio.tmp=indiv.model.bio1,
                                     sigma.factor=sigma.factor,
                                     poormetab=poormetab,
                                     fup.censored.dist=fup.censored.dist)
  indiv.B <- httk::draw_fup_clint(this.chem=this.chemcas,
                                     indiv.bio.tmp=indiv.model.bio2,
                                     sigma.factor=sigma.factor,
                                     poormetab=poormetab,
                                     fup.censored.dist=fup.censored.dist)
  #Convert to HTTK parameters for matrix A and matrix B
  indiv.A.httk <- httk::convert_httk(indiv.model.bio = indiv.A,
                                        model = model,
                                        this.chem = this.chemcas)
  indiv.B.httk <- httk::convert_httk(indiv.model.bio = indiv.B,
                                        model = model,
                                        this.chem = this.chemcas)
  
  #If model is 3compartmentss, convert Funbound.plasma to Funbound.blood
  if (model=="3compartmentss"){
    #First, get the default parameters used for the Schmitt method of estimating
    #partition coefficients.
    pschmitt <- httk::parameterize_schmitt(chem.cas=this.chemcas,
                                           species='Human')
    convert_Fup <- function(DT.bio, DT.httk, pschmitt, this.chemcas){
      DT.bio <- copy(DT.bio)
      DT.httk <- copy(DT.httk)
    #next, replace the single default value for Funbound.plasma with the vector
    #of Funbound.plasma values from the virtual population data.table.
    pschmitt$Funbound.plasma<-DT.httk[, Funbound.plasma]
    #Now, predict the partitioning coefficients using Schmitt's method. The
    #result will be a list of numerical vectors, one vector for each
    #tissue-to-plasma partitioning coefficient, and one element of each vector
    #for each individual. The list element names specify which partition
    #coefficient it is, e.g. Kliver2plasma, Kgut2plasma, etc.
    PCs <- httk::predict_partitioning_schmitt(parameters=pschmitt,
                                              chem.cas=this.chemcas,
                                              species='Human')
    Rb2p <- 1 - DT.bio$hematocrit + DT.bio$hematocrit * 
      PCs[["Krbc2pu"]] * 
      DT.httk$Funbound.plasma
    
  Funbound.blood <- DT.httk$Funbound.plasma/Rb2p
  return(list(Rb2p=Rb2p,
              Funbound.blood=Funbound.blood))
  }
  
  Fconvert.A <- convert_Fup(DT.bio=indiv.A,
                                                DT.httk=indiv.A.httk,
                                                pschmitt=pschmitt,
                                                this.chemcas=this.chemcas)
  
  Fconvert.B <- convert_Fup(DT.bio=indiv.B,
                                                DT.httk=indiv.B.httk,
                                                pschmitt=pschmitt,
                                                this.chemcas=this.chemcas)
  
  indiv.A.httk[, Funbound.plasma:=Fconvert.A$Funbound.blood]
  indiv.B.httk[, Funbound.plasma:=Fconvert.B$Funbound.blood]
  }
  #Evaluate model
  if (tolower(css.method)=='analytic') {
    f.A <- httk::calc_analytic_css(chem.cas=this.chemcas,
                                   parameters=indiv.A.httk,
                                   daily.dose=1, 
                                   output.units="uM",
                                   model=model,
                                   species=species,
                                   suppress.messages=TRUE,
                                   recalc.blood2plasma=TRUE)
    f.B <- httk::calc_analytic_css(chem.cas=this.chemcas,
                                   parameters=indiv.B.httk,
                                   daily.dose=1, 
                                   output.units="uM",
                                   model=model,
                                   species=species,
                                   suppress.messages=TRUE,
                                   recalc.blood2plasma=TRUE)
    if (model=="3compartmentss"){
      f.A <- f.A/Fconvert.A$Rb2p
      f.B <- f.B/Fconvert.B$Rb2p
    }
    }
  else if (tolower(css.method)=='full'){
    f.A <- apply(X=indiv.A.httk, 
                 MARGIN=1,
                 FUN=function(x) httk::calc_css(chem.cas=this.chemcas,
                                                parameters=as.list(x),
                                                daily.dose=1, 
                                                output.units="uM",
                                                model=model,
                                                species=species,
                                                suppress.messages=TRUE,
                                                f.change=1e-5)[['avg']])
    f.B <- apply(X=indiv.B.httk, 
                 MARGIN=1,
                 FUN=function(x) httk::calc_css(chem.cas=this.chemcas,
                                                parameters=as.list(x),
                                                daily.dose=1, 
                                                output.units="uM",
                                                model=model,
                                                species=species,
                                                suppress.messages=TRUE,
                                                f.change=1e-5)[['avg']])
    }
  #Compute normalized data (number of standard deviations from the mean)
  g0<-(f.A-mean(f.A))/sqrt(var(f.A) * (length(f.A)-1) / length(f.A))
  g0prime <- (f.B-mean(f.B))/sqrt(var(f.B) * (length(f.B)-1) / length(f.B))
  
  #Compute spurious correlation term between f.A and f.B
  p0 <- mean(g0*g0prime)
  
  #Next, define the list of HTTK parameters that constitute the physiological parameters for this model.
  #Strictly speaking this should be all of the columns in indiv.model.bio1, except for liver.density, which is a fixed value.
#   #From previous runs, these are the only parameters that the models are sensitive to
  phys.par <- switch(model,
                     'pbtk'=c('BW',
                              'hematocrit',
                              'million.cells.per.gliver',
                              'Qcardiacc',
                              'Qgfrc',
                              'Qgutf',
                              'Qkidneyf',
                              'Qliverf',
                              'Vliverc'),
                     '3compartment'=c('BW',
                                      'hematocrit',
                                      'million.cells.per.gliver',
                                      'Qgfrc',
                                      'Qgutf',
                                      'Qliverf',
                                      'Vliverc'),
                     '3compartmentss'=c('BW',
                                        'million.cells.per.gliver',
                                        'Qgfrc',
                                        'Qtotal.liverc',
                                        'Vliverc'),
                     '1compartment'=c('BW',
                                      'million.cells.per.gliver',
                                      'Qgfrc',
                                      'Qtotal.liverc',
                                      'Vliverc'))
phys.par.all <- names(indiv.model.bio1)[names(indiv.model.bio1)!=
                                            'liver.density']
  parlist <- list('Clint',
                  'Funbound.plasma',
                  phys.par.all)
  
  parlist <- c(parlist, as.list(phys.par))
  #Note: Now sensitivity is being calculated to each of the physiological parameter independently, as well as grouped.
  
  #If Clint is fixed at zero,
  #don't test sensitivity to it
  if (all(indiv.A[, Clint]==0)){
    parlist <- parlist[parlist!='Clint']
    }
  
  #Now loop over parameters (and groups thereof)
  eff.list <- lapply(parlist,
                     doforeachparam,
                     indiv.A=indiv.A,
                     indiv.B=indiv.B,
                     this.chemcas=this.chemcas,
                     model=model,
                     species=species,
                     css.method=css.method,
                     f.A=f.A,
                     f.B=f.B,
                     g0=g0,
                     g0prime=g0prime,
                     p0=p0)
  
  parnames <- parlist
  #Name the sensitivity indexes by their parameters
  #For the group of physiological parameters, just name it "phys.par"
  parnames[sapply(parnames, function(x) length(x)>1)] <- 'phys.par'
  names(eff.list) <- parnames
  
  #If Clint was zero, then sensitivity to it wasn't calculated.
  #Set all the sensitivity quantities to NA to indicate this.
  if (all(indiv.A[, Clint]==0)){
    eff.list <- c(eff.list,
                  list(Clint=list(Ti=NA,
                                  Si=NA,
                                  p0=NA,
                                  pj=NA,
                                  csj=NA,
                                  csminusj=NA,
                                  cdj=NA,
                                  cdminusj=NA,
                                  caj=NA,
                                  caminusj=NA)))
    }
  #Construct a data.table to return
  eff.dt <- as.data.table(do.call(rbind.data.frame,
                                  eff.list),
                          keep.rownames=TRUE)
  eff.dt[, chemcas:=this.chemcas]
  return(eff.dt)
  }
```

And finally, loop over the chemicals to actually do the sensitivity analysis.

```{r, eval=FALSE}
model <- "3compartmentss"
species <- "Human"
sigma.factor <- 0.3
css.method <- "analytic"
chemlist <- httk::get_cheminfo(info="CAS",
                               exclude.fup.zero=FALSE)

cluster <- parallel::makeCluster(2, # Was 40, reduced for simpler processors 
                                 outfile='globalsens_indep_parallel_out.txt')

evalout <- parallel::clusterEvalQ(cl=cluster,
                                  {library(data.table)
                                   library(httk)})
parallel::clusterExport(cl = cluster,
                        varlist = c('doforeachparam'))
Set seeds on all workers for reproducibility
parallel::clusterSetRNGStream(cluster, 
                              TeachingDemos::char2seed("Caroline Ring"))
for (poormetab in c(TRUE, FALSE)){
  for (fup.censored.dist in c(TRUE, FALSE)) {
    Now loop over the chemicals
    eff.allchems <- rbindlist(
      parLapplyLB(cl=cluster,
                  chemlist, 
                  doforeachchem_indep,
                  model=model, 
                  species=species, 
                  sigma.factor=sigma.factor,
                  css.method=css.method,
                  indep.model.bio1=indep.bio1,
                  indep.model.bio2=indep.bio2,
                  poormetab=poormetab,
                  fup.censored.dist=fup.censored.dist,
                  nsamp=10))  should be 1000, reduced for speed
    saveRDS(eff.allchems, 
            paste0('data/',
                   'sens_glenisaacs_nhanes_', 
                   'allchems_',
                   'indepMC_',
                   model,
                   '_',
                   css.method,
                   '_fup_censor_',
                   fup.censored.dist,
                   '_poormetab_',
                   poormetab,
                   "_FuptoFub",
                   '.Rdata'))
    }
  }
parallel::stopCluster(cluster)
```