---
title: "Predictions for fup using models presented in Dawson et al. 2021"

author: "Tabatabaei Sadeghi S, Guider C, Davidson-Fritz S, Dinallo R, Li L, Strock C , Wambaugh JF"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{h) Dawson's Model Evaluation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
*Please send questions to tabatabaeisadeghi.sahar@epa.gov*
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This R markdown script will allow a user to predict Fup values for chemicals based on the QSAR Models presented in Dawson et al. 2021: Designing QSARs for parameters of high throughput toxicokinetic models using open-source descriptors. Environmental Science and Technology. 

Data to accomplish this is contained in Fup folders, which each contain the random forest objects, and the means and standard deviations of the training set used to create the models. This script will read in novel chemicals for which predictors have been computed, standardize the parameters of the chemical according to the training set(s), and predict Fup for those chemicals. 

To fit predictions from the Fup model, you will need OPERA model predictions and PaDEL descriptions for your chemicals of interest. 

### OPERA predictors include:

"LogP"   "LogWS"  "LogBCF" "LogKM"  "MP"     "LogVP"  "LogD55" "LogD74" "LogOH" 

### PaDEL descriptors include:

 "CrippenLogP"       "nBase"             "XLogP"             "AATS1v"            "AATS0i"           
 "ETA_EtaP_F"        "ZMIC2"             "maxHBa"            "minsssN"           "AATS1i"           
 "ETA_dEpsilon_B"    "GATS1i"            "ZMIC1"             "SsssN"             "nAromBond"        
 "AATSC0p"           "minHBd"            "LipoaffinityIndex" "AATS3v"            "AATS2i"           
 "C2SP2"             "AATS4v"            "Kier2"             "AATS6m"            "AATS6v"           
 "minssCH2"          "AATS5v"            "AATS5i"            "ATSC2c"            "ATSC6c"           
 "ATSC3m"            "ATSC3e"            "AATSC0c"           "AATSC0v"           "MATS1c"           
 "MATS8c"            "GATS5c"            "GATS1e"            "GATS5e"            "BCUTc.1l"         
 "BCUTp.1l"          "BCUTp.1h"          "nBondsM"           "SpMin2_Bhm"        "SpMax2_Bhp"       
 "minHaaCH"          "minaaCH"           "IC1"               "MLFER_BH"          "piPC8"            
 "R_TpiPCTPC"     

These sets are included in the file "CombinedDescriptors_DED050621.RData", included in this folder. 

Several but not all needed OPERA predictors are available from the EPA CompTox Dashboard. However, PaDEL descriptors are not available at all from the dashboard.To calculate these descriptors for your data, we recommend downloading the OPERA(Open QSAR app) application from https://github.com/NIEHS/OPERA

The OPERA application is an open-source, stand-alone application that will produce both Opera predictions and PaDEL descriptions; Opera models are based entirely upon PaDEL descriptors and fingerprints. You will need either a chemical description ID(i.e., DTXSID,CASRN) or structural description(SMILES,MOL) for each chemical. Make sure to select all of the opera models you want, and select the "keep full descriptor files" box. This will generate and keep all calculations for all PaDEL descriptors, as well as your desired Opera predictors. Keep in mind that they may range from minutes to hours, depending on the number chemicals for which predictions are desired(10's to 1000's).   


If you want to use SMILES codes to run through the OPERA function, you can download them from the CompTox Dashboard. Then, you can write out the text file to be used in the OPERA application. Whatever data source you used, make sure there are no column headers or row number in your data file. For more information, read the documentation for data input into the OPERA application. 

By checking the "keep full descriptor files" box, OPERA will produce a lot of files you don't need. You only need the Opera predictions (with default suffix "_OPERA2.6Pred") and PaDEL descriptors (default suffix "_PadelDesc").  


# Read in data and prepare for prediction
Once the descriptors are calculated, load the files here, and prepare them to be used to make predictions. You have to make sure all of the descriptor names in your data match the descriptors used in the RF model objects. If you used SMILES codes, you also have make sure that the molecule ID output from Opera aligns with the DTXSID or CASRN of the chemicals in your set.  

In the example below, we use the Tox21 dataset used to produce predictions presented in the paper. We downloaded the smiles codes from the Comptox Dashboard ("Tox21_FromCompToxDashboard.xlsx") through the OPERA application. Then we loaded the OPERA descriptors and PaDEL descriptors produced from the OPERA application, and a combined list of OPERA and PaDEL descriptors used in both the Clint and fup models. We then create a combined set of Opera and Padel descriptors based on their DTXSID's. Note that if you only load SMILES into the OPERA application without a CAS or DTXSID, you have to align the molecule ID with the SMILES at this step. This is important in the case that either PaDEL or Opera predictors can't be calculated for a given chemical, and you end up with a missing chemical in the Opera prediction file.  


## LOAD PACKAGES
```{r}
packages=c("openxlsx", "randomForest", "stringr","dplyr")
sapply(packages, require, character.only=TRUE)
```


```{r}
#setting wd to the folder where the file is

setwd("C:/Users/stabatab/OneDrive - Environmental Protection Agency (EPA)/Profile/Desktop/ExpoCast_Projec/Project_Pred_Caco2_Clint_Fup/invitroTKstats/CyprotexTO1/out_fup")

level4=read.xlsx("180_Cyprotex-fup-RED-Level4.xlsx")

Operaversion = 2.9
Datalabel = "NewChemicals180_March5_2025"
load("CombinedDescriptors_DED050621.RData")#descriptors in both models
Descriptors$Opera=paste0(Descriptors$Opera, "_pred")

#adding the new opera predictions
Opera=read.csv("210Fup-DTXSID_SMILES-smi_OPERA2.8Pred.csv")
Padel=read.csv("padel-fup.csv")
Padel=unique(Padel[which(Padel$Name%in%Opera$MoleculeID),])
Padel=Padel[1:209,]

Padel=Padel[Padel$Name !="DTXSID5022306",]



Padel=Padel[Padel$Name %in% level4$DTXSID,]
Opera=Opera[Opera$MoleculeID %in%level4$DTXSID,]

#adding new padel predictors

Opera=Opera[,c(1,which(names(Opera)%in%Descriptors$Opera))]
Padel=Padel[,c(1,which(names(Padel)%in%Descriptors$Padel))]


#Merge with truncated opera and padel lists
chemlist=merge(Opera, Padel, by.x="MoleculeID", by.y="Name", all.x=TRUE)
# chemlist=chemlist[,colSums(is.na(chemlist))<nrow]


```

## Fraction unbound by plasma(fup)
Read in the fup model object and training data mean and std. 
Normalize your data using the training data. Then, use the predict function to predict fup values. After this, determine whether fup predictions
are in or out of the AD, using the the method from Roy et al. 2015. 

```{r}
fupchemlist=chemlist 
load("fup_RFmodel_30desc_select_by_sqrt_transform.RData")
fupmodvars=row.names(RF.fit$importance)
msd=read.csv("fup_mean_std_training_descriptors_DED052021.csv")

#substitute the correct names into the chemlist 
#Here, adapt descriptor names to the format expected in the fup random forest file. 
Padeldesc=which(names(fupchemlist)%in%str_remove(fupmodvars, 'padel_'))
Operadesc=which(str_remove(names(fupchemlist), "_pred")%in%str_remove(fupmodvars, 'Opera_'))
names(fupchemlist)=str_remove(names(fupchemlist), '_pred')

for(i in Padeldesc){names(fupchemlist)[i]=paste0("padel_",names(fupchemlist)[i])}
for(i in Operadesc){names(fupchemlist)[i]=paste0("Opera_",names(fupchemlist)[i])}

#Subset to only Fup model 
chemlistfupmodvars=which(names(fupchemlist)%in%fupmodvars)
IDcolstart=1
IDcolend=1
fupchemlist=fupchemlist[,c(IDcolstart:IDcolend,chemlistfupmodvars)]
chemlistfupsc=fupchemlist

# # Replace NA's with 0:
# 
# chemlistfupsc=as.data.frame(chemlistfupsc)
# chemlistfupsc[]=lapply(chemlistfupsc, function(x) ifelse(is.na(x)|is.nan(x), 0,x))


```
# Replace NA's with mean of training set:
```{r}

for (this.col in setdiff(colnames(msd),"X"))
{
  if (is.numeric(msd[1,this.col]))
  {
    chemlistfupsc[is.na(chemlistfupsc[,this.col]),this.col] <- msd[1,this.col]
  }
}


```
#Standardize by training set mean and sd
```{r}


for(i in fupmodvars){
  sub1=subset(msd, select=c(names(msd)==i))
  chemlistfupsc[,i]=(chemlistfupsc[,i]-sub1[1,1]) / sub1[2,1]
}

#Take out incomplete descriptors
fupscComplete=complete.cases(chemlistfupsc)
chemlistsc_fup_narm=chemlistfupsc[which(fupscComplete==TRUE),]
# Incompletescfup=chemlistfupsc[which(fupscComplete==FALSE),]
# write.csv(Incompletescfup,file=paste0("Incomplete_Opera_PaDEL_fup_predictions_",Datalabel,".csv"))


#Make Predictions with only five chemicals
set.seed(1255)
chemlistpredictfup=predict(RF.fit,chemlistsc_fup_narm)
chemlistsc_fup_narm$Fup.RF.pred=chemlistpredictfup^2 #square-root transformation applied to model resultsnames(chemlistsc_fup_narm$Fup.RF.pred)
```


### Apply applicability domain 
Using methodology of Roy et al 2015, apply applicability domain of fup model to predictions of new chemicals and export
```{r}
 a=chemlistsc_fup_narm
 b=a[,c("MoleculeID")]
 a=a[,which(names(a)%in%fupmodvars)]
 a=abs(a)
 SImin=apply(a,1,min) #Find the min, max, sd, and mean
 SImax=apply(a,1,max)
 SImean=apply(a,1,mean)
 SIsd=apply(a,1,sd)


  #Several are NA when SD is calculated
  SIdf=data.frame(SImin, SImax, SImean, SIsd)
  SIdf$SI90=SIdf$SImean + (1.28 * SIdf$SIsd)
  SIdf$Outlier=0 #something is assumed to be inside the AD unless it it found to not be

    #Based Algorithm specificed in Roy et al. 2015)
    for(i in 1:length(SIdf[,1])){
      if(SIdf$SImax[i]<3){ next } else
        if (SIdf$SImin[i] > 3) {
          SIdf$Outlier[i] = 1 } else if
      (SIdf$SImean[i] + (1.28 * SIdf$SIsd[i]) > 3){
        SIdf$Outlier[i]=1}
    }

chemlistsc_fup_narm=data.frame(chemlistsc_fup_narm, SIdf)


total=length(chemlistsc_fup_narm$Outlier)
inside=sum(chemlistsc_fup_narm$Outlier==0)
outside=sum(chemlistsc_fup_narm$Outlier==1)
inside_per=inside/total*100
outside_per=outside/total*100

print(paste0(signif(inside_per,3),
             "% are in the Roy et al. (2015) estimated domain o applicability"))


write.csv(chemlistsc_fup_narm, file=paste0("Novel_fup_predictions_with_AD_", Datalabel,"_descs_from_Opera",Operaversion, ".csv"))



Summary=data.frame(Domain=c("Inside AD","Outside AD"),
                   Count=c(inside,outside))

```

# Plot applicability domain summary
```{r}
applicability_domain_plot <- ggplot(Summary, aes(x = "", y = Count, fill = Domain)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  ggtitle("Applicability Domain") +
  theme_void() +
  scale_fill_manual(values=c("Inside AD"="steelblue", "Outside AD"="lightcoral"))+
  theme(plot.title = element_text(size = 14,face="bold",hjust = 0.5))


ggsave("C:\\Users\\stabatab\\OneDrive - Environmental Protection Agency (EPA)\\Profile\\Desktop\\ExpoCast_Projec\\Project_Pred_Caco2_Clint_Fup\\invitroTKstats\\CyprotexTO1\\out_fup\\applicability_domain_pie_plot.jpg", plot = applicability_domain_plot, width = 6, height = 4, units = "in", dpi = 300)

print(applicability_domain_plot)



```

### Evaulation the model 
```{r}

library(readr)
library(caret)

test.set=read.xlsx("180_Cyprotex-fup-RED-Level4.xlsx")
test.set=test.set[test.set$DTXSID !="DTXSID5022306",]

testfup=test.set$Fup.Med
predfup=(chemlistsc_fup_narm$Fup.RF.pred)
testsummary <- postResample(pred = predfup, obs = testfup)
print(testsummary)

```
### Z-Score Heatmap of Physicochemical Properties Across Cyprotex Compounds
```{r}

# Load necessary libraries
library(ggplot2)
library(reshape2)
library(pheatmap)
library(tidyr)


data=fupchemlist
data$Compound.Name=test.set$Compound.Name
data <- data[, -1]
data=data[,c(ncol(data), 1:(ncol(data)-1))]

# Remove non-numeric columns (Assuming the first column is chemical names)
descriptors <- data[, -1]  
chemical_names <- data[, 1]  


# Ensure the first column is compound names
rownames(data) <- data$Compound.Name 
data <- data[, -1] 

# Compute Z-scores for each descriptor
z_scores <- scale(data)

# Convert to dataframe
z_scores_df <- as.data.frame(z_scores)
z_scores_df$Compound <- rownames(z_scores_df)  

# Reshape for ggplot heatmap

z_melted <- z_scores_df %>% 
  pivot_longer(cols=!Compound, names_to = "Descriptors",values_to = "Z_Scores")

z_melted=as.data.frame(z_melted)

  
# Create the heatmap

heatmap_fup=ggplot(z_melted, aes(x = Compound, y = Descriptors, fill = Z_Scores)) +
  geom_tile() +
  scale_fill_gradient2(low = "#1E90FF", mid = "#E0E0E0", high = "#D50000", midpoint = 0) +
  theme_minimal() +
  labs(title = "Z-Score Heatmap of Physicochemical Properties Across Cyprotex Compounds", x = "Compounds", y = "Descriptors", fill = "Z_Scores") +
  theme(
     plot.background = element_rect(fill = "white",color=NA),
     panel.background = element_rect(fill = "white",color=NA),
     axis.text.x = element_text(angle = 90, vjust=0.5,hjust = 1, size = 5,face="bold"),
        axis.text.y = element_text(size = 8),
        plot.title = element_text(hjust = 0.5),
        legend.position = "right"
        
      ) + 
     coord_fixed((ratio=2))

ggsave("heatmap_z.score_fup.png", width=16,height = 10,dpi=700)

ggsave(
  filename  = paste0(setwd("C:/Users/stabatab/OneDrive - Environmental Protection Agency (EPA)/Profile/Desktop/ExpoCast_Projec/Project_Pred_Caco2_Clint_Fup/invitroTKstats/CyprotexTO1/out_Fup"),"heatmap_fup.png"),
  plot = heatmap_fup,
  height = 7, width = 7, units = "in", dpi = 300
)


```
### add a column for prediction and measured value
```{r}

overall=data.frame(Compound=test.set$Compound.Name,
                   Measured=testfup,
                   Predicted=predfup)

data_long=pivot_longer(overall,cols=c(Measured, Predicted), names_to = "Type", values_to = "fup_Value")



heatmap_ms.vs.pred.fup=ggplot(data_long, aes(x = Compound, y = Type, fill = fup_Value)) +
  geom_tile() +
  scale_fill_gradient2(low = "cyan", mid = "#E0E0E0", high = "deeppink3", midpoint =0,
                       limits=c(0,1),
                       breaks=seq(0,1,by=0.25),
                       labels=c("0","0.25","0.5","0.75","1"),
                       name="fup_Value") +
  theme_minimal() +
  labs(title = "Heatmap of Predicted vs. Measured fup Values Across Cyprotex Compounds ",
       x = "Compounds",
       y = "fup values",
       fill = "fup Value") +
   theme(
     plot.background = element_rect(fill = "white",color=NA),
     panel.background = element_rect(fill = "white",color=NA),
     axis.text.x = element_text(angle = 90, vjust=0.5,hjust = 1, size = 5,face="bold"),
        axis.text.y = element_text(vjust=0.5,hjust = 1,size = 10),
        plot.title = element_text(hjust = 0.5),
        legend.position = "right"
        
      ) + 
     coord_fixed((ratio=2))


ggsave("heatmap_ms.vs.pred.fup.png", width=16,height = 10,dpi=700)

ggsave(
  filename  = paste0(setwd("C:/Users/stabatab/OneDrive - Environmental Protection Agency (EPA)/Profile/Desktop/ExpoCast_Projec/Project_Pred_Caco2_Clint_Fup/invitroTKstats/CyprotexTO1/out_Fup"),"heatmap_ms.vs.pred.fup.png"),
  plot = heatmap_ms.vs.pred.fup,
  height = 7, width = 7, units = "in", dpi = 300
)




```

### check padels
```{r}
naPadel=colSums(is.na(Padel))
ind=which(naPadel > 0)
Na_namespadel=names(ind)
rownames(RF.fit$importance)
rmv=str_remove(rownames(RF.fit$importance), "^padel_|^Opera_")

Padelind=which(rmv %in% Na_namespadel)
padelcomimp=data.frame(rmv[Padelind])



```
### Distribution log p
```{r}


Trainfup=read.xlsx("S2_Dawson et al. Supporting_Information_Revision_Final_Sharing.xlsx", sheet="S2.4. fup_Training_Sets")

library(ggplot2)

# Combine datasets with a new 'group' column
Lp_Ts=data.frame(column_name=chemlistsc_fup_narm[["Opera_LogP"]])
Lp_Tr=data.frame(column_name=Trainfup[["Opera_LogP"]])
Lp_Ts$group <- "New 180 Compounds"
Lp_Tr$group <- "Training Set"

# Merge both datasets
combined_df <- rbind(Lp_Ts, Lp_Tr)

# Plot histogram with transparency to show overlap
histogram <- ggplot(combined_df, aes(x = column_name, fill = group)) +
  geom_histogram(data = Lp_Tr, aes(y = ..count.., fill = "Training Set"), 
                 bins = 30, alpha = 0.5, position = "identity") +
  geom_histogram(data = Lp_Ts, aes(y = ..count.., fill = "New 180 Compounds"), 
                 bins = 30, alpha = 0.7, position = "identity") +
  
  scale_fill_manual(name = "Group", values = c("Training Set" = "blue", 
                                               "New 180 Compounds" = "red")) +
  
  labs(title = "Comparison of Opera_LogP Distribution",
       x = "Opera_LogP",
       y = "Number of Chemicals") +
  
  theme_minimal()

# Print the histogram
print(histogram)

s=summary(Trainfup$Opera_LogP)
sdf=data.frame(Dist_logP_Training=names(s), value=as.numeric(s))

write.csv(sdf, file=paste0("Dis_LogP_Training.csv"))


ims=(RF.fit$importance)
ims_df=data.frame(Impostance=rownames(ims), value=as.vector(ims))

write.csv(ims_df, file=paste0("Importance_disc_Training.csv"))



```
###graph
```{r}

library(ggplot2)
library(scales)


fup=test.set[,c("DTXSID", "Fup.Med")]

Newtest.set=merge(fup,chemlistsc_fup_narm, by.x="DTXSID", by.y="MoleculeID", all.x=TRUE)

FigQSPRClint <- ggplot(Newtest.set) +
  geom_point(aes(x = as.numeric(testfup),
                 y = as.numeric(predfup),
             color = "Cyprotex"), alpha = 0.8, size = 5) +
  geom_abline(slope = 1, linetype = "dashed") +
  labs(y = expression(bold("QSPR Predicted ")~bolditalic("Fup")~bold('(unitless)')),
      x = expression(bold("Measured ")~bolditalic("Fup")~bold('(unitless)')),
       color = "Data Source") +
  scale_x_log10(labels = scientific_format()) +
  scale_y_log10(labels = scientific_format()) +
  scale_colour_manual(values=c("Cyproex"="purple"))+
  scale_colour_viridis_d() +
  theme_minimal() +
  theme(
    legend.position = "top",  # Move legend to the top
    legend.title = element_text(size = 18, face = "bold"),  # Adjust legend text
    legend.text = element_text(size = 16),
    axis.text = element_text(size=18),
    axis.title = element_text(size=20,face="bold"),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    axis.line = element_line(color="black",size=1),
    text = element_text(size = 16)  # Adjust text size
  ) +
  
  expand_limits(x = c(1e-2, 1e4), y = c(1e-2, 1e4))  # Ensure all points are shown

# **Save the plot**
ggsave(
  filename  = paste0(setwd("C:/Users/stabatab/OneDrive - Environmental Protection Agency (EPA)/Profile/Desktop/ExpoCast_Projec/Project_Pred_Caco2_Clint_Fup/invitroTKstats/CyprotexTO1/out_fup"),"/FigQSPRfup.png"),
  plot = FigQSPRClint,
  height = 7, width = 7, units = "in", dpi = 600
)




```




