---
title: "Pearce et al. (2017): Evaluation of Tissue Partitioning"
author: "Robert Pearce"
date: "April 30, 2023"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{c) Pearce (2017): Evaluation of Tissue Partitioning}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
*Please send questions to wambaugh.john@epa.gov*

from "Evaluation and calibration of high-throughput predictions of chemical 
distribution to tissues"

Robert G. Pearce, R. Woodrow Setzer, Jimena L. Davis,and John F. Wambaugh 

Journal of Pharmacokinetics and Pharmacodynamics volume 44, pages549–565 (2017)

https://doi.org/10.1007/s10928-017-9548-7

## Abstract

Toxicokinetics (TK) provides critical information for integrating chemical 
toxicity and exposure assessments in order to determine potential chemical risk 
(i.e., the margin between toxic doses and plausible exposures). For thousands of 
chemicals that are present in our environment, in vivo TK data are lacking. The 
publicly available R package “httk” (version 1.8, named for “high throughput 
TK”) draws from a database of in vitro data and physico-chemical properties in 
order to run physiologically-based TK (PBTK) models for 553 compounds. The PBTK 
model parameters include tissue:plasma partition coefficients (Kp) which the 
httk software predicts using the model of Schmitt 
(Toxicol In Vitro 22 (2):457–467, 2008). In this paper we evaluated and modified 
httk predictions, and quantified confidence using in vivo literature data. We 
used 964 rat Kp measured by in vivo experiments for 143 compounds. Initially, 
predicted Kp were significantly larger than measured Kp for many lipophilic 
compounds (log10 octanol:water partition coefficient > 3). Hence the approach 
for predicting Kp was revised to account for possible deficiencies in the in 
vitro protein binding assay, and the method for predicting membrane affinity 
was revised. These changes yielded improvements ranging from a factor of 10 to 
nearly a factor of 10,000 for 83 Kp across 23 compounds with only 3 Kp worsening 
by more than a factor of 10. The vast majority (92%) of Kp were predicted within 
a factor of 10 of the measured value (overall root mean squared error of 0.59 on 
log10-transformed scale). After applying the adjustments, regressions were
performed to calibrate and evaluate the predictions for 12 tissues. Predictions 
for some tissues (e.g., spleen, bone, gut, lung) were observed to be better 
than predictions for other tissues (e.g., skin, brain, fat), indicating that 
confidence in the application of in silico tools to predict chemical 
partitioning varies depending upon the tissues involved. Our calibrated model 
was then evaluated using a second data set of human in vivo measurements of 
volume of distribution (Vss) for 498 compounds reviewed by Obach et al. (Drug 
Metab Dispos 36(7):1385–1405, 2008). We found that calibration of the model 
improved performance: a regression of the measured values as a function of the 
predictions has a slope of 1.03, intercept of − 0.04, and R2 of 0.43. Through 
careful evaluation of predictive methods for chemical partitioning into tissues, 
we have improved and calibrated these methods and quantified confidence for TK 
predictions in humans and rats.

## HTTK Version

This vignette was created with **httk** v1.6. It was updated to httk v2.2.3 in
April of 2023. Although we attempt to maintain 
backward compatibility, if you encounter issues with the latest release of 
**httk**
and cannot easily address the changes, historical versions of httk are 
available from: https://cran.r-project.org/src/contrib/Archive/httk/

## Prepare for session
R package **knitr** generates html and PDF documents from this RMarkdown file,
Each bit of code that follows is known as a "chunk". We start by telling 
**knitr** how we want our chunks to look.
```{r knitrPrep, include=FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = TRUE, fig.width=5, fig.height=4)
```

### Clear the memory
It is a bad idea to let variables and other information from previous R
sessions float around, so we first remove everything in the R memory.
```{r clear_memory, eval = TRUE}
rm(list=ls()) 
```

### eval = execute.vignette
If you are using the RMarkdown version of this vignette (extension, .RMD) you
will be able to see that several chunks of code in this vignette
have the statement "eval = execute.vignette". The next chunk of code, by default,
sets execute.vignette = FALSE.  This means that the code is included (and necessary) but was
not run when the vignette was built. We do this because some steps require 
extensive computing time and the checks on CRAN limit how long we can spend
building the package. If you want this vignette to work, you must run all code,
either by cutting and pasting it into R. Or, if viewing the .RMD file, you can
either change execute.vignette to TRUE or press "play" (the green arrow) on
each chunk in RStudio.
```{r runchunks, eval = TRUE}
# Set whether or not the following chunks will be executed (run):
execute.vignette <- TRUE
```

### Load the relevant libraries
We use the command 'library()' to load various R packages for our analysis.
If you get the message "Error in library(X) : there is no package called 'X'"
then you will need to install that package: 

From the R command prompt:

install.packages("X")

Or, if using RStudio, look for 'Install Packages' under 'Tools' tab.

```{r load_libraries, eval=execute.vignette}
library(httk)
library(gdata)
library(ggplot2)
library(viridis)
library(censReg)
library(gmodels)
library(gplots)
library(scales)
library(colorspace)
library(gridExtra)
```
### Function to format scientific notation 
From https://stackoverflow.com/questions/10762287/how-can-i-format-axis-labels-with-exponents-with-ggplot2-and-scales
```{R scientific.notation, eval = execute.vignette}
scientific_10 <- function(x) {                                  
  out <- gsub("1e", "10^", scientific_format()(x))              
  out <- gsub("\\+","",out)                                     
  out <- gsub("10\\^01","10",out)                               
  out <- parse(text=gsub("10\\^00","1",out))                    
}  
```

We first filter the measured rat Kp data, pc.data. Then the old and new Kp 
predictions are made, along with error and improvement measures, 
and these are all consolidated into a table for analysis and plotting. Note 
that the final table contains log10-transformed values and error 
and improvements derived from subtracting these values. Only relevant rat 
values are used. Compounds with Funbound.plasma and partition 
coefficients of zero are removed as well as compounds with approximated 
Funbound.plasma values.

```{r setuppctable, eval=execute.vignette}
pc.table <- NULL
pc.data <- subset(pc.data,fu != 0 & Exp_PC != 0 & Tissue %in% c("Adipose","Bone","Brain","Gut",
    "Heart","Kidney","Liver","Lung","Muscle","Skin","Spleen","Blood Cells") & 
    tolower(Species) == 'rat' & !CAS %in% c('10457-90-6','5786-21-0','17617-23-1','69-23-8','2898-12-6',
    '57562-99-9','59-99-4','2955-38-6','155-97-5','41903-57-5','58-55-9','77-32-7','59-05-2','60-54-8'))
cas.list <- get_cheminfo(model='schmitt',species='rat',suppress.messages=TRUE)
cas.list <-  cas.list[cas.list %in% pc.data[,'CAS']]
ma.data.list <- subset(chem.physical_and_invitro.data,!is.na(logMA))[,'CAS']
for(this.cas in cas.list){
  parameters <- parameterize_schmitt(
    chem.cas=this.cas,
    species='rat',
    suppress.messages=TRUE)
  init.parameters <- parameters
  charge <- calc_ionization(
    chem.cas=this.cas,
    pH=7.4)$fraction_charged
  if(!this.cas %in% ma.data.list){
    init.parameters$MA <- 10^(0.999831 - 0.016578*38.7 + 0.881721 * log10(parameters$Pow))
  }
  pcs <- predict_partitioning_schmitt(
    parameters=parameters,
    species='rat',
    regression=FALSE,
    suppress.messages=TRUE
    )
  init.pcs <- predict_partitioning_schmitt(
    parameters=init.parameters,
    species='rat',
    regression=FALSE,
    suppress.messages=TRUE)
  for(this.tissue in subset(pc.data,CAS==this.cas)[,'Tissue']){
    if(this.tissue == 'Blood Cells') this.pc <- 'rbc'
    else this.pc <- this.tissue
        pc.table <- rbind(pc.table,
                      cbind(
                        as.data.frame(this.cas),
                        as.data.frame(this.tissue),
                        as.data.frame(log10(
                          init.pcs[[which(substr(names(init.pcs),
                            2,
                            nchar(names(init.pcs))-3) == 
                            tolower(this.pc))]] * 
                          init.parameters$Funbound.plasma)),
                        as.data.frame(log10(
                          pcs[[which(substr(names(pcs),
                            2,
                            nchar(names(pcs))-3) == 
                            tolower(this.pc))]] *
                          parameters$unadjusted.Funbound.plasma)),
                        as.data.frame(log10(
                          init.pcs[[which(substr(names(init.pcs),
                            2,
                            nchar(names(init.pcs))-3) == 
                            tolower(this.pc))]] *
                          init.parameters$unadjusted.Funbound.plasma)),
                        as.data.frame(log10(
                          pcs[[which(substr(names(pcs),
                            2,
                            nchar(names(pcs))-3) == tolower(this.pc))]] * 
                            parameters$Funbound.plasma)),
                        as.data.frame(log10(
                          subset(pc.data,
                            CAS==this.cas & Tissue==this.tissue)[,'Exp_PC'])),
                        as.data.frame(subset(pc.data,
                          CAS==this.cas & Tissue==this.tissue)[,'LogP']),
                        as.data.frame(charge),
                        as.data.frame(as.character(subset(pc.data,
                          CAS == this.cas)[1,'A.B.N'])),
                        as.data.frame(subset(pc.data, 
                          CAS == this.cas)[1,'fu'])))
  }
}
colnames(pc.table) <- c('CAS','Tissue',
                        'fup.correction',
                        'ma.correction',
                        'init.Predicted',
                        'Predicted',
                        'Experimental',
                        'logP',
                        'charge',
                        'type',
                        'fup')
init.error <- pc.table[,'Experimental'] - pc.table[,'init.Predicted']
fup.error <- pc.table[,'Experimental'] - pc.table[,'fup.correction']
ma.error <- pc.table[,'Experimental'] - pc.table[,'ma.correction']
final.error <- pc.table[,'Experimental'] - pc.table[,'Predicted']
fup.improvement <- abs(init.error) - abs(fup.error)
ma.improvement <- abs(init.error) - abs(ma.error)
final.improvement <- abs(init.error) - abs(final.error)
pc.table <- cbind(pc.table,fup.improvement,ma.improvement, final.improvement,
                  final.error,init.error,ma.error,fup.error)
```

```{r KpFigures, eval=execute.vignette}
init.plot <- ggplot() + 
  geom_point(data=pc.table,aes(10^(init.Predicted),10^(Experimental))) +
  geom_abline() + 
  labs(y=expression(paste("Measured ",K[p])), 
       x=expression(paste("Predicted ",K[p]))) +
  theme(axis.text=element_text(size=16),axis.title=element_text(size=16),
    plot.title=element_text(size=18,hjust = 0.5)) + 
  scale_x_log10(label=scientific_10,limits=c(0.01,10^4.5)) +
  scale_y_log10(label=scientific_10,limits=c(0.01,10^4.5)) + 
  ggtitle('(A)')
print(init.plot)
init.stats <- summary(lm(Experimental ~ init.Predicted, 
                         data=pc.table))

final.plot <- ggplot() + 
  geom_point(data=pc.table,aes(10^(Predicted),10^(Experimental))) +
  geom_abline() + 
  labs(y=expression(paste("Measured ",K[p])),
       x=expression(paste("Predicted ",K[p]))) +
  theme(axis.text=element_text(size=16),axis.title=element_text(size=16),
    plot.title=element_text(size=18,hjust=0.5)) +
  scale_x_log10(label=scientific_10,limits=c(0.01,10^4.5)) +
  scale_y_log10(label=scientific_10,limits=c(0.01,10^4.5)) + 
  ggtitle('(B)')
print(final.plot)
final.stats <- summary(lm(Experimental ~ Predicted, 
                          data=pc.table))

fup.change.plot <-  ggplot() +
  geom_point(data=pc.table[order(pc.table[,'fup.improvement'],decreasing=F),],
    aes(10^(fup.correction),10^(Experimental),color=fup.improvement)) + 
  geom_abline() +
  labs(y=expression(paste("Measured ",K[p])),
       x=expression(paste("Predicted ",K[p])),color='Improvement') +
  theme(axis.text=element_text(size=16),axis.title=element_text(size=16)) +
  scale_x_log10(label=scientific_10,limits=c(0.01,10^4.5)) + 
  scale_y_log10(label=scientific_10,limits=c(0.01,10^4.5)) +
  scale_color_viridis(direction=-1,option='inferno')
print(fup.change.plot)
fup.stats <- summary(lm(Experimental ~ fup.correction, 
                     data=pc.table))

ma.subset <- subset(pc.table,!CAS %in% ma.data.list)
ma.change.plot <- ggplot() +
    geom_point(data=ma.subset[order(ma.subset[,'ma.improvement']
      ,decreasing=F),], 
      aes(10^(ma.correction),10^(Experimental),color=ma.improvement)) + 
    geom_abline() +
    labs(y=expression(paste("Measured ",K[p])),
      x=expression(paste("Predicted ",K[p])),color='Improvement') +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16)) +
    scale_x_log10(label=scientific_10,limits=c(0.01,10^4.5)) + 
    scale_y_log10(label=scientific_10,limits=c(0.01,10^4.5)) +
    scale_color_viridis(direction=-1,option='inferno')
print(ma.change.plot)
ma.stats <- summary(lm(Experimental ~ ma.correction, 
                    data=ma.subset))

fup.table <- data.frame(Test=c("Initial Tissue PC Accuracy","Fup Lipid Correction","Membrane Affinity","Final"),
                           RSquared = signif(c(init.stats$adj.r.squared,
                                         fup.stats$adj.r.squared,
                                         ma.stats$adj.r.squared,
                                         final.stats$adj.r.squared),3),
                           RMSLE = signif(c(mean(init.stats$residuals^2,na.rm=TRUE)^(1/2),
                                     mean(fup.stats$residuals^2,na.rm=TRUE)^(1/2),
                                     mean(ma.stats$residuals^2,na.rm=TRUE)^(1/2),
                                     mean(final.stats$residuals^2,na.rm=TRUE)^(1/2)),3)
                          )
knitr::kable(fup.table)
```
Now we calculate and plot the regressions for all tissues, together with their 95% confidence intervals.

```{r performPCregressions, eval=execute.vignette}
regressions <- NULL

for(tissue in as.character(unique(pc.table[,'Tissue']))){
  fit <- lm(Experimental ~ Predicted ,data=subset(pc.table,Tissue==tissue))
  smry <- summary(fit)
  est <- estimable(fit, cm=diag(2), beta0=c(0,1), joint.test=TRUE)
  regressions <- rbind(regressions,cbind(tissue,as.data.frame(fit$coefficients[['(Intercept)']]),
      as.data.frame(fit$coefficients[['Predicted']]),
      as.data.frame(smry$coefficients[['Predicted','Pr(>|t|)']]),
      as.data.frame(smry$sigma),as.data.frame(smry$r.squared),
      as.data.frame(smry[[11]][1,1]),as.data.frame(smry[[11]][2,2]),
      as.data.frame(smry[[11]][1,2]),as.data.frame(smry$df[2]),as.data.frame(est[[3]])))
}
colnames(regressions) <- c('Tissue','Intercept','Slope','P-value','SE','R-squared',
                           'Int Var','Slp Var','Cov','df','estimable')

for (this.col in 2:10) regressions[,this.col] <- signif(regressions[,this.col], 3)
regressions <- regressions[order(regressions[,1]),]

knitr::kable(regressions, caption = "Table 1: The regressions for each tissue, after fup and membrane affinity adjustments, of the log10-transformed measured Kp regressed on
predicted Kp")

write.table(regressions,
            file=paste0("Pearce2017PCCalibration=",Sys.Date(),".txt"),
            sep="/t",
            row.names=FALSE)
```

```{r PCRegressionFigure, eval=execute.vignette}
x.cf <- seq(-2,3.5,.01)
for(tissue in  as.character(unique(pc.table[,'Tissue'])))
{
  conf <- qt(0.975,df=subset(regressions,Tissue==tissue)[['df']]+1) *
      subset(regressions,Tissue==tissue)[['SE']] *
      sqrt(subset(regressions,Tissue==tissue)[['Int Var']] +
      x.cf^2 * subset(regressions,Tissue==tissue)[['Slp Var']] +
      2 * x.cf * subset(regressions,Tissue==tissue)[['Cov']] + 1)
  line <- subset(regressions,Tissue==tissue)[['Intercept']] +
      x.cf * subset(regressions,Tissue==tissue)[['Slope']]
  y.cf <- line + conf
  y.ncf <- line - conf

  cf <- cbind(as.data.frame(x.cf),as.data.frame(y.cf),as.data.frame(y.ncf))
  if(tissue == 'Blood Cells'){
    eval(parse(text= paste('Blood <- ggplot() + 
                               geom_abline(linetype = "dashed") +
                               geom_point(data=subset(pc.table,Tissue == \'',tissue,'\'),aes(10^(Predicted),
                               10^(Experimental)))  +  theme(axis.text=element_text(size=14),
                               axis.title=element_text(size=14),plot.title=element_text(size=14)) +
                               scale_x_log10(label=scientific_10,limits=c(0.01,1000)) + 
                               scale_y_log10(label=scientific_10,limits=c(0.01,1000)) +
                               ylab(ifelse(tissue=="Brain",expression(paste("Inferred ",K[p])),"")) +
                               xlab(ifelse(tissue=="Skin", expression(paste("Predicted ",K[p])),"")) +
                               geom_line(data=cf,aes(10^(x.cf),10^(y.cf))) +
                               geom_line(data=cf,aes(10^(x.cf),10^(y.ncf))) +
                               geom_abline(intercept=subset(regressions,Tissue==tissue)[[\'Intercept\']],
                               slope=subset(regressions,Tissue==tissue)[[\'Slope\']]) +
                               ggtitle(\'Red Blood Cells\')',sep='')))
  }else{
    eval(parse(text= paste(tissue,' <- ggplot() + labs(y=expression(paste("Measured ",K[p]))
                               ,x=expression(paste("Predicted ",K[p]))) + geom_abline(linetype = "dashed") +
                               geom_point(data=subset(pc.table,Tissue == \'',tissue,'\'),
                               aes(10^(Predicted),10^(Experimental))) + theme(axis.text=element_text(size=14),
                               axis.title=element_text(size=14),plot.title=element_text(size=14)) +
                               scale_x_log10(label=scientific_10,limits=c(0.01,1000)) + 
                               scale_y_log10(label=scientific_10,limits=c(0.01,1000)) +
                               ylab(ifelse(tissue=="Brain",expression(paste("Inferred ",K[p])),"")) +
                               xlab(ifelse(tissue=="Skin", expression(paste("Predicted ",K[p])),"")) + 
                               geom_line(data=cf,aes(10^(x.cf),10^(y.cf))) +
                               geom_line(data=cf,aes(10^(x.cf),10^(y.ncf))) +
                               geom_abline(intercept=subset(regressions,Tissue==tissue)[[\'Intercept\']],
                               slope=subset(regressions,Tissue==tissue)[[\'Slope\']]) +
                               ggtitle(\'',tissue,'\')',sep='')))
  }
}

grid.arrange(Adipose, Blood, Bone, Brain, Gut, Heart, Kidney, Liver, Lung, Muscle, Skin, Spleen, nrow=4)
```
In vivo volume of distribution data are compared with the predictions, and 
errors are calculated.  Regressions and improvements are calculated and plotted.

```{r Vdevalaution, eval=execute.vignette}
obach <- subset(Obach2008,CAS %in% get_cheminfo(model='schmitt'))
vd.table <- NULL

for(this.cas in obach[,'CAS']){
  parameters <- parameterize_schmitt(
    chem.cas=this.cas,
    suppress.messages=TRUE)
  init.parameters <- parameters
  if(!this.cas %in% ma.data.list){
    init.parameters$MA <- 10^(0.999831 - 0.016578*37 + 0.881721 * log10(parameters$Pow))
  }
  pcs <- predict_partitioning_schmitt(
    parameters=parameters,
    regression=FALSE,
    suppress.messages=TRUE)
  init.pcs <- predict_partitioning_schmitt(
    parameters=init.parameters,
    regression=FALSE,
    suppress.messages=TRUE)
  reg.pcs <- predict_partitioning_schmitt(
    parameters=parameters,
    regression=TRUE,
    suppress.messages=TRUE)
  vdist <- calc_vdist(
    parameters=c(pcs,Funbound.plasma=parameters$Funbound.plasma),
    suppress.messages=TRUE)
  init.vdist <- calc_vdist(
    parameters=c(
      init.pcs,
      Funbound.plasma=parameters$unadjusted.Funbound.plasma),
    suppress.messages = TRUE)
  reg.vdist <- calc_vdist(
    parameters=c(reg.pcs,Funbound.plasma=parameters$Funbound.plasma),
    suppress.messages = TRUE)
  vd.table <- rbind(
    vd.table,
    cbind(as.data.frame(this.cas),as.data.frame(log10(init.vdist)),
    as.data.frame(log10(vdist)),as.data.frame(log10(reg.vdist)),
    as.data.frame(log10(subset(obach,CAS==this.cas)[['VDss (L/kg)']]))))
}
colnames(vd.table) <- c(
  'CAS',
  'init.vdist',
  'corrected.vdist',
  'calibrated.vdist',
  'Experimental')
init.error <- vd.table[,'Experimental'] - vd.table[,'init.vdist']
correction.error <- vd.table[,'Experimental'] - vd.table[,'corrected.vdist']
calibration.error <- vd.table[,'Experimental'] - vd.table[,'calibrated.vdist']
correction.improvement <- abs(init.error) - abs(correction.error)
calibration.improvement <- abs(correction.error) - abs(calibration.error)
vd.table <- cbind(vd.table,correction.improvement,calibration.improvement,
                  init.error,correction.error,calibration.error)
```

```{r VdistRegressions, eval=execute.vignette}
fit <- lm(Experimental ~ calibrated.vdist,data=vd.table)
smry <- summary(fit)
calibrated.reg <- cbind(as.data.frame(fit$coefficients['(Intercept)']),
                        as.data.frame(fit$coefficients['calibrated.vdist']),
                        as.data.frame(smry$coefficients['calibrated.vdist','Pr(>|t|)']),
                        as.data.frame(smry$sigma),as.data.frame(smry$r.squared))
fit <- lm(Experimental ~ init.vdist,data=vd.table)
smry <- summary(fit)
init.reg <- cbind(as.data.frame(fit$coefficients['(Intercept)']),
                  as.data.frame(fit$coefficients['init.vdist']),
                  as.data.frame(smry$coefficients['init.vdist','Pr(>|t|)']),
                  as.data.frame(smry$sigma),as.data.frame(smry$r.squared))
fit <- lm(Experimental ~ corrected.vdist,data=vd.table)
smry <- summary(fit)
corrected.reg <- cbind(as.data.frame(fit$coefficients['(Intercept)']),
                       as.data.frame(fit$coefficients['corrected.vdist']),
                       as.data.frame(smry$coefficients['corrected.vdist','Pr(>|t|)']),
                       as.data.frame(smry$sigma),as.data.frame(smry$r.squared))
colnames(init.reg) <- colnames(corrected.reg) <-
colnames(calibrated.reg) <- c('Intercept','Slope','P-value','Std Err','R-squared')
```

```{r VdistPlots, eval=execute.vignette}
init.vd.plot <- ggplot(vd.table,aes(10^(init.vdist),10^(Experimental))) + geom_point() +
    geom_abline(intercept = init.reg[['Intercept']], slope = init.reg[["Slope"]]) +
    geom_abline(linetype = "dashed") + xlab("Predicted Volume of Distribution") +
    ylab("Measured Volume of Distribution") + theme(axis.text=element_text(size=16),
    axis.title=element_text(size=16),plot.title=element_text(size=18,hjust = 0.5)) +
    scale_x_log10(label=scientific_10,limits=c(10^(-1.5),10^(8.5))) + 
    scale_y_log10(label=scientific_10,limits=c(10^(-1.5),10^(8.5))) +
    ggtitle('(A)')
print(init.vd.plot)

correction.plot <- ggplot() +
    geom_point(data=vd.table[order(vd.table[,'correction.improvement'],decreasing=F),],
    aes(10^(corrected.vdist),10^(Experimental),color=correction.improvement)) +
    geom_abline(intercept = corrected.reg[['Intercept']],slope = corrected.reg[["Slope"]]) +
    geom_abline(linetype = "dashed") + xlab("Predicted Volume of Distribution") +
    ylab("Measured Volume of Distribution") + theme(legend.position = c(.95, .95),
    legend.justification = c("right", "top"),legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),axis.text=element_text(size=16),
    axis.title=element_text(size=16),plot.title=element_text(size=18,hjust = 0.5)) +
    scale_x_log10(limits=c(10^(-1.5),10^(3))) + scale_y_log10(limits=c(10^(-1.5),10^(3))) +
    ggtitle('(B)') + scale_color_viridis(direction=-1,option='inferno')
print(correction.plot)

calibration.plot <- ggplot() +
    geom_point(data=vd.table[order(vd.table[,'calibration.improvement'],decreasing=F),],
    aes(10^(calibrated.vdist),10^(Experimental),color=calibration.improvement)) +
    geom_abline(intercept = calibrated.reg[['Intercept']],slope = calibrated.reg[["Slope"]]) +
    geom_abline(linetype = "dashed") + xlab("Predicted Volume of Distribution") +
    ylab("Measured Volume of Distribution") + theme(legend.position = c(.95, .95),
    legend.justification = c("right", "top"),legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6),axis.text=element_text(size=16),
    axis.title=element_text(size=16),plot.title=element_text(size=18,hjust = 0.5)) +
    scale_x_log10(limits=c(10^(-1.5),10^(3))) + scale_y_log10(limits=c(10^(-1.5),10^(3))) +
    ggtitle('(C)') + scale_color_viridis(direction=-1,option='inferno')
print(calibration.plot)
```

Now we pull in the in vivo blood to plasma ratios, use these to calculate the 
inferred red blood cell to plasma ratios, and then make predictions for these 
values.
A censored regressions is performed, and predictions are plotted against errors.

```{r bloodtoplasmaevaluation, eval=execute.vignette}
rb2p.data <- subset(chem.physical_and_invitro.data,!is.na(Human.Rblood2plasma))
measured.rb2p <- NULL
measured.krbc <- NULL
predicted.rb2p <- NULL
predicted.krbc <- NULL
cas <- NULL
charge <- NULL
fup <- NULL
logP <- NULL
pka_donor <- NULL
pka_accept <- NULL
for(this.cas in rb2p.data[rb2p.data[,'CAS'] %in% 
  get_cheminfo(model='schmitt', suppress.messages = TRUE),'CAS'])
{
  rb2p <- get_rblood2plasma(chem.cas=this.cas)
  krbc <- (rb2p + .44 - 1) / 0.44
  measured.rb2p <- c(measured.rb2p,rb2p)
  measured.krbc <- c(measured.krbc,krbc)
  parameters <- parameterize_schmitt(
    chem.cas=this.cas,
    suppress.messages = TRUE)
  pcs <- predict_partitioning_schmitt(
    parameters=parameters,
    suppress.messages=TRUE)
  predicted.krbc <- c(predicted.krbc,pcs[['Krbc2pu']] * parameters$Funbound.plasma)
  cas <- c(cas,this.cas)
  charge <- c(charge,calc_ionization(chem.cas=this.cas,pH=7.4)$fraction_charged)
  fup <- c(fup,parameters$unadjusted.Funbound.plasma)
  logP <-  c(logP,log10(parameters$Pow))
  pka_donor <- c(pka_donor,paste(parameters$pKa_Donor,collapse=','))
  pka_accept <- c(pka_accept,paste(parameters$pKa_Accept,collapse=','))
}
predicted.rb2p <-  1 - 0.44 + 0.44 * predicted.krbc
rb2p.table <- cbind(as.data.frame(cas),as.data.frame(predicted.rb2p),as.data.frame(measured.rb2p))
colnames(rb2p.table) <- c('cas','predicted.rb2p','measured.rb2p')
error <- log10(rb2p.table[,'measured.rb2p']) - log10(rb2p.table[,'predicted.rb2p'])
rb2p.table <- cbind(rb2p.table,error,charge,fup,logP)

error <- log10(measured.krbc) -  log10(predicted.krbc)
krbc.table <- cbind(as.data.frame(cas),as.data.frame(predicted.krbc),as.data.frame(measured.krbc),
                    as.data.frame(error),charge,fup,logP,pka_donor,pka_accept)
```

```{r RBCStats, eval=execute.vignette}
pdta <- data.frame(x = predicted.krbc,
                   y = measured.krbc)
pdta$y[pdta$y <= 0.1] <- 0.1
pdta$Censoring <- factor(c("Not Censored","Censored")[as.numeric(pdta$y <= 0.1) + 1])
y <- measured.krbc
x <- cbind(rep(1, length(y)),-1 * log10(predicted.krbc))
colnames(x) <- c("Intercept","Predicted")
cc <- as.numeric(y <= 0.1)
y[y < 0.1] <- 0.1
y <- -log10(y)

out <- censReg(y~x, data = pdta, left=0.1)
out$betas <- out$estimate
```

```{r RBCFigure, eval=execute.vignette}
censored.regression <- ggplot() +
    geom_point(data=pdta,aes(x=x,y=y, color=Censoring)) +
    scale_x_log10(limits=c(.0009,40)) + scale_y_log10(limits=c(.1,4),breaks=c(.1,.5,2.5)) +
    labs(y=expression(paste("Inferred ",K[p])),x=expression(paste("Predicted ",K[p]))) +
    geom_abline(intercept=0, slope=1, linetype='dashed') +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16),
    plot.title=element_text(size=18,hjust=0.5),legend.position = c(0.11, .8)) +
    geom_abline(slope=out$betas[2],intercept=-out$betas[1]) + ggtitle('(B)')
print(censored.regression)

rb2p.plot <- ggplot(rb2p.table,aes(predicted.rb2p,measured.rb2p)) +
    geom_point()  + scale_x_log10(lim=c(.52,18)) + 
    scale_y_log10(lim=c(.52,2.5),breaks=c(0.5,1,2)) + geom_abline(linetype='dashed') +
    labs(y=expression(paste("Measured Whole Blood ",K[p])),
    x=expression(paste("Predicted Whole Blood ",K[p]))) +
    theme(axis.text=element_text(size=16),axis.title=element_text(size=16),
    plot.title=element_text(size=18,hjust=0.5)) + ggtitle('(A)')
print(rb2p.plot)
```

Lastly, we make the heatmap.
```{r summarytable, eval=execute.vignetteE}
heatmap.table <- NULL
for(this.cas in get_cheminfo(model='schmitt')){
    parms <- parameterize_schmitt(
      chem.cas=this.cas,
      suppress.messages = TRUE)
    pcs <- predict_partitioning_schmitt(
      parameters=parms, 
      suppress.messages = TRUE)
    heatmap.table <- cbind(heatmap.table,log10(unlist(pcs)[1:11]*parms$Funbound.plasma))
}
rownames(heatmap.table) <-  c('Adipose','Bone','Brain','Gut','Heart',
                              'Kidney','Liver','Lung','Muscle','Skin','Spleen')
colnames(heatmap.table) <- rep("",dim(heatmap.table)[2])
```

```{r summaryheatmap, eval=execute.vignetteE}
pal <- function (n, h = c(260, -328), c = 80, l = c(30, 100), power = 1.5,
    fixup = TRUE, gamma = NULL, alpha = 1, ...)
{
    if (!is.null(gamma))
        warning("'gamma' is deprecated and has no effect")
    if (n < 1L)
        return(character(0L))
   h <- rep(h, length.out = 2L)
    c <- c[1L]
    l <- rep(l, length.out = 2L)
    power <- rep(power, length.out = 2L)
    rval <- seq(1, -1, length = n)
    rval <- hex(polarLUV(L = l[2L] - diff(l) * abs(rval)^power[2L],
        C = c * abs(rval)^power[1L], H = ifelse(rval > 0, h[1L],
            h[2L])), fixup = fixup, ...)
    if (!missing(alpha)) {
        alpha <- pmax(pmin(alpha, 1), 0)
        alpha <- format(as.hexmode(round(alpha * 255 + 1e-04)),
            width = 2L, upper.case = TRUE)
        rval <- paste(rval, alpha, sep = "")
    }
    return(rval)
}

hclust.ave <- function(x) hclust(x, method="ward.D2")
heatmap.2(heatmap.table,col=pal,trace="none", hclustfun=hclust.ave,
          key.xlab=expression(paste("log10 ",K[p]," Value")),
          key.ylab=expression(paste("Number of ",K[p])),
          key.title="Partition Coefficient",xlab="Chemicals",cex.lab=2,margins=c(2,5))
```