---
title: 'Scherer: Plot IVDisp Fractions'
author: "Meredith Scherer"
date: "2024-02-06"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{i) Scherer: Plot IVDisp Fractions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
*Please send questions to wambaugh.john@epa.gov*

## HTTK Version

This vignette was created with **httk** v2.3.0. Although we attempt to maintain backward compatibility, if you encounter issues with the latest release of **httk** and cannot easily address the changes, historical versions of httk are available from: https://cran.r-project.org/src/contrib/Archive/httk/

## Prepare for session
R package **knitr** generates html and PDF documents from this RMarkdown file,
Each bit of code that follows is known as a "chunk". We start by telling 
**knitr** how we want our chunks to look.
```{r configure_knitr, eval = TRUE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
options(rmarkdown.html_vignette.check_title = FALSE)
```

### Clear the memory
It is a bad idea to let variables and other information from previous R
sessions float around, so we first remove everything in the R memory.
```{r clear_memory, eval = TRUE}
rm(list=ls()) 
```

### eval = execute.vignette
If you are using the RMarkdown version of this vignette (extension, .RMD) you
will be able to see that several chunks of code in this vignette
have the statement "eval = execute.vignette". The next chunk of code, by default,
sets execute.vignette = FALSE.  This means that the code is included (and necessary) but was
not run when the vignette was built. We do this because some steps require 
extensive computing time and the checks on CRAN limit how long we can spend
building the package. If you want this vignette to work, you must run all code,
either by cutting and pasting it into R. Or, if viewing the .RMD file, you can
either change execute.vignette to TRUE or press "play" (the green arrow) on
each chunk in RStudio.
```{r runchunks, eval = TRUE}
# Set whether or not the following chunks will be executed (run):
execute.vignette <- FALSE
```

### Load the relevant libraries
If you get the message "Error in library(X) : there is no package called 'X'"
then you will need to install that package: 

From the R command prompt:

install.packages("X")

Or, if using RStudio, look for 'Install Packages' under 'Tools' tab.
```{r InitialPrep, eval = execute.vignette}
library(ggplot2)
```

## Create Pie Chart

This code produces a pie chart for an example in vitro distribution data set.

```{r MakeTable, eval = execute.vignette}

# Example data set with compartments and fraction present in each compartment

input_table <- data.frame(compartment = c("Free in medium", "Bound in plasma", "Associated with cells", "In headspace", "Soaked to well plastic"), fraction=c(0.1, 0.240597245, 0.600, 0.048, 0.01140276))

```


```{r MakePlot, eval = execute.vignette}

# Pie chart function to plot the data set
piechart(input_table)

```






