---
title: "Crizer et al. (2024): Analysis of PFAS Hepatic Clearance Data"
author: "David Crizer, Marci G. Smeltz, John F. Wambaugh, and Barbara A. Wetmore"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Crizer (2024): PFAS Clearance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  *Please send questions to wambaugh.john@epa.gov*
  
  from "Category-Based Toxicokinetic Evaluations of Data-Poor Per and Polyfluoroalkyl Substances (PFAS) by Gas Chromatography Coupled with Mass Spectrometry"

Anna Kreutz, Matthew Scott Clifton, W. Matthew Henderson, Marci G. Smeltz, John F. Wambaugh, and Barbara A. Wetmore

## Abstract

Concern over per- and polyfluoroalkyl substances (PFAS) has increased significantly as more is learned about their widespread environmental presence, persistence, and bioaccumulative potential. The limited measurement, monitoring, toxicokinetic (TK), and toxicologic data currently available is inadequate to inform risk evaluations across this diverse domain. In this study, 73 PFAS were selected for *in vitro* TK evaluation to expand our knowledge across lesser studied PFAS that include alcohol, amide, and acrylate functional groups. Targeted analytical methods were developed using gas chromatography coupled with tandem mass spectrometry (GC-MS/MS); ultracentrifugation was used to measure human plasma protein binding and human hepatocyte suspensions were used in a substrate depletion approach to measure in vitro clearance (Clint). Method development was successful in plasma for 43 PFAS, with fraction unbound in plasma ($f_{up}$) values ranging from 0.004 for 1,6-dibromododecafluorohexane and 1 for 2-aminohexafluoropropan-2-ol, With a median $f_{up}$ of 0.09 (i.e., 91% bound), these PFAS are highly bound, but exhibit 10-fold lower binding than that recently reported for legacy perfluoroalkyl acids (Smeltz et al., 2022). Physico-chemical property trends analyses revealed that increased binding was associated with increases in mass, van der Waals volumes, and values of Log Pow and Log D at pH 7.4. Thirty of the PFAS evaluated in the hepatocyte assay showed abiotic loss, with many exceeding 60% loss within 60 min of media addition. Metabolic clearance was noted for 11 of the 13 successfully evaluated, with rates up to 49.9 l/min/million cells for 1H,1H,10H,10H-Perfluorodecane-1,10-diol. The Chemical Transformation Simulator revealed potential metabolites and other transformation products to consider during *in vitro* assay data interpretation. This integration of quantitative analytic methods with *in vitro* toxicokinetics has provided critical learnings as New Approach Methods are considered to evaluate PFAS for which volatility, metabolism, and other routes of transformation are likely to modulate their environmental fate. 

## HTTK Version

This vignette was created with **httk** v2.3.0. Although we attempt to maintain 
backward compatibility, if you encounter issues with the latest release of 
**httk**
  and cannot easily address the changes, historical versions of httk are 
available from: https://cran.r-project.org/src/contrib/Archive/httk/
  
  ## Prepare for session
```{r}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Session Preparation

### Clear the memory
It is a bad idea to let variables and other information from previous R
sessions float around, so we first remove everything in the R memory.
```{r clear_memory, eval = TRUE}
rm(list=ls()) 
```

### eval = execute.vignette
If you are using the RMarkdown version of this vignette (extension, .RMD) you
will be able to see that several chunks of code in this vignette
have the statement "eval = execute.vignette". The next chunk of code, by default,
sets execute.vignette = FALSE.  This means that the code is included (and necessary) but was
not run when the vignette was built. We do this because some steps require 
extensive computing time and the checks on CRAN limit how long we can spend
building the package. If you want this vignette to work, you must run all code,
either by cutting and pasting it into R. Or, if viewing the .RMD file, you can
either change execute.vignette to TRUE or press "play" (the green arrow) on
each chunk in RStudio.
```{r runchunks, eval = TRUE}
# Set whether or not the following chunks will be executed (run):
execute.vignette <- TRUE
```

### Load the relevant libraries
If you get the message "Error in library(X) : there is no package called 'X'"
then you will need to install that package: 
  
  From the R command prompt:
  
  install.packages("X")

Or, if using RStudio, look for 'Install Packages' under 'Tools' tab.

```{r load_packages, eval = TRUE}
#
#
# Setup
#
#
library(ggplot2)
library(httk)
library(grid)
library(scales)
library(viridis)
library(ggpubr)
library(readxl)
library(knitr)

LOC.WD <- "c:/users/jwambaug/git/httk-dev/work/HTTKPFAS/"
PFAS.MODEL.USED <- "sumclearancespfas"
NONPFAS.MODEL.USED <- "sumclearances"
```

Set the sizes of the labels and points in the figure (these settings impact the PNG differently than the figure within RStudio):
```{r fig_properties, eval = TRUE}
FONT.SIZE <- 24
POINT.SIZE <- 4
```

### Function to format scientific notation 
From https://stackoverflow.com/questions/10762287/how-can-i-format-axis-labels-with-exponents-with-ggplot2-and-scales
```{R scientific.notation, eval = execute.vignette}
scientific_10 <- function(x) {                                  
  out <- gsub("1e", "10^", scientific_format()(x))              
  out <- gsub("\\+","",out)                                     
  out <- gsub("10\\^01","10",out)                               
  out <- parse(text=gsub("10\\^00","1",out))                    
}  
```

### Function to calculate root mean squared log10 error
```{r RMSLE_function, eval = TRUE}
RMSLE <- function(tab,colx,coly) 
{ 
    return(signif(mean(log10(tab[,colx]/tab[,coly])^2,na.rm=TRUE)^(1/2),3))
}
```

# Analysis of Crizer et al. (2023) $Cl_{int}$ data

## Methods

### Evaluate Ratio Blood:Plasma Predictions
```{r rb2peval, eval=TRUE}
rb2p.table <- subset(chem.physical_and_invitro.data,
                     Human.Rblood2plasma.Reference %in% "Poothong 2017")
rb2p.table <- rb2p.table[,colnames(rb2p.table) %in%
                                     c("Compound","DTXSID",
                                       "Human.Rblood2plasma"
                                     )]
for (this.chem in rb2p.table$DTXSID)
{
  if (this.chem %in% get_2023pfasinfo(info="dtxsid", suppress.messages=TRUE,
                                   model=PFAS.MODEL.USED))
  {
    rb2p.table[rb2p.table$DTXSID==this.chem,"HTTK.Rb2p"] <- 
      calc_rblood2plasma(dtxsid=this.chem,
                         class.exclude=FALSE)
  }
  p <- try(parameterize_pfas1comp(dtxsid=this.chem))
  if(!is(p, "try-error")) 
    rb2p.table[rb2p.table$DTXSID==this.chem,"ML.Rb2p"] <- 
    p$Rblood2plasma
}
kable(rb2p.table)
```

### Evaluate Membrane Affinity Predictions
```{r MAeval, eval=TRUE}
logma.table <- subset(chem.physical_and_invitro.data,
                     logMA.Reference %in% "Droge 2019")
logma.table <- logma.table[,colnames(logma.table) %in%
                                     c("Compound","DTXSID",
                                       "logMA"
                                     )]
for (this.chem in logma.table$DTXSID)
{
  logma.table[logma.table$DTXSID==this.chem,"HTTK.logMA"] <- 
      signif(log10(calc_ma(dtxsid=this.chem,
                           pfas.calibration=FALSE,
                           suppress.messages=TRUE)),3)
}
kable(logma.table)
```

```{r fit_logma, eval = TRUE}
fit.logma <- lm(logMA~HTTK.logMA,data=logma.table)
summary(fit.logma)

logma.table$Fit.logMA <- signif(fit.logma$coefficients[1] +
  fit.logma$coefficients[2]*logma.table$HTTK.logMA,3)
```

```{r fig_logma, eval = TRUE}
fig.logma  <- ggplot(data=logma.table) +
  geom_point(aes(
    x=HTTK.logMA,
    y=logMA
    ),
    size=POINT.SIZE) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
  geom_line(aes(x=HTTK.logMA,
    y=Fit.logMA), color="red")+
  ylab("Droge (2019) logMA") + 
  xlab("HTTK log Membrane Affinity") +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))

print(fig.logma)
```


## New In Vitro TK Parameters

```{r plot_clint_fup}
new.pfas <- get_2023pfasinfo(info="dtxsid",
                                      model=PFAS.MODEL.USED,
                                      suppress.messages=TRUE)
# What chems does sumclearances model work for overall?
good.chems <- get_cheminfo(info="dtxsid", model = PFAS.MODEL.USED)
cat(paste("were incorporated from the scientific literature into R package \\“httk\\” for",
          sum(new.pfas %in% good.chems),
          "PFAS\n"))
cat(paste("Out of",
          length(new.pfas),
          "PFAS with new in vitro TK data reported,",
          sum(new.pfas %in% good.chems),
          "had the Clint, fup, and physicochemical",
          "property predictions needed to make HTTK predictions.\n"))

  

clint.fup <- subset(chem.physical_and_invitro.data,
                    !is.na(Human.Funbound.plasma) &
                    !is.na(Human.Clint) &
                    (regexpr("PFAS",Chemical.Class)==-1 |
                    DTXSID %in% new.pfas))
clint.fup$Human.Clint <- sapply(clint.fup$Human.Clint, 
                                function(x) as.numeric(strsplit(x,",")[[1]][1]))
clint.fup$Human.Funbound.plasma <- sapply(clint.fup$Human.Funbound.plasma, 
                                function(x) as.numeric(strsplit(x,",")[[1]][1]))
clint.fup[clint.fup$Chemical.Class=="","Chemical.Class"] <- "Non-PFAS"
clint.fup[clint.fup$Chemical.Class=="PFAS","Chemical.Class"] <- "Newly Measured PFAS"
clint.fup.pfas <- subset(clint.fup,Chemical.Class!="Non-PFAS")
clint.fup.pfas <- merge(clint.fup.pfas,subset(httk::dawson2023,Species=="Human" &
                              Sex=="Female" &
                              DosingAdj=="Oral")[,c("DTXSID","ClassPredFull")],
      by="DTXSID",all.x=TRUE)
colnames(clint.fup.pfas)[colnames(clint.fup.pfas)=="ClassPredFull"] <- "ML.Class"
save(clint.fup, 
     clint.fup.pfas,
     file=paste0(LOC.WD,"Clint-Fup-Values.RData"))
cat(paste("(fraction unbound in plasma) in",
          sum(!is.na(clint.fup.pfas$Human.Funbound.plasma)),
          "PFAS\n"))
cat(paste("(intrinsic hepatocyte clearance) in",
          sum(!is.na(clint.fup.pfas$Human.Clint)),
          "PFAS\n"))
```

```{r clint.fup.figures}
load(paste0(LOC.WD,"Clint-Fup-Values.RData"))
fig.clint.fup  <- ggplot(data=clint.fup) +
  geom_point(aes(
    x=Human.Funbound.plasma,
    y=Human.Clint,
    colour=Chemical.Class,
    shape=Chemical.Class),
    size=POINT.SIZE) +
  ylab(expression(italic("In Vitro")~"Measured Cl"[int]~"("*mu*"L/min/10"^'6'~"hepatocytes)")) +
  xlab(expression(italic("In Vitro")~"Measured f"[up])) +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE)) +
        scale_color_viridis(discrete=2)

print(fig.clint.fup)

fig.fup.hist <- ggplot(data=clint.fup) +
  geom_histogram(aes(
    x=Human.Funbound.plasma,
    colour=Chemical.Class,
    fill=Chemical.Class)) +
  xlab(expression(italic("In Vitro")~"Measured f"[up])) +
  scale_x_log10(label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))+
        scale_fill_viridis(discrete=2)+
        scale_color_viridis(discrete=2)

print(fig.fup.hist)

fig.clint.hist <- ggplot(data=clint.fup) +
  geom_histogram(aes(
    x=Human.Clint+10^-5,
    colour=Chemical.Class,
    fill=Chemical.Class)) +
  xlab(expression(italic("In Vitro")~"Measured Cl"[int]~"("*mu*"L/min/10"^'6'~"hepatocytes)")) +
  scale_x_log10(label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))+
        scale_fill_viridis(discrete=2)+
        scale_color_viridis(discrete=2)+
  coord_flip()

print(fig.clint.hist)

fig.henry.hist <- ggplot(data=clint.fup) +
  geom_histogram(aes(
    x=10^logHenry,
    colour=Chemical.Class,
    fill=Chemical.Class)) +
  xlab(expression("Henry's Law Constant (Atm.-m"^"3"*"/mole)")) +
  scale_x_log10(label=scientific_10)+
  geom_vline(xintercept =10^-4.5,linetype="dashed",color="blue")+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))+
        scale_fill_viridis(discrete=2)+
        scale_color_viridis(discrete=2)

print(fig.henry.hist)
```

```{r multipanelnewclintfupdata}
ggarrange(fig.clint.fup, fig.clint.hist, fig.fup.hist, fig.henry.hist, 
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2,
          common.legend = TRUE, legend = "bottom")
tiff(file = paste0(LOC.WD,"Fig_NewClintFup.tiff"),
         height = 5*200, width = 5*200, compression = "lzw")
ggarrange(fig.clint.fup, fig.clint.hist, fig.fup.hist, fig.henry.hist, 
          labels = c("A", "B", "C", "D"),
          ncol = 2, nrow = 2,
          common.legend = TRUE, legend = "bottom")
dev.off()
```
## Examine Kbloodair
```{r calc_Kbloodair, eval=TRUE}
pfas.data <- subset(clint.fup,Chemical.Class=="Newly Measured PFAS")
pfas.data[,"logKblood2air"] <- apply(pfas.data,
                                     1,
                                     function(x) try(log10(calc_kair(
                                       dtxsid=x["DTXSID"],
                                       suppress.messages=TRUE)$Kblood2air)))
```

```{r kblood2air_results, eval=TRUE}
good.Kblood2air <- pfas.data$logKblood2air
good.logHenry <- pfas.data$logHenry
good.logHenry <- good.logHenry[!is.na(good.Kblood2air)]
good.Kblood2air <- good.Kblood2air[!is.na(good.Kblood2air)]
cat(paste("Out of",
          length(good.logHenry),
          "PFAS for which the Henry's Law Constant could be predicted,",
          signif(sum(good.logHenry >
                       -4.5)/length(good.logHenry)*100,2),
          "% were more volatile than acetone.\n"))
most.volatile <- pfas.data[pfas.data$logKblood2air %in% min(good.Kblood2air), 
                           "Compound"]
cat(paste("The most volatile chemical was",most.volatile,
          "with a predicted logKblood2air of", signif(min(good.Kblood2air),3),
          ".\n"))
```

# HTTK Interspecies Clearance Differences

## Remove Rat-specific HTTK data:
```{r removeoldhttk}
chem.physical_and_invitro.data$Rat.Clint <- NA
chem.physical_and_invitro.data$Rat.Funbound.plasma <- NA
```

## Add bioavailability predictions
```{r load_bioavailability_preds}
load_honda2023()
```

## Compare default httk against in vivo clearance data:
```{r invivoeval, eval=TRUE}
cl.table <- httk::pfas.clearance
human.ML.table <- subset(httk::dawson2023,
                         Species=="Human" & 
                         Sex=="Female" &
                         DosingAdj=="Oral")

for (this.chem in unique(cl.table$DTXSID))
  if (this.chem %in% get_2023pfasinfo(info="dtxsid",
                                      model=PFAS.MODEL.USED,
                                      suppress.messages=TRUE))
  {
    chem.subset <- subset(cl.table, DTXSID==this.chem)
    for (this.species in unique(cl.table$Species))
    {
      # Css in units of mg/L for 1 mg/kg/day
      httk.normal <- calc_total_clearance(dtxsid=this.chem,
                               species=this.species,
                               model=NONPFAS.MODEL.USED,
                               default.to.human=TRUE,
                               class.exclude=FALSE,
                               physchem.exclude=FALSE,
                               suppress.messages=TRUE)
      httk.pfas <- calc_total_clearance(dtxsid=this.chem,
                               species=this.species,
                               model=PFAS.MODEL.USED,
                               default.to.human=TRUE,
                               class.exclude=FALSE,
                               physchem.exclude=FALSE,
                               suppress.messages=TRUE)
      if (any(cl.table$DTXSID==this.chem &
               cl.table$Species==this.species))
      {
        cl.table[cl.table$DTXSID==this.chem &
               cl.table$Species==this.species,
               "HTTK.Cltot.Lphpkgbw"] <- httk.normal
        cl.table[cl.table$DTXSID==this.chem &
               cl.table$Species==this.species,
               "HTTKPFAS.Cltot.Lphpkgbw"] <- httk.pfas
      } else {
        new.row <- cl.table[1,]
        new.row[] <- NA
        new.row[,"DTXSID"] <- this.chem
        new.row[,"Species"] <- this.species
        new.row[,"Sex"] <- "Female"
        new.row[,"HTTK.Cltot.Lphpkgbw"] <- httk.normal
        new.row[,"HTTKPFAS.Cltot.Lphpkgbw"] <- httk.pfas
        cl.table <- rbind(cl.table,new.row)
        new.row[,"Sex"] <- "Male"
        cl.table <- rbind(cl.table,new.row)
      }
    }
  }
```

## Add in Dawson 2023 predictions for later
```{r invivoeval2, eval=TRUE}
for (this.chem in unique(cl.table$DTXSID))
  {
    chem.subset <- subset(cl.table, DTXSID==this.chem)
    for (this.species in unique(cl.table$Species))
    {
      p <- try(parameterize_pfas1comp(dtxsid=this.chem,
                                  species=this.species,
                                  suppress.messages=TRUE))
      # Clearance in units of L/h/kgBW
      if(!is(p, "try-error")) 
        cl.table[cl.table$DTXSID==this.chem &
               cl.table$Species==this.species,
               "MachineLearning.Cltot.Lphpkgbw"] <- p$kelim*p$Vd
    }
  }
```

### Calculate human:species clearance ratios
```{r invivo_interspecies_eval_scaling, eval=TRUE}
cl.table <- as.data.frame(cl.table)
for (this.sex in c("Female","Male"))
{
  human.data <- subset(cl.table, Species=="Human" &
                            Sex==this.sex)
  for (this.species in unique(cl.table$Species))
  if (this.species != "Human")
  {
    species.data <- subset(cl.table, Species==this.species &
                                     Sex==this.sex)
    for (this.chem in union(human.data$DTXSID, species.data$DTXSID))
    {
      cl.table.index <- cl.table$Sex==this.sex &
               cl.table$Species==this.species &
               cl.table$DTXSID==this.chem
      human.table.index <- human.data$Sex==this.sex &
               human.data$DTXSID==this.chem
      species.table.index <- species.data$Sex==this.sex &
               species.data$DTXSID==this.chem
      if (any(human.table.index) & any(species.table.index))
      {
        cl.table[cl.table.index, "TK.Ratio.InVivo"] <- 
        signif(species.data[species.table.index, "ClLphpkgbw"] /
        human.data[human.table.index, "ClLphpkgbw"], 4)
        
        cl.table[cl.table.index, "TK.Ratio.HTTK"] <- 
        signif(species.data[species.table.index, "HTTK.Cltot.Lphpkgbw"] /
        human.data[human.table.index, "HTTK.Cltot.Lphpkgbw"], 4) 
        
        cl.table[cl.table.index, "TK.Ratio.ML"] <- 
        signif(species.data[species.table.index, "MachineLearning.Cltot.Lphpkgbw"] /
        human.data[human.table.index, "MachineLearning.Cltot.Lphpkgbw"], 4)
        
        cl.table[cl.table.index, "TK.Ratio.HTTKPFAS"] <- 
        signif(species.data[species.table.index, "HTTKPFAS.Cltot.Lphpkgbw"] /
        human.data[human.table.index, "HTTKPFAS.Cltot.Lphpkgbw"], 4) 
      }
    }
  }
}  
```

## Add non-PFAS clearances from literature
```{r cl_lit_Data}
huh.data <- as.data.frame(read_excel(paste0(
  LOC.WD,"Huh-2011-HumanCL.xlsx")))
# Huh Data is in ml/min/kg:
Human.CL.mLpminpkg <- sapply(huh.data[,4],function(x) signif(as.numeric(strsplit(x," ")[[1]][1]),3))
huh.data$Human.CL.Lphpkg <- signif(Human.CL.mLpminpkg/1000*60, 3)
huh.data$Rat.CL.Lphpkg <- signif(
  as.numeric(huh.data$a)*(0.25)^as.numeric(huh.data$b)*60/1000, 3)

httk.chem.list <- get_cheminfo(model=PFAS.MODEL.USED, 
                               info="DTXSID",
                               suppress.messages = TRUE)
for (this.chem in huh.data[,"DTXSID"])
  if (!is.na(this.chem))
  {
  this.row <- which(huh.data[,"DTXSID"]==this.chem)
  new.row <- cl.table[1,]
  new.row[1,] <- NA
  new.row[1, "DTXSID"] <- huh.data[this.row, "DTXSID"]
  new.row[1, "Species"] <- "Rat"
  new.row[1, "Sex"] <- "Female"
  new.row[1, "ClLphpkgbw"] <- huh.data[this.row, "Human.CL.Lphpkg"]

  new.row[1, "TK.Ratio.InVivo"] <- 
    huh.data[this.row, "Rat.CL.Lphpkg"] /
    huh.data[this.row, "Human.CL.Lphpkg"]
  if (this.chem %in% httk.chem.list)
    if (!(this.chem %in% new.pfas) |
        this.chem %in% dawson2023$DTXSID)
  {
    # Calculate human clearance:
    HTTK.Human.Cltot.Lphpkgbw <-
      calc_total_clearance(dtxsid = this.chem,
                           model=NONPFAS.MODEL.USED,
                           species="Human",
                           suppress.messages=TRUE)
    # Add human row (with NA ratio):
    new.row[1, "HTTK.Cltot.Lphpkgbw"] <- HTTK.Human.Cltot.Lphpkgbw 
    cl.table <- rbind(cl.table,new.row)
    # Make a row for rat:
    new.row[1, "ClLphpkgbw"] <- huh.data[this.row, "Rat.CL.Lphpkg"]
    HTTK.Rat.Cltot.Lphpkgbw <- 
      calc_total_clearance(dtxsid = this.chem,
                           model=NONPFAS.MODEL.USED,
                           species="Rat",
                           default.to.human=TRUE,
                           suppress.messages=TRUE)
    new.row[1, "HTTK.Cltot.Lphpkgbw"] <- HTTK.Rat.Cltot.Lphpkgbw
    new.row[1, "TK.Ratio.HTTK"] <- signif(HTTK.Rat.Cltot.Lphpkgbw /
                                          HTTK.Human.Cltot.Lphpkgbw, 4)
    # Add rat row:
    cl.table <- rbind(cl.table,new.row)
  }
  }
```

```{r annotate_chemical_class, eval=TRUE}
cl.table$Class <- "PFAS"
cl.table[!(cl.table$DTXSID %in% new.pfas), "Class"] <- "Non-PFAS"
cl.table$Class <- factor(cl.table$Class, levels=c("PFAS","Non-PFAS"))
cat(paste("Also in Figure 2 Panel B are clearances based upon",
  sum(cl.table$Class!="PFAS")/2,
  "non-PFAS chemicals\n"))
cl.table <- merge(cl.table,subset(httk::dawson2023,DosingAdj=="Oral")[,c("DTXSID","Species","Sex","ClassPredFull")],by=c("DTXSID","Species","Sex"))
colnames(cl.table)[colnames(cl.table)=="ClassPredFull"] <- "ML.Class"
save(cl.table, file=paste0(LOC.WD,"CL-Table.RData"))
write.table(cl.table, file=paste0(LOC.WD,"CL-Table.txt"),sep="\t",row.names=FALSE,quote = FALSE)
```

```{r fit_non_pfas_trend, eval=TRUE}
load(paste0(LOC.WD,"CL-Table.RData"))
nonpfasratiotrend <- lm(log10(TK.Ratio.InVivo)~log10(TK.Ratio.HTTK),
                        data=subset(cl.table, Species=="Rat" & Class!="PFAS"))
summary(nonpfasratiotrend)
```
```{r fig_clhttk, eval = TRUE}
load(paste0(LOC.WD,"CL-Table.RData"))
fig.clhttk.invivo  <- ggplot(data=subset(cl.table, Class=="PFAS")) +
  geom_point(aes(
    x=HTTK.Cltot.Lphpkgbw,
    y=ClLphpkgbw,
    color=Species,
    shape=Sex),
    size=POINT.SIZE) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
#  geom_line(aes(x=logP,y=logPfit1),color="black") + 
#  geom_line(aes(x=logP,y=logPfit2),color="blue",linetype = "dashed") +
  ylab("PFAS In Vivo Clearance (L/h/kg bw)") + 
  xlab("HTTK Clearance (L/h/kg bw)") +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
   scale_color_viridis(discrete=4)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))+
  annotate("text", size=FONT.SIZE/3, x = 10^-4, y = 10^-1, 
           label = paste("RMSLE:",
                         RMSLE(subset(cl.table, Class=="PFAS"),
                               "HTTK.Cltot.Lphpkgbw","ClLphpkgbw")))+
  theme(legend.title = element_text(size = FONT.SIZE/3), 
               legend.text = element_text(size = FONT.SIZE/2))

print(fig.clhttk.invivo)

```

```{r fig_ratio_httk, eval = TRUE}
load(paste0(LOC.WD,"CL-Table.RData"))
fig.ratio.httk  <- ggplot(data=subset(cl.table, Species=="Rat")) +
  geom_point(aes(
    x=TK.Ratio.HTTK,
    y=TK.Ratio.InVivo,
    color=Class,
    shape=Class),
    size=POINT.SIZE) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
#  geom_line(aes(x=logP,y=logPfit1),color="black") + 
#  geom_line(aes(x=logP,y=logPfit2),color="blue",linetype = "dashed") +
  ylab("In Vivo Derived Human:Rat Clearance Ratio") + 
  xlab("HTTK Predicted Clearance Ratio") +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE)) +
  scale_color_viridis(discrete=2)

print(fig.ratio.httk)
#  ggsave(paste0(LOC.WD,
#    "Fig_rat_cl_ratio.tiff"),
#         height =11, width =12, dpi = 300)
```








```{r multipanelclearancefigureforpaper}
ggarrange(fig.clhttk.invivo, fig.ratio.httk,
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          common.legend = FALSE, legend = "bottom")
tiff(file = paste0(LOC.WD,"Fig_PFASRatioEval.tiff"),
         height = 2.5*200, width = 5*200, compression = "lzw")
ggarrange(fig.clhttk.invivo, fig.ratio.httk, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          hjust=-2.5,
          common.legend = FALSE, legend = "top",
          font.label=list(size=20))
dev.off()
```

```{r summary_table}
ratio.table <- data.frame()
ratio.table["Non-PFAS In Vivo Clearance Ratio","Median"] <- 
  median(subset(cl.table, 
                Species=="Rat" &
                  Class != "PFAS")$TK.Ratio.InVivo,na.rm=TRUE)
ratio.table["Non-PFAS In Vivo Clearance Ratio","Min"] <- 
  min(subset(cl.table, 
                Species=="Rat" &
                  Class != "PFAS")$TK.Ratio.InVivo,na.rm=TRUE)
ratio.table["Non-PFAS In Vivo Clearance Ratio","Max"] <- 
  max(subset(cl.table, 
                Species=="Rat" &
                  Class != "PFAS")$TK.Ratio.InVivo,na.rm=TRUE)
ratio.table["PFAS In Vivo Clearance Ratio","Median"] <- 
  median(subset(cl.table, 
                Species=="Rat" &
                  Class == "PFAS")$TK.Ratio.InVivo,na.rm=TRUE)
ratio.table["PFAS In Vivo Clearance Ratio","Min"] <- 
  min(subset(cl.table, 
                Species=="Rat" &
                  Class == "PFAS")$TK.Ratio.InVivo,na.rm=TRUE)
ratio.table["PFAS In Vivo Clearance Ratio","Max"] <- 
  max(subset(cl.table, 
                Species=="Rat" &
                  Class == "PFAS")$TK.Ratio.InVivo,na.rm=TRUE)
ratio.table["PFAS Default HTTK Clearance Ratio","Median"] <- 
  median(subset(cl.table, 
                Species=="Rat" &
                  Class == "PFAS")$TK.Ratio.HTTK)
ratio.table["PFAS Default HTTK Clearance Ratio","Min"] <- 
  min(subset(cl.table, 
                Species=="Rat" &
                  Class == "PFAS")$TK.Ratio.HTTK)
ratio.table["PFAS Default HTTK Clearance Ratio","Max"] <- 
  max(subset(cl.table, 
                Species=="Rat" &
                  Class == "PFAS")$TK.Ratio.HTTK)
ratio.table <- signif(ratio.table, 2)
write.table(ratio.table, file=paste0(LOC.WD,"Ratio-Table.txt"),
            sep="\t")
kable(ratio.table)
```

## Evaluating Machine Learning Model for PFAS TK

```{r make_ml_clint_table}
load(paste0(LOC.WD,"Clint-Fup-Values.RData")
clint.fup.pfas[is.na(clint.fup.pfas$Human.Clint.pValue),"Human.Clint.pValue"] <- 0
ml.clint.table <- data.frame()
ml.clint.table["Class 1", "Description"] <- "Rapidly Cleared"
ml.clint.table["Class 2", "Description"] <- "Moderate"
ml.clint.table["Class 3", "Description"] <- "Slow"
ml.clint.table["Class 4", "Description"] <- "Very Slow"
ml.clint.table["Class 1", "Half-Life"] <- "< 12 h"
ml.clint.table["Class 2", "Half-Life"] <- "< 1 week"
ml.clint.table["Class 3", "Half-Life"] <- "< 2 months"
ml.clint.table["Class 4", "Half-Life"] <- "> 2 months"
ml.clint.table["Class 1", "Number Crizer (2024) Chemicals"] <- 
  dim(subset(clint.fup.pfas,
             ML.Class==1))[1]
ml.clint.table["Class 2", "Number Crizer (2024) Chemicals"] <- 
  dim(subset(clint.fup.pfas,
             ML.Class==2))[1]
ml.clint.table["Class 3", "Number Crizer (2024) Chemicals"] <- 
  dim(subset(clint.fup.pfas,
             ML.Class==3))[1]
ml.clint.table["Class 4", "Number Crizer (2024) Chemicals"] <- 
  dim(subset(clint.fup.pfas,
             ML.Class==4))[1]
ml.clint.table["Class 1", "Number Non-Zero Clint"] <-
  dim(subset(clint.fup.pfas,
             ML.Class==1 &
             Human.Clint > 0 &
             Human.Clint.pValue < 0.05))[1]
ml.clint.table["Class 2", "Number Non-Zero Clint"] <-
  dim(subset(clint.fup.pfas,
             ML.Class==2 &
             Human.Clint > 0 &
             Human.Clint.pValue < 0.05))[1]
ml.clint.table["Class 3", "Number Non-Zero Clint"] <-
  dim(subset(clint.fup.pfas,
             ML.Class==3 &
             Human.Clint > 0 &
             Human.Clint.pValue < 0.05))[1]
ml.clint.table["Class 4", "Number Non-Zero Clint"] <-
  dim(subset(clint.fup.pfas,
             ML.Class==4 &
             Human.Clint > 0 &
             Human.Clint.pValue < 0.05))[1]
write.table(ml.clint.table, file=paste0(LOC.WD,
                                        "ML-Clint-table.txt"), sep="\t")
kable(ml.clint.table)
```

```{r rat_human_scaling}
half.lives <- NULL
for (this.chem in new.pfas)
{
  new.row <- data.frame(DTXSID=this.chem,
                        Human.HL=try(calc_half_life(
                          dtxsid=this.chem,
                          species="human",
                          model="sumclearances",
                          class.exclude=FALSE,
                          physchem.exclude=FALSE,
                          suppress.messages=TRUE)),
                        Rat.HL=try(calc_half_life(
                          dtxsid=this.chem,
                          species="rat",
                          model="sumclearances",
                          default.to.human=TRUE,
                          class.exclude=FALSE,
                          physchem.exclude=FALSE,
                          suppress.messages=TRUE)))
  new.row$Type <- "HTTK"
  half.lives <- rbind(half.lives, new.row)
  param.human <- try(parameterize_pfas1comp(dtxsid=this.chem,
                                  species="human",
                                  suppress.messages=TRUE))
  param.rat <- try(parameterize_pfas1comp(dtxsid=this.chem,
                                  species="rat",
                                  suppress.messages=TRUE))
  if ("kelim" %in% names(param.human) &
      "kelim" %in% names(param.rat))
    new.row <- data.frame(DTXSID=this.chem,
                        Human.HL=log(2)/param.human$kelim,
                        Rat.HL=log(2)/param.rat$kelim)
  new.row$Type <- "ML"
  half.lives <- rbind(half.lives, new.row)
}
half.lives$Ratio <- half.lives$Human.HL / half.lives$Rat.HL
save(half.lives,file=paste0(LOC.WD,"All-Half-Lives.RData"))
```

```{r half_life_fig}
load(paste0(LOC.WD,"All-Half-Lives.RData"))
fig.species.scaling <- ggplot(data=half.lives) +
  geom_histogram(aes(
    x=Ratio,
    colour=Type,
    fill=Type),
    bins=10) +
  xlab("Human:Rat PK Half-Life Ratio") +
  scale_x_log10(label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))+
        scale_fill_viridis(discrete=2)+
        scale_color_viridis(discrete=2)


print(fig.species.scaling)
```

```{r wallis_2023_halflives}
Dawson.known.human.HLs <- c(
"DTXSID1037303",
"DTXSID1037303",
"DTXSID3031860",
"DTXSID3031860",
"DTXSID3031862",
"DTXSID3031862",
"DTXSID3031864",
"DTXSID3031864",
"DTXSID4059916",
"DTXSID4059916",
"DTXSID5030030",
"DTXSID5030030",
"DTXSID7040150",
"DTXSID7040150",
"DTXSID70880215",
"DTXSID70880215",
"DTXSID8031863",
"DTXSID8031863",
"DTXSID8031865",
"DTXSID8031865",
"DTXSID80892506",
"DTXSID80892506"
)

wallis <- as.data.frame(read_excel(paste0(LOC.WD,"Wallis-2023-PFASHL.xlsx")))
wallis$HL.Low <- sapply(wallis[,4],function(x) as.numeric(strsplit(x,", ")[[1]][1]))
wallis$HL.High <- sapply(wallis[,4],function(x) as.numeric(strsplit(x,", ")[[1]][2]))

abraham <- as.data.frame(read_excel(paste0(LOC.WD,"Abraham-2024-PFASHL.xlsx")))
abraham$HL.Low <- sapply(abraham[,"Halflife.95.d"],function(x) as.numeric(strsplit(x,"–")[[1]][1]))
abraham$HL.High <- sapply(abraham[,"Halflife.95.d"],function(x) as.numeric(strsplit(x,"–")[[1]][2]))
abraham$HL.High[is.na(abraham$HL.High)] <- abraham$HL.Low*100

wallis$Reference <- "Wallis 2023"
abraham$Reference <- "Abraham 2024"

human.invivo.hl <- wallis[,c("PREFERRED_NAME","DTXSID","CASRN","Reference","HL.Low","HL.High")]
human.invivo.hl <- rbind(human.invivo.hl,
                         abraham[,c("PREFERRED_NAME","DTXSID","CASRN","Reference","HL.Low","HL.High")])
human.invivo.hl$Status <- "Novel"
human.invivo.hl[human.invivo.hl$DTXSID %in% Dawson.known.human.HLs, "Status"] <- "Known"
                    
for (this.chem in human.invivo.hl$DTXSID)
{
  param.human <- try(parameterize_pfas1comp(dtxsid=this.chem,
                                  species="human",
                                  suppress.messages=TRUE,
                                  restrict.doa="none"))
  if ("kelim" %in% names(param.human))
    human.invivo.hl[human.invivo.hl$DTXSID == this.chem, "HL.ML"] <- log(2)/param.human$kelim/24
}
```

```{r results_from_new_model_run}
human.invivo.hl[human.invivo.hl$DTXSID=="DTXSID90723993","HL.ML"] <- 29000/24
human.invivo.hl[human.invivo.hl$DTXSID=="DTXSID50723994","HL.ML"] <- 29000/24
human.invivo.hl[human.invivo.hl$DTXSID=="DTXSID20892348","HL.ML"] <- 53/24
for (this.col in c("HL.Low","HL.High","HL.ML"))
{
  temp.col <- human.invivo.hl[, this.col]
  bad.rows <- is.na(temp.col)
  temp.col[bad.rows] <- -1
  new.col <- paste0(this.col,".Class") 
    human.invivo.hl[, new.col] <- 4
  human.invivo.hl[temp.col < 60,
       new.col] <- 3
  human.invivo.hl[temp.col < 7,
       new.col] <- 2
  human.invivo.hl[temp.col< 0.5,
       new.col] <- 1
  human.invivo.hl[bad.rows, new.col] <- NA
}
human.invivo.hl$Agree <- human.invivo.hl$HL.Low.Class ==
                           human.invivo.hl$HL.ML.Class |
                         human.invivo.hl$HL.High.Class ==
                           human.invivo.hl$HL.ML.Class 
human.invivo.hl$Underestimate <- human.invivo.hl$HL.High.Class <
                           human.invivo.hl$HL.ML.Class
```

```{r human_hl_stats}
human.invivo.hl$HL.GeoMean <- exp((log(human.invivo.hl$HL.Low)+log(human.invivo.hl$HL.High))/2)
fit.overall <- lm(log10(HL.GeoMean) ~ log10(HL.ML),
                  data=human.invivo.hl)
fit.known <- lm(log10(HL.GeoMean) ~ log10(HL.ML),
                data=subset(human.invivo.hl,Status=="Known"))
fit.novel <- lm(log10(HL.GeoMean) ~log10(HL.ML), 
                data=subset(human.invivo.hl,Status=="Novel"))
hl.known <- subset(human.invivo.hl,Status=="Known")
RMSLE.known <- (mean((log10(hl.known$HL.GeoMean) - log10(hl.known$HL.ML))^2,na.rm=TRUE))^(1/2)
agree.known <- sum(hl.known$Agree)/dim(hl.known)[1]
under.known <- sum(hl.known$Underestimate)/dim(hl.known)[1]
hl.novel <- subset(human.invivo.hl,Status=="Novel" & !is.na(HL.ML))
RMSLE.novel <- (mean((log10(hl.novel$HL.GeoMean) - log10(hl.novel$HL.ML))^2,na.rm=TRUE))^(1/2)
agree.novel <- sum(hl.novel$Agree)/dim(hl.novel)[1]
agree.count.novel <- sum(hl.novel$Agree)
under.novel <- sum(hl.novel$Underestimate)/dim(hl.known)[1]
under.count.novel <- sum(hl.novel$Underestimate)
num.novel.pfas.hl <- dim(hl.novel)[1]
cat(paste(num.novel.pfas.hl,"novel PFAS human half-lives.\n"))
save(human.invivo.hl,
     hl.known,
     hl.novel,
     file=paste0(LOC.WD,"Human-Invivo-HL.RData"))
```

```{r human_hl_eval_fig}
load(paste0(LOC.WD,"Human-Invivo-HL.RData"))
set.seed(1234)
human.invivo.hl$HL.ML.jitter <- human.invivo.hl$HL.ML*(1.0+runif(dim(human.invivo.hl)[1],-0.5,0.5))

human.invivo.hl.fig <- ggplot(data=human.invivo.hl) +
  geom_segment(aes(
    x=HL.ML.jitter,
    xend=HL.ML.jitter,
    y=HL.Low,
    yend=HL.High,
    colour=Status),
    linewidth=3,
    alpha=0.5) +
  geom_point(aes(
    x=HL.ML.jitter,
    y=HL.GeoMean,
    shape=Reference,
    colour=Status),
    size=2) +
    geom_abline(slope=1,intercept=0,linetype="dashed")+
  xlab("Machine Learning Half-Life Prediction") +
  ylab("In Vivo Measurment") +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
  theme_bw()  +
  annotate("text", x = 5, y = 10^4, size=6, hjust=0,
           label = paste("RMSLE for ",
                         num.novel.pfas.hl, 
                         " novel PFAS: ",
                         signif(RMSLE.novel,2),"\n"#,
                         #agree.count.novel,
                         #" agree\n",
                         #under.count.novel,
                         #" underestimated",
                         ,sep=""))+
  theme( text  = element_text(size=FONT.SIZE))+
        scale_color_viridis(discrete=2)+
    theme(legend.title = element_text(size = FONT.SIZE/3), 
               legend.text = element_text(size = FONT.SIZE/2))

print(human.invivo.hl.fig)

#tiff(file = paste0(LOC.WD,"InVivoComparison.tiff"),
#     width=9,
#     height=7,
#     units="in",
#     res=300)
#print(human.invivo.hl.fig)
#dev.off()

subset(human.invivo.hl, is.na(HL.ML))
```

```{r multipanelMLfigureforpaper}
ggarrange(fig.species.scaling, human.invivo.hl.fig,
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          common.legend = FALSE, legend = "bottom")
tiff(file = paste0(LOC.WD,"Fig_MLModelEval.tiff"),
         height = 2.5*200, width = 5*200, compression = "lzw")
ggarrange(fig.species.scaling, human.invivo.hl.fig,
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          hjust=-2.5,
          common.legend = FALSE, legend = "top",
          font.label=list(size=20))
dev.off()
```

```{r fig_clml, eval = TRUE}
load(paste0(LOC.WD,"CL-Table.RData"))
fig.clml.invivo  <- ggplot(data=subset(cl.table, Class=="PFAS")) +
  geom_jitter(aes(
    x=MachineLearning.Cltot.Lphpkgbw,
    y=ClLphpkgbw,
    color=Species,
    shape=Sex),
    size=POINT.SIZE,
    height=0,width=0.1) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
#  geom_line(aes(x=logP,y=logPfit1),color="black") + 
#  geom_line(aes(x=logP,y=logPfit2),color="blue",linetype = "dashed") +
  ylab("In Vivo Clearance (L/h/kg bw)") + 
  xlab("Machine Learning Clearance (L/h/kg bw)") +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
   scale_color_viridis(discrete=4)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))+
  annotate("text", x = 3*10^-5, y = 10^-1,size=FONT.SIZE/3, 
           label = paste("RMSLE:",
                         RMSLE(cl.table,
                               "MachineLearning.Cltot.Lphpkgbw",
                               "ClLphpkgbw")))

print(fig.clml.invivo)
```

```{r fig_ratio_ml, eval = TRUE}
load(paste0(LOC.WD,"CL-Table.RData"))
fig.ratio.ml  <- ggplot(data=subset(cl.table, Species!="Human" & Class=="PFAS")) +
  geom_point(aes(
    x=TK.Ratio.ML,
    y=TK.Ratio.InVivo,
    color=Species,
    shape=Sex),
    size=POINT.SIZE) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
#  geom_line(aes(x=logP,y=logPfit1),color="black") + 
#  geom_line(aes(x=logP,y=logPfit2),color="blue",linetype = "dashed") +
  ylab("In Vivo Observed Clearance Ratio") + 
  xlab("Machine Learning Predicted Clearance Ratio") +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
   scale_color_viridis(discrete=3)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))

print(fig.ratio.ml)
  ggsave(paste0(LOC.WD,
    "Fig_ml_cl_ratio.tiff"),
         height = 11, width = 12, dpi = 300)
```

## New HTTK Model for PFAS
```{r fig_clhttkpfas, eval = TRUE}
load(paste0(LOC.WD,"CL-Table.RData"))
fig.clhttkpfas.invivo  <- ggplot(data=subset(cl.table, Class=="PFAS")) +
  geom_point(aes(
    x=HTTKPFAS.Cltot.Lphpkgbw,
    y=ClLphpkgbw,
    color=Species,
    shape=Sex),
    size=POINT.SIZE) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
#  geom_line(aes(x=logP,y=logPfit1),color="black") + 
#  geom_line(aes(x=logP,y=logPfit2),color="blue",linetype = "dashed") +
  ylab("PFAS In Vivo Clearance (L/h/kg bw)") + 
  xlab("HTTK-PFAS Clearance (L/h/kg bw)") +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
   scale_color_viridis(discrete=4)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))+
  annotate("text", size=FONT.SIZE/3, x = 10^-6, y = 10^-1, 
           label = paste("RMSLE:",
                         RMSLE(subset(cl.table, Class=="PFAS"),
                               "HTTKPFAS.Cltot.Lphpkgbw","ClLphpkgbw")))+
  theme(legend.title = element_text(size = FONT.SIZE/3), 
               legend.text = element_text(size = FONT.SIZE/2))

print(fig.clhttkpfas.invivo)

```



```{r multipanelclearancenewmodelfigureforpaper}
multiclerancenewmodel <- ggarrange(fig.clml.invivo+ rremove("ylab") , fig.clhttkpfas.invivo+ rremove("ylab") ,
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          common.legend = TRUE, legend = "bottom")
multiclerancenewmodel <- annotate_figure(multiclerancenewmodel , 
                left = textGrob(expression(paste("PFAS ",
                  italic("In Vivo"),
                  "Clearance (L/h/kg bw)")), 
                                rot = 90, vjust = 1, gp = gpar(cex = 1.75)),
                )
print(multiclerancenewmodel)
tiff(file = paste0(LOC.WD,"Fig_PFASClearanceEval.tiff"),
         height = 2.5*200, width = 5*200, compression = "lzw")
print(multiclerancenewmodel)
dev.off()
```

```{r fig_ratio_httkpfas, eval = TRUE}
load(paste0(LOC.WD,"CL-Table.RData"))
fig.ratio.httkpfas  <- ggplot(data=subset(cl.table, Species!="Human" & Class=="PFAS")) +
  geom_point(aes(
    x=TK.Ratio.HTTKPFAS,
    y=TK.Ratio.InVivo,
    color=Species,
    shape=Sex),
    size=POINT.SIZE) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
#  geom_line(aes(x=logP,y=logPfit1),color="black") + 
#  geom_line(aes(x=logP,y=logPfit2),color="blue",linetype = "dashed") +
  ylab(expression(paste(italic("In Vivo"),
    "Observed Clearance Ratio"))) + 
  xlab("HTTK-PFAS Predicted Clearance Ratio") +
  scale_x_log10(label=scientific_10)+
  scale_y_log10(label=scientific_10)+
  scale_color_viridis(discrete=3)+
  annotate("text", size=FONT.SIZE/3, x = 5*10^2, y = 3, 
           label = paste("RMSLE:",
                         RMSLE(subset(cl.table, Species!="Human" & Class=="PFAS"),
                               "TK.Ratio.HTTKPFAS","TK.Ratio.InVivo")))+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))

print(fig.ratio.httkpfas)
  ggsave(paste0(LOC.WD,
    "Fig_httkpfas_cl_ratio.tiff"),
         height = 5*1.5, width = 6*1.5, dpi = 300)
```

```{r MakeSupTableChange}
pfas.httk.table <- get_2023pfasinfo(info="all")
for (this.chem in pfas.httk.table$DTXSID)
{
  this.row <- pfas.httk.table$DTXSID==this.chem
  pfas.httk.table[this.row, "Css.HTTK.Classic"] <- try(calc_analytic_css(
                                                dtxsid=this.chem,
                                                class.exclude=FALSE,
                                                physchem.exclude=FALSE,
                                                suppress.messages=TRUE,
                                                model="3compartmentss"))
  pfas.httk.table[this.row, "Css.HTTK.Exhale"] <- try(calc_analytic_css(
                                                dtxsid=this.chem,
                                                class.exclude=FALSE,
                                                suppress.messages=TRUE,
                                                model="sumclearances"))
  pfas.httk.table[this.row, "Css.HTTK.PFAS"] <- try(calc_analytic_css(
                                                dtxsid=this.chem,
                                                suppress.messages=TRUE,
                                                model="sumclearancespfas"))
  pfas.httk.table[this.row, "Css.ML"] <- try(calc_analytic_css(
                                                dtxsid=this.chem,
                                                suppress.messages=TRUE,
                                                model="pfas1compartment"))
  fractions <- try(calc_clearance_frac(dtxsid=this.chem, 
                                       model="sumclearancespfas",
                                       suppress.messages=TRUE,
                                      fraction.params = c("Clint",
                                                          "Qgfrc",
                                                          "Qalvc")))
  if (!is(fractions, "try-error"))
  {
    pfas.httk.table[this.row, "Frac.Metabolism"] <- fractions[["Clint"]]
    pfas.httk.table[this.row, "Frac.Glomerular"] <- fractions[["Qgfrc"]]
    pfas.httk.table[this.row, "Frac.Exhalation"] <- fractions[["Qalvc"]]
  }
}
pfas.httk.table$Css.HTTK.Classic <- as.numeric(pfas.httk.table$Css.HTTK.Classic)
pfas.httk.table$Css.HTTK.Exhale <- as.numeric(pfas.httk.table$Css.HTTK.Exhale)
pfas.httk.table$Css.HTTK.PFAS <- as.numeric(pfas.httk.table$Css.HTTK.PFAS)
pfas.httk.table$Css.ML <- as.numeric(pfas.httk.table$Css.ML)
pfas.httk.table$Css.Exhale.FoldChange <- 
  signif(pfas.httk.table$Css.HTTK.Exhale/pfas.httk.table$Css.HTTK.Classic, 3)
pfas.httk.table$Css.PFAS.FoldChange <- 
  signif(pfas.httk.table$Css.HTTK.PFAS/pfas.httk.table$Css.HTTK.Classic, 3)
pfas.httk.table <- 
  merge(pfas.httk.table,subset(httk::dawson2023,
                               Species=="Human" &
                               Sex=="Female" &
                               DosingAdj=="Oral")[
  ,c("DTXSID","ClassPredFull")],by="DTXSID",all.x=TRUE)
colnames(pfas.httk.table)[colnames(pfas.httk.table)=="ClassPredFull"] <- "ML.Class"
pfas.httk.table$InVitroMetabolism <- "No"
clint.pvalue <- pfas.httk.table$Human.Clint.pValue
clint.pvalue[is.na(clint.pvalue)] <- 0
pfas.httk.table[pfas.httk.table$Human.Clint > 0 &
                clint.pvalue < 0.05, 
                "InVitroMetabolism"] <- "Yes"


write.table(pfas.httk.table, 
          row.names=FALSE, 
          quote=FALSE,
          sep="\t",
          file=paste(LOC.WD,"new-pfas-css-uM.txt",sep=""))
save(pfas.httk.table,file=paste0(LOC.WD,"PFAS-Css-Table.RData"))
```

```{r cssfigpanela, eval=TRUE}
load(paste0(LOC.WD,"PFAS-Css-Table.RData"))
pfas.httk.table$ML.Class <- as.factor(pfas.httk.table$ML.Class)
fig.css.httk  <- ggplot(data=subset(pfas.httk.table,!is.na(ML.Class))) +
  geom_point(aes(
    x=Css.HTTK.Classic,
    y=Css.HTTK.PFAS,
    shape=InVitroMetabolism,
    color=ML.Class
    ),
    size=POINT.SIZE) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
  geom_abline(slope=1,intercept=1,linetype="dotted", color="blue")+
  geom_abline(slope=1,intercept=-1,linetype="dotted", color="blue") +
  ylab(expression(paste("HTTK-PFAS",
                         ~C[ss],
                         " (",
                         mu,
                         "M)"))) + 
  xlab(expression(paste("Default HTTK",
                         ~C[ss],
                         " (",
                         mu,
                         "M)"))) + 
  scale_x_log10(limits=c(10^-2,10^6),label=scientific_10)+
  scale_y_log10(limits=c(10^-2,10^6),label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE)) +
  scale_color_viridis(discrete=3)+ 
  guides(color=guide_legend(title="ML Halflife Class"),
         shape=guide_legend(title="In Vitro Metabolism"))

print(fig.css.httk)
#  ggsave(paste0(LOC.WD,
#    "Fig_httkpfas_cl_ratio.tiff"),
#         height = 11, width = 12, dpi = 300)
```

```{r cssfigpanelb, eval=TRUE}
load(paste0(LOC.WD,"PFAS-Css-Table.RData"))
pfas.httk.table$ML.Class <- as.factor(pfas.httk.table$ML.Class)
fig.css.ml  <- ggplot(data=subset(pfas.httk.table,!is.na(ML.Class))) +
  geom_point(aes(
    x=Css.ML,
    y=Css.HTTK.PFAS,
    shape=InVitroMetabolism,
    color=ML.Class
    ),
    size=POINT.SIZE) +
  geom_abline(slope=1,intercept=0,linetype="dashed")+
  geom_abline(slope=1,intercept=1,linetype="dotted", color="blue")+
  geom_abline(slope=1,intercept=-1,linetype="dotted", color="blue") +
  ylab(expression(paste("HTTK-PFAS",
                         ~C[ss],
                         " (",
                         mu,
                         "M)"))) + 
  xlab(expression(paste("Machine Learning",
                         ~C[ss],
                         " (",
                         mu,
                         "M)"))) + 
  scale_x_log10(limits=c(10^-2,10^6),label=scientific_10)+
  scale_y_log10(limits=c(10^-2,10^6),label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE)) +
  scale_color_viridis(discrete=3)+ 
  guides(color=guide_legend(title="ML Halflife Class"),
         shape=guide_legend(title="In Vitro Metabolism"))

print(fig.css.ml)
#  ggsave(paste0(LOC.WD,
#    "Fig_httkpfas_cl_ratio.tiff"),
#         height = 11, width = 12, dpi = 300)
```

```{r multipanelcssfigureforpaper}
multipanelcss <- ggarrange(fig.css.httk + rremove("ylab") , 
                                   fig.css.ml + rremove("ylab") ,
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          common.legend = TRUE, legend = "bottom")
multipanelcss <- annotate_figure(multipanelcss, 
                left = textGrob(expression(paste("HTTK-PFAS",
                         ~C[ss],
                         " (",
                         mu,
                         "M)")), 
                rot = 90, 
                vjust = 1, 
                gp = gpar(cex = 1.75)),
                )
print(multipanelcss)
tiff(file = paste0(LOC.WD,"Fig_Css.tiff"),
         height = 2.5*200, width = 5*200, compression = "lzw")
print(multipanelcss)
dev.off()
```

## Other Analyses
```{r firstpass, eval=TRUE}
firstpass.table <- NULL
for (this.chem in get_2023pfasinfo(info="dtxsid", suppress.messages=TRUE,
                                   model=PFAS.MODEL.USED))
{
      params <- parameterize_sumclearances(dtxsid=this.chem,
                                         suppress.messages=TRUE,
                                         class.exclude=FALSE)
      params2 <- parameterize_schmitt(dtxsid=this.chem,
                                      suppress.messages=TRUE,
                                      class.exclude=FALSE)                           
      ion <- calc_ionization(
        pH=7.4,
        pKa_Donor=params2[["pKa_Donor"]],
        pKa_Accept=params2[["pKa_Accept"]])
      fraction_neutral <- ion$fraction_neutral
      if (fraction_neutral < 0.5) params[['Rblood2plasma']] <- 0.5
      else params[['Rblood2plasma']] <- 20
  params[["Clmetabolismc"]] <- calc_hep_clearance(parameters=params,
          hepatic.model='unscaled',
          suppress.messages=TRUE)#L/h/kg body weight
  this.row <- data.frame(DTXSID=this.chem,
                         Fsystemic=calc_hep_bioavailability(
                           parameters=params)
  )
  firstpass.table <- rbind(firstpass.table, this.row)
}
```

```{r firstpass_fig, eval=TRUE}
fig.firstpass  <- ggplot(data=firstpass.table) +
  geom_histogram(aes(
    x=Fsystemic),bins=10) +
  ylab("Number of Chemicals") + 
  xlab("Fraction Systemically Available") +
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))

print(fig.firstpass)
```

```{r fig_clspecies_invivo, eval = TRUE}
fig.clspecies.invivo  <- ggplot(data=cl.table) +
  geom_point(aes(
    x=Species,
    y=ClLphpkgbw,
    color=Species,
    shape=Sex),
    size=POINT.SIZE) +
  geom_line(aes(x=Species,
              y=ClLphpkgbw,
              group = DTXSID)) +
#  geom_line(aes(x=logP,y=logPfit1),color="black") + 
#  geom_line(aes(x=logP,y=logPfit2),color="blue",linetype = "dashed") +
  ylab("In Vivo Clearance (L/h/kg bw)") + 
  xlab("Species") +
  scale_y_log10(label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))

print(fig.clspecies.invivo)
```

```{r fig_clspecies_httk, eval = TRUE}
fig.clspecies.httk  <- ggplot(data=cl.table) +
  geom_point(aes(
    x=Species,
    y=HTTK.Cltot.Lphpkgbw,
    color=Species,
    shape=Sex),
    size=POINT.SIZE) +
#  geom_segment(color="#69b3a2", 
#                  aes(
#                    x=Species,
#                    y=HTTK.Cltot.Lphpkgbw,
#                    xend = c(tail(Species, n = -1), NA), 
#                    yend = c(tail(HTTK.Cltot.Lphpkgbw, n = -1), NA)
#                  ))+
geom_line(aes(x=Species,
              y=HTTK.Cltot.Lphpkgbw,
              group = DTXSID)) +
ylab("HTTK Clearance (L/h/kg bw)") +
  xlab("Species") +
  scale_y_log10(label=scientific_10)+
  theme_bw()  +
  theme( text  = element_text(size=FONT.SIZE))

print(fig.clspecies.httk)
```


```{r clearance_fraction}
FigClearanceFrac1<- ggplot(data=pfas.httk.table)+
   geom_histogram(binwidth = 0.1,alpha=1,fill="Red",aes(Frac.Metabolism))+ 
  xlab(" ") +
  ylab(" ")+
  ggtitle("Hepatic Metabolism")+
  scale_x_continuous(limits=c(-.05,1.05))+
  scale_y_continuous(limits=c(0,100))
FigClearanceFrac2<- ggplot(data=pfas.httk.table)+
   geom_histogram(binwidth = 0.1,alpha=1,fill="Blue",aes(Frac.Glomerular))+
  xlab(" ") +
  ylab("Number of chemicals")+
  ggtitle("Glomerular Filtration")+
  scale_x_continuous(limits=c(-.05,1.05))+
  scale_y_continuous(limits=c(0,100))
FigClearanceFrac3<- ggplot(data=pfas.httk.table)+
   geom_histogram(binwidth = 0.1,alpha=1,fill="Green",aes(Frac.Exhalation))+
  xlab("Fraction of Total Clearance") +
  ylab(" ")+
  ggtitle("Exhalation")+
  scale_x_continuous(limits=c(-.05,1.05))+
  scale_y_continuous(limits=c(0,100))
ggarrange(FigClearanceFrac1, FigClearanceFrac2, FigClearanceFrac3,
          ncol = 1, nrow = 3,
          common.legend = TRUE, legend = "right")
tiff(file = paste0(LOC.WD,"Fig_FracClearance.tiff"),
         height = 3*100, width = 5*100, compression = "lzw")
ggarrange(FigClearanceFrac1, FigClearanceFrac2, FigClearanceFrac3,
          ncol = 1, nrow = 3,
          common.legend = TRUE, legend = "right")
dev.off()
```

```{r two_panel_Css_histogram}
Fig.Cssa <- ggplot(data=pfas.httk.table)+
  geom_histogram(binwidth = 0.01,aes(Css.Exhale.FoldChange))+
  xlab("Css Fold Change Due to Exhalation")
Fig.Cssb <- ggplot(data=pfas.httk.table)+
  geom_histogram(binwidth = 10,aes(Css.PFAS.FoldChange))+
  xlab("Css Fold Change Due to PFAS Resorption")
ggarrange(Fig.Cssa, Fig.Cssb, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          common.legend = TRUE, legend = "bottom")
tiff(file = paste0(LOC.WD,"Fig_CssFoldChange.tiff"),
         height = 3*100, width = 5*100, compression = "lzw")
ggarrange(Fig.Cssa, Fig.Cssb, 
          labels = c("A", "B"),
          ncol = 2, nrow = 1,
          common.legend = TRUE, legend = "top")
dev.off()
```

