---
title: "Ring et al. 2017 Global sensitivity analysis plotting"
author: "Caroline Ring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Global sensitivity analysis plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = '#>')
```

This vignette includes the code to create the global sensitivity analysis plots included in the HTTK-Pop paper.
```{r load_packages, eval= FALSE}
library(data.table)
library(httk)
library(ggplot2) 
library(scales)
```
First, read in the global sensitivity analysis data. Do this for all four combinations of the `poormetab` and `fup.censor` conditions.
```{r read_sens_data, eval=FALSE}
model <- '3compartmentss'
css.method <- 'analytic'

#All combinations of poormetab and fup.censor
pmfc <- expand.grid(poormetab=c(TRUE, FALSE),
                    fup.censor=c(TRUE, FALSE),
                    fuptofub=TRUE)
#Initialize list of data.tables
dt.list <- vector(mode="list", length=nrow(pmfc))

for (i in 1:nrow(pmfc)){
  dt.list[[i]] <- readRDS(paste0('data/',
                   'sens_glenisaacs_nhanes_', 
                   'allchems_',
                   'allgroups_',
                   model,
                   '_',
                   css.method,
                   '_fup_censor_',
                   pmfc[i, 'fup.censor'],
                   '_poormetab_',
                   pmfc[i, 'poormetab'],
                   ifelse(pmfc[i, 'fuptofub'], "_FuptoFub", ""),
                   '.Rdata'))
  dt.list[[i]][, poormetab:=pmfc[i, 'poormetab']]
  dt.list[[i]][, fup.censor:=pmfc[i, 'fup.censor']]
  dt.list[[i]][, fuptofub:=pmfc[i, 'fuptofub']]
}

dat <- rbindlist(dt.list)
rm(dt.list)
setnames(dat, c('chemcas', 'rn'),
         c('CAS', 'param'))
```
Next, add measured values of `Funbound.plasma` and `CLint` to the data table. Get these measured values from `httk`.
```{r get_measured_vals, eval=FALSE}
cheminfo.dt <- as.data.table(httk::get_cheminfo(model='3compartmentss',
                                          info=c('CAS','Compound',
                                                 'Funbound.plasma'),
                                          exclude.fup.zero=FALSE))
setnames(cheminfo.dt,
         'Human.Funbound.plasma',
         'Funbound.plasma'
)

cheminfo.dt[, Clint:=sapply(CAS, 
                            function(x) parameterize_steadystate(chem.cas=x,
                                                                 species='Human')$Clint)]

dat <- merge(dat,
                  cheminfo.dt[, .(CAS,
                                  Compound,
                                  Funbound.plasma,
                                  Clint)],
                  by='CAS')
  
  setnames(dat, 'Clint', 'CLint')
  dat[param=='Clint', param:='CLint']
  dat[param=='Funbound.plasma', param:='Fub']
  dat[, param:=factor(param, 
                              levels=c('CLint',
                                       'Fub',
                                       'phys.par'))]
```
Next, set up a few things to prepare for plotting. First, change the factor levels for `poormetab` and `fup.censor` to be more informative than just TRUE and FALSE.
```{r change_factor_levels, eval=FALSE}
  
  #change poormetab factor levels to be informative
  dat[, poormetab:=factor(poormetab,
                              levels=c(FALSE, TRUE),
                              labels=c('excluded',
                                       'included'))]
  #change fup.censor factor levels to be informative
  dat[, fup.censor:=factor(fup.censor,
                               levels=c(FALSE, TRUE),
                               labels=c('<lod=lod/2',
                                        '<lod censored'))]
  #change fup.censor name is be more informative
  setnames(dat, 'fup.censor', 'Fub')
```
And set up a labeller function to more informatively label the plots by parameter.
```{r my_labeller_fun, eval=FALSE}
my_labeller_fun <- function(DF){
 if (variable=='param'){
   return(paste('i =', value))
 }else{
   return(paste(variable,value))
 }
}
```
Next, since we're going to plot measured CLint on a log scale, we need to handle the case where measured CLint was 0. It's a bit kludgy, but we'll just plot CLint = 0 as CLint = 1e-8, so that those points will actually show up on the log-scale plot.
```{r handle_clint_zero, eval=FALSE}
dat[CLint==0, CLint:=1e-8] #plot them as 1e-8 but label them as zero
#Hack the log-scaled x-axis so that CLint==0 points have some buffer space for display
CLint.breaks <- c(1e-8, 1e-6, 1e-4, 1e-2, 1e0, 1e2, 1e4)
#Label the CLint breaks
CLint.labels <- c(expression(0^{paste(' ')}), expression(10^{-6}), 
                  expression(10^{-4}),
                  expression(10^{-2}), expression(10^0),
                  expression(10^2), expression(10^4))
```
And define the colormap to use for the sensitivity values. I chose a multi-hue sequential palette from ColorBrewer, because the variations in hue make it easier to see where sensitivity index values fall into certain bins, and the yellow-green-blue palette shows up well against a gray plot background.
```{r define_sens_colormap, eval=FALSE}
ylgnbu <- RColorBrewer::brewer.pal(n=9, name="YlGnBu")
```
Finally, let's actually make the plots.
```{r global_sens_plots, eval=FALSE}
for (ecg in c('Total',
                  'Age.6.11',
                  'Age.12.19',
                  'Age.20.65',
                  'Age.GT65',
                  'BMIgt30',
                  'BMIle30',
                  'Males',
                  'Females',
                  'ReproAgeFemale')){
  for (fpb in c(TRUE)){
    dt.tmp <- dat[param %in% c('CLint',
                                       'Fub',
                                       'phys.par') &
                         ExpoCast.group==ecg &
                         fuptofub==fpb, ]
    setorder(dt.tmp,Si)
    p<-ggplot(data=dt.tmp) + 
        geom_point(aes(x=CLint,
                       y=Funbound.plasma,
                       color=Si,
                       fill=Si,
                       order=Si),
                   shape=21)+
        scale_color_gradientn(colours=ylgnbu,
                              guide='colourbar',
                              limits=c(0,1), #specify limits, breaks, and labels to get the colorbar tick marks where I want them
                              breaks=c(0,0.25,0.5,0.75,1),
                              labels=as.character(c(0,0.25,0.5,0.75,1)),
                              oob=squish,
                              name='First-order \nCss sensitivity \nto param i')+ #there may be some very small negative indexes; just treat them as zero
        scale_fill_gradientn(colours=ylgnbu,
                             guide=FALSE, #don't plot separate colorbar for fill
                             na.value=NA, #don't fill points for NAs
                             limits=c(0,1), #specify limits, breaks, and labels to get the colorbar tick marks where I want them
                             breaks=c(0,0.25,0.5,0.75,1),
                             labels=as.character(c(0,0.25,0.5,0.75,1)),
                             oob=squish)+
        scale_x_continuous(trans='log10', #use the hack to get CLint==0 points to not be clipped
                           breaks=CLint.breaks,
                           labels=CLint.labels) +
        scale_y_continuous(limits=c(0,1),
                           oob=squish) + #Funbound.plasma shouldn't be >1 -- if it is, just pretend it's 1
#         facet_grid(poormetab*Fub~param,
#                    labeller=my_labeller_fun)+
  facet_grid(poormetab*Fub~param, 
             labeller=labeller(poormetab=function(x) paste("CLint: PM ", x),
                                        Fub = label_both,
                                        param=function(x) paste("i = ", x))) +
        labs(x='Measured CLint (uL/min/million cells)',
             y='Measured Fub',
             title=ecg) +
        theme(legend.title=element_text(size=12), #bump up the legend text size
              legend.text=element_text(size=12),
              strip.background=element_blank()) 
  ggsave(filename=paste0('pdf_figures/',
                  paste('Si',
                        'httkpop',
                        'group',
                        ecg,
                        'model',
                        model,
                        "FuptoFub",
                        fpb,
                        sep='_'),
                  '.pdf'),
         plot=p,
         height=8.5,
         width=11)
print(p)
    }
}
```

We'd also like to plot the sensitivity analysis results for independent Monte Carlo -- both showing the same data as for correlated Monte Carlo (sensitivity to the physiological parameters as a group), and showing sensitivity to each of the physiological parameters independently.

So first, we read in the independent MC data.

```{r, eval=FALSE}
#Initialize list of data.tables
dt.list <- vector(mode="list", length=nrow(pmfc))

for (i in 1:nrow(pmfc)){
  dt.list[[i]] <- readRDS(paste0('data/',
                   'sens_glenisaacs_nhanes_', 
                   'allchems_',
                   'indepMC_',
                   model,
                   '_',
                   css.method,
                   '_fup_censor_',
                   pmfc[i, 'fup.censor'],
                   '_poormetab_',
                   pmfc[i, 'poormetab'],
                    ifelse(pmfc[i, 'fuptofub'], "_FuptoFub", ""),
                   '.Rdata'))
  dt.list[[i]][, poormetab:=pmfc[i, 'poormetab']]
  dt.list[[i]][, fup.censor:=pmfc[i, 'fup.censor']]
  dt.list[[i]][, fuptofub:=pmfc[i, 'fuptofub']]
}

dat_indep <- rbindlist(dt.list)
setnames(dat_indep, c('chemcas', 'rn'),
         c('CAS', 'param'))
 dat_indep <- merge(dat_indep,
                  cheminfo.dt[, .(CAS,
                                  Compound,
                                  Funbound.plasma,
                                  Clint)],
                  by='CAS')
  
  setnames(dat_indep, 'Clint', 'CLint')
  dat_indep[param=='Clint', param:='CLint']
  dat_indep[param=='Funbound.plasma', param:='Fub']
  dat_indep[, param:=factor(param, 
                          levels=c('CLint',
                                   'Fub',
                                   'phys.par',
                                   unique(param)[!(unique(param) %in%  
                                                     c('CLint',
                                                       'Fub',
                                                       'phys.par'))]))]
  #change poormetab factor levels to be informative
  dat_indep[, poormetab:=factor(poormetab,
                              levels=c(FALSE, TRUE),
                              labels=c('excluded',
                                       'included'))]
  #change fup.censor factor levels to be informative
  dat_indep[, fup.censor:=factor(fup.censor,
                               levels=c(FALSE, TRUE),
                               labels=c('<lod=lod/2',
                                        '<lod censored'))]
  #change fup.censor name is be more informative
  setnames(dat_indep, 'fup.censor', 'Fub')
dat_indep[CLint==0, CLint:=1e-8] #plot them as 1e-8 but label them as zero
```

Now do the plotting.
```{r indep_params, eval=FALSE}
for (fpb in c(TRUE)){
  dt.tmp <- dat_indep[param %in% c('CLint',
                                       'Fub',
                                       'phys.par') &
                           fuptofub==fpb, ]
  setorder(dt.tmp, Si)
p<-ggplot(data=dt.tmp) + 
        geom_point(aes(x=CLint,
                       y=Funbound.plasma,
                       color=Si,
                       fill=Si,
                       order=Si),
                   shape=21)+
        scale_color_gradientn(colours=ylgnbu,
                              guide='colourbar',
                              limits=c(0,1), #specify limits, breaks, and labels to get the colorbar tick marks where I want them
                              breaks=c(0,0.25,0.5,0.75,1),
                              labels=as.character(c(0,0.25,0.5,0.75,1)),
                              oob=squish,
                              name='First-order \nCss sensitivity \nto param i')+ #there may be some very small negative indexes; just treat them as zero
        scale_fill_gradientn(colours=ylgnbu,
                             guide=FALSE, #don't plot separate colorbar for fill
                             na.value=NA, #don't fill points for NAs
                             limits=c(0,1), #specify limits, breaks, and labels to get the colorbar tick marks where I want them
                             breaks=c(0,0.25,0.5,0.75,1),
                             labels=as.character(c(0,0.25,0.5,0.75,1)),
                             oob=squish)+
        scale_x_continuous(trans='log10', #use the hack to get CLint==0 points to not be clipped
                           breaks=CLint.breaks,
                           labels=CLint.labels) +
        scale_y_continuous(limits=c(0,1),
                           oob=squish) + #Funbound.plasma shouldn't be >1 -- if it is, just pretend it's 1
#         facet_grid(poormetab*Fub~param,
#                    labeller=my_labeller_fun)+
  facet_grid(poormetab*Fub~param, 
             labeller=labeller(poormetab=function(x) paste("CLint: PM ", x),
                                        Fub = label_both,
                                        param=function(x) paste("i = ", x))) +
        labs(x='Measured CLint (uL/min/million cells)',
             y='Measured Fub',
             title='Independent MC') +
        theme(legend.title=element_text(size=12), #bump up the legend text size
              legend.text=element_text(size=12),
              strip.background=element_blank()) 
  ggsave(filename=paste0('pdf_figures/',
                  paste('Si',
                        'indepMC',
                        'model',
                        model,
                        "FuptoFub",
                        fpb,
                        sep='_'),
                  '.pdf'),
         plot=p,
         height=8.5,
         width=11)
print(p)
}
```


And plot each physiological parameter independently.

```{r indep_phys, eval=FALSE}
for (fpb in c(TRUE)){
  dt.tmp <- dat_indep[param %in% c('million.cells.per.gliver',
                                        'Qgfrc',
                                        'Qtotal.liverc',
                                        'Vliverc') &
                           fuptofub==fpb, ]
  setorder(dt.tmp, Si)
p<-ggplot(data=dt.tmp) + 
        geom_point(aes(x=CLint,
                       y=Funbound.plasma,
                       color=Si,
                       fill=Si,
                       order=Si),
                   shape=21)+
        scale_color_gradientn(colours=ylgnbu,
                              guide='colourbar',
                              limits=c(0,1), #specify limits, breaks, and labels to get the colorbar tick marks where I want them
                              breaks=c(0,0.25,0.5,0.75,1),
                              labels=as.character(c(0,0.25,0.5,0.75,1)),
                              oob=squish,
                              name='First-order \nCss sensitivity \nto param i')+ #there may be some very small negative indexes; just treat them as zero
        scale_fill_gradientn(colours=ylgnbu,
                             guide=FALSE, #don't plot separate colorbar for fill
                             na.value=NA, #don't fill points for NAs
                             limits=c(0,1), #specify limits, breaks, and labels to get the colorbar tick marks where I want them
                             breaks=c(0,0.25,0.5,0.75,1),
                             labels=as.character(c(0,0.25,0.5,0.75,1)),
                             oob=squish)+
        scale_x_continuous(trans='log10', #use the hack to get CLint==0 points to not be clipped
                           breaks=CLint.breaks,
                           labels=CLint.labels) +
        scale_y_continuous(limits=c(0,1),
                           oob=squish) + #Funbound.plasma shouldn't be >1 -- if it is, just pretend it's 1
#         facet_grid(poormetab*Fub~param,
#                    labeller=my_labeller_fun)+
  facet_grid(poormetab*Fub~param, 
             labeller=labeller(poormetab=function(x) paste("CLint: PM ", x),
                                        Fub = label_both,
                                        param=function(x) paste("i = ", x))) +
        labs(x='Measured CLint (uL/min/million cells)',
             y='Measured Fub',
             title='Independent MC') +
        theme(legend.title=element_text(size=12), #bump up the legend text size
              legend.text=element_text(size=12),
              strip.background=element_blank()) 
  ggsave(filename=paste0('pdf_figures/',
                  paste('Si',
                        'indepMC',
                        'physparams',
                        'model',
                        model,
                        "FuptoFub",
                        fpb,
                        sep='_'),
                  '.pdf'),
         plot=p,
         height=8.5,
         width=11)
print(p)
}
```
