---
title: "Ring et al. 2017 Serum creatinine spline fits and residuals"
author: "Caroline Ring"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Serum creatinine spline fits and residuals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, 
                      comment = '#>', 
                      fig.width=11.25, 
                      fig.height=19)
```

```{r, eval = FALSE}
library(httk)
library(survey)
library(data.table)
library(ggplot2)
all.reth <- levels(nhanes_mec_svy$variables[, ridreth1])
all.gendr <- levels(nhanes_mec_svy$variables[, riagendr])
```

Plot serum creatinine vs. age with spline fits.
```{r serumcreat_spline, eval=FALSE}
par(mfcol=c(5,2),
    mar=c(2,2,2.5,2)+0.1,
    oma=c(4,4,4,0),
    mgp=c(1,1,0))

for (gendr in all.gendr){
  for (r in all.reth){
    grsub <- subset(nhanes_mec_svy, riagendr==gendr & ridreth1==r & !is.na(ridexagm) & !is.na(loglbxscr))
    svyplot(loglbxscr~ridexagm, grsub, style="transparent", 
            basecol="gray50", xlab=NA, ylab=NA, cex.axis=1.5)
    lines(sort(unique(grsub$variables[,ridexagm])), predict(spline_serumcreat[gender==gendr & reth==r,
                                                               sc_spline[[1]]],
                                  sort(unique(grsub$variables[,ridexagm])))$y)
    title(paste(gendr, r), line=0.5, cex.main=1.5)
  }
}
title(xlab="Age, months",
      ylab="Log serum creatinine, log g/dL",
      outer=TRUE,
      line=2, cex.lab=2)
title(main="Serum creatinine vs. age, with spline fits",
      outer=TRUE,
      line=1, cex.main=2)
```

```{r serumcreat_spline_pdf, echo=FALSE, eval=FALSE}
pdf("pdf_figures/serumcreat_spline.pdf", width=8.5, height=11)
par(mfcol=c(5,2),
    mar=c(2,2,2.5,2)+0.1,
    oma=c(4,4,4,0),
    mgp=c(1,1,0))

for (gendr in all.gendr){
  for (r in all.reth){
    grsub <- subset(nhanes_mec_svy, riagendr==gendr & ridreth1==r & !is.na(ridexagm) & !is.na(loglbxscr))
    svyplot(loglbxscr~ridexagm, grsub, style="transparent", 
            basecol="gray50", xlab=NA, ylab=NA, cex.axis=1.5)
    lines(sort(unique(grsub$variables[,ridexagm])), predict(spline_serumcreat[gender==gendr & reth==r,
                                                               sc_spline[[1]]],
                                  sort(unique(grsub$variables[,ridexagm])))$y)
    title(paste(gendr, r), line=0.5, cex.main=1.5)
  }
}
title(xlab="Age, months",
      ylab="Log serum creatinine, log g/dL",
      outer=TRUE,
      line=2, cex.lab=2)
title(main="Serum creatinine vs. age, with spline fits",
      outer=TRUE,
      line=1, cex.main=2)
dev.off()
```

Plot log serum creatinine residuals vs. age.
```{r serumcreat_resid, eval=FALSE}
par(mfcol=c(5,2),
    mar=c(2,2,2.5,2)+0.1,
    oma=c(4,4,4,0),
    mgp=c(1,1,0))

for (gendr in all.gendr){
  for (r in all.reth){
    grsub <- subset(nhanes_mec_svy, riagendr==gendr & ridreth1==r & !is.na(ridexagm) & !is.na(loglbxscr))
    grsub <- update(grsub, loglbxscr_resid=loglbxscr- 
                      predict(spline_serumcreat[gender==gendr & 
                                                    reth==r,
                                                  sc_spline[[1]]],
                              grsub$variables[,ridexagm])$y)
    svyplot(loglbxscr_resid~ridexagm, grsub, style="transparent", 
            xlab=NA, ylab=NA, cex.axis=1.5)
    title(paste(gendr, r), line=0.5, cex.main=1.5)
    }
  }
title(xlab="Age, months",
      ylab="Log serum creatinine residual, log g/dL", 
      outer=TRUE,
      line=2, cex.lab=2)
title(main = "Serum creatinine residuals vs. age",
      outer=TRUE, line=1, cex.main=2)
```

Make QQ plots of log serum creatinine residual quantiles vs. normal quantiles.
```{r, eval=FALSE}
par(mfcol=c(5,2),
    mar=c(2,2,2.5,2)+0.1,
    oma=c(4,4,4,0),
    mgp=c(1,1,0))

for (gendr in all.gendr){
  for (r in all.reth){
    grsub <- subset(nhanes_mec_svy, riagendr==gendr & ridreth1==r & !is.na(ridexagm) & !is.na(loglbxscr))
    grsub <- update(grsub, logscresid=loglbxscr- 
                      predict(spline_serumcreat[gender==gendr & 
                                                    reth==r,
                                                  sc_spline[[1]]],
                              grsub$variables[,ridexagm])$y)
qprobs <- ppoints(n=length(grsub$prob))
norm.q <- qnorm(p=qprobs)
logscresid.q <- svyquantile(~logscresid, 
                            design=grsub,
                            quantiles=qprobs)
plot(y=logscresid.q,
     x=norm.q, #Plot weighted points
     type='p',
     col="gray70",
     xlab=NA, ylab=NA)
#Add qqline: passing through first and third quartiles
x <- qnorm(p=c(0.25, 0.75))
y <- as.vector(svyquantile(~logscresid, 
                           design=grsub,
                           quantiles=c(0.25, 0.75)))
m <- diff(y)/diff(x)
intercept <- -m*x[1]+y[1]
abline(a=intercept, b=m)
title(paste(gendr, r), line=0.5, cex.main=1.5)
}
}
title(xlab='Normal quantiles',
     ylab='log SC resid quantiles',
     outer=TRUE,
     line=2,
     cex.lab=2)
```

Clearly, the log residuals aren't normally distributed for most combinations of race and gender. For this reason, I decided to use a kernel density estimate of the residual distribution as a way to non-parametrically estimate the residual distribution, rather than assuming a particular theoretical parametric distribution.

# KDE for residuals
Here are the KDEs for the log serum creatinine residuals for each gender and race/ethnicity combination.
```{r, eval=FALSE}

par(mfcol=c(5,2),
    mar=c(2,2,2.5,2)+0.1,
    oma=c(4,4,4,0),
    mgp=c(1,1,0))

for (gendr in all.gendr){
  for (r in all.reth){
     grsub <- subset(nhanes_mec_svy, riagendr==gendr & ridreth1==r & !is.na(ridexagm) & !is.na(loglbxscr))
    grsub <- update(grsub, logscresid=loglbxscr- 
                      predict(spline_serumcreat[gender==gendr & 
                                                    reth==r,
                                                  sc_spline[[1]]],
                              grsub$variables[,ridexagm])$y)
     tmp.kde <- ks::kde(x=grsub$variables[, logscresid],
                   w=grsub$variables[, wtmec6yr])
    plot(tmp.kde, xlab=NA, ylab=NA, xlim=c(-1.72,2.7), drawpoints=TRUE, col.pt="black")
    title(paste(gendr, r), line=0.5, cex.main=1.5)
    }
  }
title(xlab='Log serum creatinine residual',
     ylab='KDE density',
     outer=TRUE,
     line=2,
     cex.lab=2)
```