---
title: "HTTK Fup LOD test"
author: "Caroline Ring"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
data_dir <- 'C:/Users/cring/OneDrive - Environmental Protection Agency (EPA)/httk fup LOD/'
```


# Introduction

This Rmd document contains analysis to check the effect of lowering the default Fup limit of detection and/or minimum value in httk.


# Planned analyses

Test each of the models in turn: one-compartment, three-compartment, sum-of-clearances, and PBTK.

Can also try gestational PBTK.

Test the "default" (non-Monte Carlo) solutions, and also test the HTTK-Pop solutions.

Test both analytical Css and numerical Css solutions.

Test with restrictive clearance both on and off.

Can do this using `calc_css()` (and `calc_analytic_css()`).

The `...` argument to `calc_css()` is the additonal arguments passed to the model solver, which includes `minimum.Funbound.plasma`.


Plan:

Loop over models

For each model:
Get available chemicals
Loop over chemicals

Call parametrize function
Call `calc_analytic_css()`
Call `calc_css()`
Compile the parameters and Css output into a data table


A quick test: here's a compound, "Amitraz", whose Fup distribution is "3.02e-08,1.85e-13,0.00627". Compare parameters for the 1compartment model.

```{r}

#with a smaller Fup min
pars1 <- parameterize_1comp(chem.name = "Amitraz",
                            minimum.Funbound.plasma = 1e-12)

#with default Fup min
pars2 <- parameterize_1comp(chem.name = "Amitraz")

parsdf <- rbind(as.data.frame(pars1),
                as.data.frame(pars2))

parsdf
```

Compare results calling `calc_vdist()` directly 

```{r}
calc_vdist(chem.name = "Amitraz", minimum.Funbound.plasma = 1e-12)
calc_vdist(chem.name = "Amitraz")
```

Try calc_css?

```{r}
suppressWarnings(css1 <- calc_css(chem.name = "Amitraz", model = "1compartment"))

suppressWarnings(css2 <- calc_css(chem.name = "Amitraz", model = "1compartment", minimum.Funbound.plasma = 1e-12)
)

cssdf <- rbind(as.data.frame(css1), as.data.frame(css2))

cssdf
```

OK. now that the minimum is being propagated through `calc_css()`, we see a big difference in Css. Wow, is it really that big? The volume of distribution change wasn't that huge. What baout kelim?

```{r}
suppressWarnings(kelim1 <- calc_elimination_rate(chem.name = "Amitraz"))

suppressWarnings(kelim2 <- calc_elimination_rate(chem.name = "Amitraz", minimum.Funbound.plasma = 1e-12))

c(kelim1, kelim2)
```

Yes, okay, wow, the elimination rate really does change a lot (get a lot slower) when we allow a very small Fup.

Let's check other models with this chemical.

PBTK

```{r}
suppressWarnings(css1 <- calc_css(chem.name = "Amitraz", model = "pbtk"))

suppressWarnings(css2 <- calc_css(chem.name = "Amitraz", model = "pbtk", minimum.Funbound.plasma = 1e-12)
)

suppressWarnings(css3 <- calc_css(chem.name = "Amitraz", model = "pbtk", restrictive.clearance = FALSE))

suppressWarnings(css4 <- calc_css(chem.name = "Amitraz", model = "pbtk", minimum.Funbound.plasma = 1e-12, restrictive.clearance = FALSE)
)

cssdf <- rbind(as.data.frame(css1), as.data.frame(css2),
               as.data.frame(css3), as.data.frame(css4))

cssdf
```


I find it very curious that the results are identical with and without restrictive clearance. Seems like the restrictive clearance argument might not be getting passed?

3-compartment

```{r}
suppressWarnings(css1 <- calc_css(chem.name = "Amitraz", model = "3compartment"))

suppressWarnings(css2 <- calc_css(chem.name = "Amitraz", model = "3compartment", minimum.Funbound.plasma = 1e-12)
)

cssdf <- rbind(as.data.frame(css1), as.data.frame(css2))

cssdf
```

Note that we can't do fetal PBTK because for some reason, `calc_css()` call to `solve_model()` crashes beacuse it apparently requests too much precision. This is beyond the scope of this problem so I will just drop fetal PBTK from this analysis.

For the non-population (default individual) simulations, I believe the only chemicals that will show a difference are the ones whose median Funbound.plasma values were less than the default minimum.Funbound.plasma. So I will save the initial Funbound.plasma values. 


```{r}
fup_small <- 1e-12

model_vect <- c("1compartment",
                "3compartment",
                "pbtk")

restrictive_vect <- c(TRUE,
                      FALSE)

fup_min_vect <- c(1e-4,
                  fup_small)

run_opts <- expand.grid(model = model_vect,
                        restrictive = restrictive_vect,
                        fup_min = fup_min_vect)

#get chemicals for each model

run_tbl <- run_opts %>% dplyr::rowwise(model, restrictive, fup_min) %>%
  dplyr::reframe(get_cheminfo(info = c("DTXSID",
                                              "Compound",
                                              "Funbound.plasma"),
                                     species = "Human",
                                     model = model,
                                     median.only = TRUE,
                                     suppress.messages = TRUE))

#function to apply for each chemical and set of options
calc_chem_css <- function(this_model,
                      this_restrictive,
                      this_fup_min,
                      this_dtxsid
                      ){
               
               #get css
               this_chem_css <- as.data.frame(
                 suppressWarnings(
                   calc_css(dtxsid = this_dtxsid,
                                        model = this_model,
                                     minimum.Funbound.plasma = this_fup_min,
                                     restrictive.clearance = this_restrictive,
                                     suppress.messages = TRUE)
                 )
               )

               return(this_chem_css)
}


system.time(run_out <- run_tbl %>% 
              dplyr::slice_head(n = 100) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(calc_chem_css(this_model = model,
                          this_restrictive = restrictive,
                          this_fup_min = fup_min,
                          this_dtxsid = DTXSID)))
```

This benchmark for the first 100 rows suggests that we are going to need to use multidplyr, since it is running at approximately 2 seconds per row and we have 11,484 rows.

```{r}
library(multidplyr)
cluster <- multidplyr::new_cluster(parallel::detectCores()-1)

#load httk on each cluster
multidplyr::cluster_send(cluster,
             devtools::load_all("../httk"))

multidplyr::cluster_copy(cluster, "calc_chem_css")

#evaluate calc_css in parallel
system.time(run_out <- run_tbl %>% 
              dplyr::slice_head(n = 100) %>%
  dplyr::rowwise() %>%
    multidplyr::partition(cluster = cluster) %>%
  dplyr::mutate(calc_chem_css(this_model = model,
                          this_restrictive = restrictive,
                          this_fup_min = fup_min,
                          this_dtxsid = DTXSID)) %>%
    dplyr::collect()
  
  )

parallel::stopCluster(cluster)
```

This takes about 70 seconds, rather than more than 200. A decent speed-up.

So then let's run the whole thing.


```{r}

#evaluate calc_css in parallel
system.time(run_out <- run_tbl %>% 
  dplyr::rowwise() %>%
    multidplyr::partition(cluster = cluster) %>%
  dplyr::mutate(calc_chem_css(this_model = model,
                          this_restrictive = restrictive,
                          this_fup_min = fup_min,
                          this_dtxsid = DTXSID)) %>%
    dplyr::collect()
  
  )

#save
saveRDS(run_out,
        paste0(data_dir,
               "fup_lod_css_analysis.Rds"))


```

Nice! With the parallel processing, the system time was
   user  system elapsed 
   4.75   14.86 5608.76
   
   

Try a plot:

```{r}
plot_tbl <- run_out %>% dplyr::mutate(fup_min_cat = factor(fup_min, levels = c(1e-4, 1e-12), labels = c("old_fup_min", "new_fup_min")))

plot_tbl2 <- plot_tbl %>% tidyr::pivot_wider(id_cols = c(model, restrictive, Compound, DTXSID, Human.Funbound.plasma), names_from = fup_min_cat, values_from = c(avg, frac, max, the.day))
```

```{r}
ggplot(plot_tbl2) +
  geom_point(aes(x = avg_old_fup_min,
                 y = avg_new_fup_min,
                 color = Human.Funbound.plasma)) +
  geom_abline(intercept = 0, slope = 1) +
  scale_color_viridis_c(trans = "log10") +
  scale_x_log10() +
  scale_y_log10() +
  annotation_logticks(sides = "bl") +
  facet_grid(rows = vars(model),
             cols = vars(restrictive),
             labeller = labeller(.rows = label_value,
                                 .cols = label_both)) +
  xlab("Css with Fup min = 1e-4") +
  ylab("Css with Fup min = 1e-12")
```



