---
title: "HTTK Fup LOD test"
author: "Caroline Ring"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This Rmd document contains analysis to check the effect of lowering the default Fup limit of detection and/or minimum value in httk.


# Planned analyses

Test each of the models in turn: one-compartment, three-compartment, sum-of-clearances, and PBTK.

Can also try gestational PBTK.

Test the "default" (non-Monte Carlo) solutions, and also test the HTTK-Pop solutions.

Test both analytical Css and numerical Css solutions.

Test with restrictive clearance both on and off.

Can do this using `calc_css()` (and `calc_analytic_css()`).

The `...` argument to `calc_css()` is the additonal arguments passed to the model solver, which includes `minimum.Funbound.plasma`.


Plan:

Loop over models

For each model:
Get available chemicals
Loop over chemicals

Call parametrize function
Call `calc_analytic_css()`
Call `calc_css()`
Compile the parameters and Css output into a data table


A quick test: here's a compound, "Amitraz", whose Fup distribution is "3.02e-08,1.85e-13,0.00627". Compare parameters for the 1compartment model.

```{r}

#with a smaller Fup min
pars1 <- parameterize_1comp(chem.name = "Amitraz",
                            minimum.Funbound.plasma = 1e-12)

#with default Fup min
pars2 <- parameterize_1comp(chem.name = "Amitraz")

parsdf <- rbind(as.data.frame(pars1),
                as.data.frame(pars2))

parsdf
```

Compare results calling `calc_vdist()` directly 

```{r}
calc_vdist(chem.name = "Amitraz", minimum.Funbound.plasma = 1e-12)
calc_vdist(chem.name = "Amitraz")
```

Try calc_css?

```{r}
css1 <- calc_css(chem.name = "Amitraz", model = "1compartment")

css2 <- calc_css(chem.name = "Amitraz", model = "1compartment", minimum.Funbound.plasma = 1e-12)

cssdf <- rbind(as.data.frame(css1), as.data.frame(css2))

cssdf
```

Again, exactly identical, suggesting that the minimum is not being propagated correctly.

Checking the code for `calc_css()`, that's true -- it's not being passed to the parameterize function. So I'm adding an explicit `minimum.Funbound.plasma` argument (because that one will get passed to both the parameterize and the model solver), as well as an `parameterize.args` argument (as for `calc_analytic_css()`) to pass any other parameterize arguments.


```{r}
fup_small <- 1e-12

model_vect <- c("1compartment",
                "3compartment",
                "3compartmentss",
                "pbtk")

lapply(model_vect,
       function(this_model){
         model_chems <- get_cheminfo(info = c("DTXSID",
                                              "Compound",
                                              "Funbound.plasma")
                                     species = "Human",
                                     fup.lod.default = fup_small,
                                     model = this_model)
         
         param_fun <- httk:::model.list[[this_model]][["parameterize.func"]]
         
         
         params_restrict <- sapply(model_chems$DTXSID,
                          function(this_dtxsid){
                             do.call(param_fun,
                                     args = list(dtxsid = this_dtxsid,
                                                 minimum.Funbound.plasma = fup_small,
                                                 restrictive.clearance = TRUE)
                           )
                          }
          
       })
```

We can probably just publish one of the modelsi n the main text and put the rest in supplemental



